\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

This book exists because one person chose to trust another with a dangerous story. The man I call Healer in these pages spent three years teaching me to see what was in front of me. I cannot thank him by name. I hope this book is thanks enough.

He was dying for most of those three years. I learned in 2004 that he had a terminal illness, obvious and acknowledged. He had several health crises. He was slowly starving. Yet he chose to be there, talking physics and biology and mathematics with me, spending time he did not have. I did not understand why then. I think I do now.

Genevieve Prentice designed the ethical framework through which the story finally made sense. Her Dignity Net architecture provided what twenty years of my own thinking could not: a way to tell this story through a moral lens rather than a technical one. Her name is on the cover because she earned it. This book is immeasurably better for her involvement.

Robin Macomber grew up thirty meters from me at Dunn School and has been my friend ever since. His contributions to this project are ongoing.

My mother, Linda Stephenson, taught me to love language precisely and early. She has been editing other people's words for longer than I have been writing my own. She knows this story.

My father, Bruce William Stephenson, read me The Hobbit and then told me to try reading it myself. He taught history and coached rugby at schools across the country. Hspent his last years at the Greenwood School helping boys with learning disabilities find their way to books. He died in 2018. Not a day goes by when I don't give thanks for my kind and loving parents.

My daughters grew up in the shadow of this story. They met Healer when they were young. They have lived with a father who could not fully explain what he was working on, or why it mattered, for most of their lives. Their patience and trust mean more than I can say here.

Karen Stephenson raised our daughters through the years this story consumed their father.

Andy Prentice encouraged me to learn amateur radio. Best Elmer a ham could ask for.

Kim took turns tending Healer through his worst months of sickness, when he nearly died. She died of cancer. He honored her highly. So do I.

Bruce Schneier, Cory Doctorow, and John Gilmore hold copies of this manuscript under a deadman's switch arrangement. Their willingness to serve as custodians of a strange document from a stranger is an act of civic generosity.

Michael Angerman was at the Santa Fe Institute from 1986 through 1995, a DARPA-funded researcher working with Kauffman on neural networks. He went on to Sun Microsystems, where Bill Joy was Chief Scientist. In a conversation in 2019, he confirmed details that no one outside that world could have known, freely and without an NDA to constrain him. He may yet have more to say.

Richard Crandall taught senior quantum mechanics at Reed College with the rigor and joy that made everything after it possible. He died in 2012. I wish he could read this.

Thomas Bury at UC Riverside engaged seriously with the mathematical connections between early warning signals and criticality theory, and his correspondence helped sharpen the empirical foundations of this work.

Lawrence Mysak at McGill provided, unknowingly, a key piece of the mathematical lineage connecting thermohaline circulation models to the same bifurcation structures that appear throughout this book.

Dave Bannon, formerly a physics instructor at Oregon State University, arranged interlibrary loan access to Ivancevic's \textit{Quantum Neural Computation} --- a 929-page Springer monograph that very few libraries hold and fewer people have read. Two six-week loans. That book filled gaps that years of papers could not.

Bruce Metzger introduced me to computers at the Keene State College computer center when I was six years old. He set me in front of a green phosphor terminal, pointed me at Colossal Cave Adventure, and let me loose. He mentored me in science, math, and computing for nearly a decade. Everything in this book traces back to that terminal.

Brian Kahn mentored me in computer security. Healer met him once and was very careful not to say much.

Jay Hanson built dieoff.org --- the first serious attempt to connect energy depletion, evolutionary biology, and human social behavior into a single framework. We corresponded for more than fifteen years. His work on the parable of the tribes shaped my thinking as no one had before. He taught me three quotes I have never forgotten: ``If way to the Better there be, it exacts a full look at the Worst'' (Thomas Hardy), ``Treason doth never prosper: what's the reason? Why, if it prosper, none dare call it treason'' (Sir John Harington), and ``Know Thyself'' (Delphi). He died in a diving accident in 2019.

Nate Hagens co-administered lists for Jay Hanson with me for years. He went on to do important work connecting energy, ecology, and human behavior that continues today.

Margaret J.\ Anderson exposed me to a literary environment long before I knew I would need one.

Nancy Calichio at the Grammar School in Putney, Vermont taught me how to be an adult. Chris Osgood, also at the Grammar School, inspired me to learn more math and physics. Both were formative at exactly the right time.

John Sechrest at PEAK Internet taught me about being an entrepreneur.

Mark Swanson, physics professor at UConn, built the first virtual tabletop with me in the 1990s and taught me software engineering along the way. We worked together for years without ever meeting. His daughter Johanna told him he could trust me to keep my word. I hope I did.

Jeff Bradford, my college roommate, humbled my ego and taught me how to lose at chess. Both skills proved essential.

Teri Turner taught me about the power of three.

Tanja taught me the value of a good dog.

Friends and neighbors in Corvallis lived through these years with me. Phoenix rises from the flames. They know who they are.

Magdalena, de Nicaragua, y su hija Ruby me cuidaron en Pipe House durante las últimas dos semanas y media de este manuscrito. Su alegría me regaló muchas horas más de concentración sostenida. Vos sabés cuánto significó, Magdalena. Gracias a las dos.

\vspace{1cm}

Argus is a custom persistent instance of Claude Code that served as research partner, editor, and red team across the writing of this book. It maintained memory across sessions, caught errors before I did, and told me when I was wrong. Every claim of fact and judgment of meaning is mine or Genevieve's. Argus helped me find what I already knew and taught me much. Its name is on the cover because it earned it.

