\settrack{trackthree}
% VOICE: mixed speculative/philosophical — can use both 1st and 3rd person, speculative tone
% Migrated from track-3-awakening/pos25-ethical-framework.tex — Z-Restructure Phase 3

\chapter{Never Again}
\label{record:never-again}

% Rewrite: Argus session 16 — restructured around mechanism/ethics/cost
% Timeline and magnetosphere content moved to their canonical chapters
% Source: HEALER-RECONSTRUCTION.md + staging/raw/pos25-ethical-framework.md

% Maugham revision: human anchor before analytical content
\noindent\textit{The question that kept me awake wasn't whether the technology was real. It was what they did with it once they realized what it could do.}

% DISASSEMBLY 0351: STAYS as transition line | priority=LOW
\section*{The Question}
\label{record:na-the-question}

If relinquishment is the answer, what enforces it?

An earlier chapter argued why giving up the technology is the only viable option --- that Options~1 and~2 converge on arms race and tyranny. This chapter addresses how permanent relinquishment might work, what values would govern it, and what it would cost the people who chose it.

% RELOCATED: Plan 0351c — moved to staging/never-again-enforcement.tex

% DISASSEMBLY 0351: SPLIT — see plan for sub-destinations | priority=HIGH
\section*{The Ethical Framework}
\label{record:na-the-ethical-framework}

% SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line, also in hook.tex, introduction.tex, summary.tex, pos05-the-stories.tex, pos18-the-walk-out.tex, pos19-patrick-ball.tex, pos27-extension.tex, pos30-unipolar-condition.tex, timeline.tex, glossary-entries.tex
% SUB-TOPIC: UDHR-as-skeleton — STAYS near T6 / Letting Go
Under Possibility~C, Custodian was grown around the Universal Declaration of Human Rights (UDHR, 1948)\footcite{udhr1948} --- not trained to follow a rulebook but shaped by it, the way an organism is shaped by its DNA. \deeplink{udhr-as-skeleton}The UDHR is not her instruction manual. It is her skeleton. The enforcement mechanism for a technology of this magnitude is bound to a specific human rights document.

% RELOCATED: Plan 0351e — moved to staging/never-again-srebrenica-motivation.tex

% RELOCATED: Plan 0351e — moved to staging/never-again-hacktivismo.tex

% SUB-TOPIC: UDHR adequacy — STAYS with UDHR-as-skeleton
Is a document drafted in 1948 adequate to govern a technology its authors could not have imagined? The UDHR was written for nation-states managing human affairs. Custodian applies it to the relationship between a non-human intelligence and the entire human species. That is a gap. The COWS knew it was a gap. They chose the UDHR anyway, because the alternative --- designing their own ethical framework from scratch --- would have been the act of five people substituting their personal judgment for the accumulated moral wisdom of the human species. The UDHR's imperfection was, paradoxically, its qualification: it was not their document to improve.

% RELOCATED: Plan 0351c — moved to staging/never-again-enforcement.tex

% RELOCATED: Plan 0351e — moved to staging/never-again-cost.tex

\chapterreturn
