% Dossier: The Handler
% Split from appendix/dossier.tex — Part B moved to Part I after The Departure
% Placement: after pos07 (The Departure) — reader has spent 4 chapters with Healer, now he's gone

\settrack{trackbridge}

\chapter*{The Handler}
\addcontentsline{toc}{chapter}{The Handler}
\label{record:handler}

\begin{quote}\small
\textit{The following is assembled from two sources: Tom Clancy's character Timothy Hanley in} Rainbow Six \textit{(1998), which Healer told me was modeled on him after Clancy interviewed him in person; and my own observations from 2.7 years of close contact. Clancy got the military background right. He left out the interesting parts.}
\par\raggedleft --- Bruce Stephenson, 2026\end{quote}

\vspace{1cm}

\begin{center}
\textsc{Five Eyes --- Operator Profile (Approximation)}\\[0.3em]
\textsc{Classification: \textbf{[REDACTED]} / HUMINT / REL TO FVEY}
\end{center}

\vspace{0.5cm}

\noindent\textbf{File No:} \textbf{[REDACTED]}\\
\textbf{Subject:} \textbf{[REDACTED]} (operational alias: ``Healer'')\\
\textbf{Nationality:} Australian\\
\textbf{DOB:} 14 April 1965 (age 38 at assessment)\\
\textbf{Cover Legend:} Walkabout veteran, humanitarian worker\\
\textbf{Prepared for:} \textbf{[REDACTED]}\\
\textbf{Source classification:} Character biography from published fiction (Clancy, 1998) supplemented by subject \textsc{Stephenson}'s direct observation (2003--2006). Items marked \textsc{[clancy]} from novel. Items marked \textsc{[observed]} from student. Items marked \textbf{[REDACTED]} are neither.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{I. Biographical Summary --- per Clancy}
\label{record:hd-biographical-summary-per-clancy}

\textsc{[clancy]} White male, born 14 April 1965, \textbf{Margaret River, Western Australia} --- wine country. Father: winery foreman. Raised rural.

\textsc{[observed]} Subject is of mixed European/M\={a}ori ancestry. Clancy's description is incomplete.

\textsc{[clancy]} Australian Defence Forces Academy (ADFA), 1983--1987. Graduated with distinction.

\textsc{[clancy]} \textbf{1st Squadron, Special Air Service Regiment (SASR).} Australia's tier-one special operations unit. Selection rate: $\sim$15\%. Passed on first attempt.

\textsc{[clancy]} \textbf{Australian Intelligence Corps}, 1993--1996. Direct action to intelligence operations.

\textsc{[clancy]} Cross-trained with \textbf{US Delta Force} and \textbf{British 22 SAS}. Counter-terrorism specialist. Selected for \textsc{rainbow} --- multinational counter-terrorism unit.

\textsc{[observed]} Among many other deployments, subject participated in the compromised Bravo Two Zero operation. In published accounts he is known as Steven ``Legs'' Lane, who supposedly died of hypothermia after swimming across the Euphrates. He and another survived by finding a source of warmth after their perilous swim. His ``death'' freed him for classified work.

\textsc{[observed]} In 1996, subject stood on the summit of K2 and resolved to use his abilities for something other than ``killing bad guys.'' He anchored this vow to the Universal Declaration of Human Rights. This is the origin point of the Relinquishment.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{II. What Clancy Left Out}
\label{record:hd-what-clancy-left-out}

\textbf{[REDACTED]}

\textbf{[REDACTED]}

\textbf{[REDACTED]}

\textsc{[observed]} Subject possesses deep fluency in mathematics, physics, and computational theory inconsistent with an SAS/Intelligence Corps career path. No military curriculum produces this. Subject does not explain the discrepancy.

\textsc{[observed]} Subject claims to have helped invent ``a new form of mathematics'' in collaboration with ``other very smart people.'' This mathematics has ``a new definition of zero.'' Subject does not name the mathematics, the collaborators, or the institution.

\textbf{[REDACTED]}

\textsc{[observed]} Subject demonstrated partial eidetic memory during conversations with \textsc{Stephenson}. Mutual testing confirmed. This appears to have been a baseline assessment of the student's capacity before committing to extended pedagogy.

\textbf{[REDACTED]}

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{III. The Safe House}
\label{record:hd-the-safe-house}

\textsc{[clancy]} Margaret River, Western Australia. Wine country. Father: winery foreman.

\textsc{[observed]} A property in rural Australia was offered as a relocation site for \textsc{Stephenson}'s wife and daughters. The offer was not casual.

\textsc{[clancy]} Clancy encoded the location into the character's backstory. As one does.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{IV. Operational Characteristics}
\label{record:hd-operational-characteristics}

\textsc{[observed]} OPSEC: \textbf{Exceptional.} SAS + Intelligence Corps training evident in every interaction. No paper trail, no digital footprint, no photographs. \textsc{Stephenson}: ``\textsc{fantastic} OPSEC.''

\textsc{[observed]} Pedagogy: 2.7-year guided deduction across \hovertip{five scientific fields} through conversation alone. No written materials. Information delivery calibrated to student's processing capacity --- never overloading, never underwhelming. In retrospect, calibration kept student in critical region for optimal integration.

\textsc{[observed]} Subject rarely answered direct questions. Asked questions that caused the student to derive the answer independently. Not Socratic method --- more precise. Subject appeared to know exactly which question would produce which deduction.

\textsc{[observed]} Cover legend (walkabout veteran, humanitarian worker) maintained consistently. Subject genuinely performed humanitarian work. Cover was true. Also incomplete.

\textbf{[REDACTED]}

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{V. Skills Assessment}
\label{record:hd-skills-assessment-2}

{%
\small
\begin{tabular}{p{4cm} p{3.5cm} p{6cm}}
\hline
\textbf{Skill} & \textbf{Level} & \textbf{Source} \\
\hline
Special operations & \textsc{tier one} & \textsc{[clancy]} \\
Counter-terrorism & \textsc{specialist} & \textsc{[clancy]} \\
Intelligence operations & \textsc{advanced} & \textsc{[clancy]} \\
Cross-force interoperability & \textsc{confirmed} & \textsc{[clancy]} --- Delta, 22 SAS \\
Mathematics & \textbf{[REDACTED]} & \textbf{[REDACTED]} \\
Physics & \textbf{[REDACTED]} & \textbf{[REDACTED]} \\
Computational theory & \textbf{[REDACTED]} & \textbf{[REDACTED]} \\
Guided deduction pedagogy & \textsc{exceptional} & \textsc{[observed]} \\
Operational security & \textsc{exceptional} & \textsc{[observed]} \\
Partial eidetic memory & \textsc{confirmed} & \textsc{[observed]} --- mutual testing \\
Cover maintenance & \textsc{seamless} & \textsc{[observed]} \\
Languages & \textbf{[REDACTED]} & \textbf{[REDACTED]} \\
\hline
\end{tabular}
}%

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{VI. Assessment}
\label{record:hd-assessment}

\textsc{[clancy]} provides the military skeleton: ADFA, SASR, Intelligence Corps, \textsc{rainbow}. Real career path. Clancy did his homework --- or his source did.

\textsc{[observed]} provides the rest: mathematics, pedagogy, five-field convergence, ``new definition of zero.''

\textbf{[REDACTED]} is where it gets interesting.

An SAS operator with Intelligence Corps credentials who also helped invent new mathematics, who also conducts multi-year guided deductions across five scientific fields, who calibrates information delivery to keep a student at the critical boundary of processing capacity --- this is not a career path that exists in any military personnel system. Unless \textbf{[REDACTED]}.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{VII. Current Status}
\label{record:hd-current-status}

\textbf{[REDACTED]}

\vspace{1cm}\hrule\vspace{0.5cm}

\begin{flushright}\small\textit{%
Filed by: \textbf{[REDACTED]}\\
Distribution: \textbf{[REDACTED]} / FVEY EYES ONLY\\[0.5em]
Clancy, T.\ (1998). Rainbow Six. G.P.\ Putnam's Sons. pp.\ [character introduction].\\
\textsc{Stephenson}, B.\ (2003--2006). Direct observation. 2.7 years.%
}\end{flushright}

\chapterreturn
