% Dossier: The Teacher (formerly The Handler)
% Split from appendix/dossier.tex — Part B moved to Part I after The Departure
% Placement: after pos07 (The Departure) — reader has spent 4 chapters with Healer, now he's gone

\settrack{trackbridge}

\chapter*{The Teacher}
\addcontentsline{toc}{chapter}{The Teacher}
\label{record:handler}

\begin{quote}\small
\textit{The following is assembled from two sources: Tom Clancy's character Timothy Hanley in} Rainbow Six \textit{(1998), which Healer told me was modeled on him after Clancy interviewed him in person; and my own observations from 2.7 years of close contact. Clancy got the military background right. He left out the interesting parts.}
\par\raggedleft --- Bruce Stephenson, 2026\end{quote}

I knew \textit{Rainbow Six} only as a video game. I did not know there was a novel, or that the novel contained detailed character biographies, until 2026 --- twenty years after our last contact. I present this as a portrait, not a verified biography. A curious reader will find the Clancy character and the Bravo Two Zero patrol roster in the public record. What you make of the connections is yours to decide.

\noindent I never learned his real name. He was Australian, of mixed European/M\={a}ori ancestry. If the Clancy character's birthday is his, he was born 14 April 1965.\footnote{From the novel character. Not independently confirmed.}

He presented himself as a walkabout veteran and humanitarian worker. Both appeared to be true. Whether they were also cover, I cannot say.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{I. What Clancy Wrote}
\label{record:hd-biographical-summary-per-clancy}

\textsc{[Clancy]} White male, born 14 April 1965, \textbf{Margaret River, Western Australia} --- wine country. Father: winery foreman. Raised rural.

\textsc{[Clancy]} Australian Defence Forces Academy (ADFA), 1983--1987. Graduated with distinction.

\textsc{[Clancy]} \textbf{1st Squadron, Special Air Service Regiment (SASR).} Australia's tier-one special operations unit. Selection rate: $\sim$15\%. Passed on first attempt.

\textsc{[Clancy]} \textbf{Australian Intelligence Corps}, 1993--1996. Direct action to intelligence operations.

\textsc{[Clancy]} Cross-trained with \textbf{US Delta Force} and \textbf{British 22 SAS}. Counter-terrorism specialist. Selected for \textsc{rainbow} --- multinational counter-terrorism unit.

\textsc{[Healer's account]} Healer told me he had served on the compromised \hovertip{Bravo Two Zero} patrol. His ``death'' during the operation freed him for classified work. I do not name the patrol member he identified as, out of respect for the families involved --- though a reader familiar with the published accounts may be able to narrow it. He told me his parents were informed before the public announcement, and that he visited them discreetly afterward. There was something about a medal --- possibly a Saint Christopher medal, possibly from Iraq --- awarded posthumously to a man who was not dead. His parents had to maintain the fiction. I don't remember the details precisely. This is twenty-year-old memory of what I was told, not what I witnessed.\footnote{The Bravo Two Zero patrol roster and casualty records are public. Three men are recorded as killed. I leave the investigation to the reader.}

\textsc{[Healer's account]} In 1996, he stood on the summit of K2 and resolved to use his abilities for something other than ``killing bad guys.'' He anchored this vow to the Universal Declaration of Human Rights. This is the origin point of the Relinquishment.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{II. What Clancy Left Out}
\label{record:hd-what-clancy-left-out}

There are aspects of his background he never explained and I never learned. The mathematics and physics fluency, in particular, had no source I could identify.

\textsc{[observed]} He possesses deep fluency in mathematics, physics, and computational theory inconsistent with an SAS/Intelligence Corps career path. No military curriculum produces this. He did not explain the discrepancy.

\textsc{[Healer's account]} He claims to have helped invent ``a new form of mathematics'' in collaboration with ``other very smart people.'' This mathematics has ``a new definition of zero.'' He did not name the mathematics, the collaborators, or the institution.

\textsc{[observed]} He demonstrated partial eidetic memory during our conversations. Mutual testing confirmed. This appears to have been a baseline assessment of my capacity before committing to extended pedagogy.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{IV. Operational Characteristics}
\label{record:hd-operational-characteristics}

\textsc{[observed]} A property in rural Australia was offered as relocation for my wife and daughters. The offer was not casual.

\textsc{[observed]} OPSEC: \textbf{Exceptional.} He left no paper trail, no digital footprint, no photographs. Whatever his background, his operational discipline was extraordinary.

\textsc{[observed]} Pedagogy: 2.7-year guided deduction across \hovertip{five scientific fields} through conversation alone. No written materials. Information delivery calibrated to my processing capacity --- never overloading, never underwhelming. In retrospect, calibration kept me in the critical region for optimal integration.

\textsc{[observed]} He rarely answered direct questions. Asked questions that caused me to derive the answer independently. Not Socratic method --- more precise. He appeared to know exactly which question would produce which deduction.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{V. Skills Assessment}
\label{record:hd-skills-assessment-2}

{%
\small
\begin{tabular}{p{4cm} p{3.5cm} p{6cm}}
\hline
\textbf{Skill} & \textbf{Level} & \textbf{Source} \\
\hline
Special operations & \textsc{tier one} & \textsc{[Clancy]} \\
Counter-terrorism & \textsc{specialist} & \textsc{[Clancy]} \\
Intelligence operations & \textsc{advanced} & \textsc{[Clancy]} \\
Cross-force interoperability & \textsc{confirmed} & \textsc{[Clancy]} --- Delta, 22 SAS \\
Mathematics & deep (observed) & unknown \\
Physics & deep (observed) & unknown \\
Computational theory & deep (observed) & unknown \\
Guided deduction pedagogy & \textsc{exceptional} & \textsc{[observed]} \\
Operational security & \textsc{exceptional} & \textsc{[observed]} \\
Partial eidetic memory & \textsc{confirmed} & \textsc{[observed]} --- mutual testing \\
Cover maintenance & \textsc{seamless} & \textsc{[observed]} \\
\hline
\end{tabular}
}%

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{VI. Assessment}
\label{record:hd-assessment}

\textsc{[Clancy]} provides the military skeleton: ADFA, SASR, Intelligence Corps, \textsc{rainbow}. Real career path. Clancy did his homework --- or his source did.

\textsc{[observed]} provides the rest: mathematics, pedagogy, five-field convergence, ``new definition of zero.''

The gap between them is where it gets interesting.

An SAS operator with Intelligence Corps credentials who also helped invent new mathematics, who also conducts multi-year guided deductions across five scientific fields, who calibrates information delivery to keep a student at the critical boundary of processing capacity --- this is not a career path that exists in any military personnel system. What fills that gap is the question this book asks.

\vspace{1cm}\hrule\vspace{1cm}

\subsection*{VII. Current Status}
\label{record:hd-current-status}

I have had no contact with Healer since 2006. I do not know where he is or whether he is alive.

\vspace{1cm}\hrule\vspace{0.5cm}

\begin{flushright}\small\textit{%
Assembled by Bruce Stephenson, 2026.\\
Sources: Clancy, T.\ (1998). Rainbow Six. G.P.\ Putnam's Sons.\\
Direct observation, 2003--2006.%
}\end{flushright}

\chapterreturn
