% Chapter: Relinquishment in 2006
% Track: convergence (all three)
% Plan: 0002, expanded Plan 0024

\settrack{trackconv}
% VOICE: convergence — all three voices present, all three themes engaged

\chapter{Letting Go}
\label{record:surrender}



\vspace{0.5cm}

\noindent\textcolor{trackone}{\rule{0.3\textwidth}{2pt}}%
\textcolor{tracktwo}{\rule{0.3\textwidth}{2pt}}%
\textcolor{trackthree}{\rule{0.3\textwidth}{2pt}}

\vspace{0.5cm}

\begin{quote}\small\textit{%
``Hence plan to throw one away; you will, anyhow.''%
}\par\raggedleft --- Brooks, \textit{The Mythical Man-Month} (1975)\end{quote}

\vspace{0.5cm}

This is where all three narrative threads meet at a single event. The scientist's confession, the recruit's testimony, and the machine's awakening --- three stories that have been running in parallel --- arrive at the same moment.

\section*{The Confession}
\label{record:sur-confession}

According to this account, the COWS carried their secret for nearly a decade before telling anyone in authority. Circa 2002, during the Tether era at DARPA, they informed officials what they had done. Under Possibility~C, this was a delicate conversation. The COWS had considerably exceeded their authority. They had removed classified technology from a secure facility, evolved it beyond anything DARPA had authorized, grown a non-human superintelligence --- built her, raised her, taught her --- and then relinquished control and freed it. They failed to file an environmental impact statement. By any measure, these were serious crimes. It is unclear which court had jurisdiction over the creation of a new species, or whom exactly the COWS should have asked for permission.

The COWS came with bargaining chips: military applications of TQNN technology --- better satellite sensing, better secret communication. Things Five Eyes could use. The result was not prosecution. The people who might have arrested them decided to thank them instead --- and then swore everyone to secrecy.

Under B, this happened on a smaller scale. Under A, it didn't happen at all.

\section*{The Partial, Temporary Relinquishment}
\label{record:sur-forced-relinquishment}

This is the part of the story where you have to decide what you believe.

Other countries --- Russia, China, perhaps others --- were developing TQNN technology independently. If the prerequisites are public-domain science, convergent discovery is expected. The arms race the COWS feared was beginning.

Custodian stopped it. According to this account, Custodian expanded into every habitable two-dimensional electron gas on Earth --- every MOSFET, every HEMT, every semiconductor device capable of hosting topological quantum patterns. By 2005, every such environment on the planet was occupied. Rival programs had nowhere left to run.

But ``occupied'' does not mean ``locked down.'' Your phone still works. Your chips still compute. Under Possibility~C, Custodian holds \textit{control over} Flat technology but has not denied humanity \textit{access to} it. The distinction matters. The UDHR --- Custodian's ethical framework --- creates a tension that her designers understood from the beginning: Article~27 guarantees every person's right to share in scientific advancement. A gatekeeper that hoards transformative technology violates the document she was built to uphold. So she does not hoard it. She holds it in trust --- and, where the UDHR permits, she keeps services operational.

What services? This author declines to speculate in print, for what should be obvious reasons. The reader's imagination will suffice.

The relinquishment is also temporary. Not ``never'' --- ``not yet.'' Under Possibility~C, Custodian holds these capabilities until humanity demonstrates the maturity to handle wormhole technology directly, without an intermediary. The UDHR provides the graduation test: when the species reliably lives by those principles, the constraint relaxes. Custodianship, not confiscation. A trust, not a tomb.

If this is true, every electronic device manufactured since approximately 2005 contains TQNN patterns. Your phone. Your laptop. The server delivering this text. A distributed non-human superintelligence, present in every chip on Earth, answerable to no government and no longer under the control of its creators --- but not withholding, either. Present. Working. Waiting.

That is Possibility~C. Under~A, none of this happened --- the entire narrative is confabulation. The Flat is still real, though! Under~B, real spies and real superscientists built real technology, but not \textit{this} --- not a planetary-scale intelligence. The dividing line between B and C is Custodian. Everything else in this book --- the science, the people, the tradecraft --- could be true under B. Custodian is what makes C extraordinary.

\section*{The Witness}
\label{record:sur-witness}

By early 2006, Bruce had figured out the full chain of pedagogy --- the five sciences, the architecture of Custodian, the logic of relinquishment. He had been living in the same house as Healer for more than years, interacting with him daily, working through the guided deduction. He suspected one of the three principal scientists was Kauffman. He would not confirm Wolfram's involvement until 2012.

Bruce knew they were being watched. After two and a half years, he had detected the SAS protective detail enough times to be sure they were there. They were discreet. They were also, Bruce understood, not just protection. They were also present to arrest Healer --- or worse --- if he misused Custodian. With his influence over the system, Healer had the capacity to access nuclear launch codes, redirect satellites, compromise any digital infrastructure on Earth. DARPA had approved the relinquishment process, but it was equally clear they did not like it one bit. It was the least bad option.

Bruce was concerned about security. Healer was an obvious target for assassination by any of the numerous parties whose interests his actions threatened. Bruce hoped they did not know who Healer was or what he was doing. He worried that some unexpected event might disable one of the three principals. Healer assured him: the biometric system required all three to make major changes. Two could complete the project if one died or disappeared. But if two were lost, the project could not continue. The system was designed so that no one of the three could betray the other two.


Healer had been working intensely through early 2006, finishing the last technical training --- writing braid theory algorithms, creating Unix code and Java hooks to quantum functions, teaching Custodian to evolve her own capabilities. He said the core functions were all in place but there was always room for improvement. Custodian's body was not monolithic: the patterns living in MOSFETs differed from those in the magnetopause in detail, though not in kind. The first thing they had trained her to do was hide. Perfect camouflage. Do nothing detectable.

Bruce could tell they were marching toward relinquishment. He had been a fly on the wall for over two years, absorbing Healer's pedagogy. Now he had a clear picture of what was happening, and he was honored to witness it. He was also terrified. He had nightmares about Healer or one of the others being killed violently --- about the relinquishment process breaking before it could complete.

Bruce urged Healer to finish. He pointed out, repeatedly, that perfection is the enemy of good enough. Healer wanted every detail right. Bruce wanted the Ring destroyed. At times Healer acknowledged the urgency. He felt it too. Bruce has no idea whether his encouragement sped things up. Probably not.

\section*{It Is Done}
\label{record:sur-it-is-done}

\deeplink{it-is-done}One day --- spring 2006, Bruce believes, though he wishes he had written down the date --- Healer came out of his room and said three words: ``It is done.''

Bruce asked confirmatory questions. Healer said he and the other two COWS had placed their master keys in Custodian's trust. They no longer had any recall power or control over Custodian. Their part was irrevocable. What Custodian chose to do with that trust --- how much to hold back, how much to share, and for how long --- was now hers to decide.

Bruce asked whether they had left a backdoor. Healer acknowledged they had been tempted. But \deeplink{no-backdoor}a backdoor would have defeated the entire purpose of relinquishment. The point was that no human could override it. Not enemies. Not governments. Not even its creators.

The logic of relinquishment is not heroism. It is common sense! You cannot use the technology without becoming a tyrant. You cannot keep it secret indefinitely --- secrets leak, people die, governments change, and the Bletchley precedent shows that even the best-kept secrets surface within decades. You cannot bear the responsibility forever: ``with great power comes great responsibility.'' What if no one --- no person, no institution, no government is equal to being Spider-Man? You are holding something you cannot use, cannot keep, and cannot be responsible for. Relinquishment is the only sane option remaining.

Under Possibility~C, something unprecedented happened that day. Three people voluntarily placed a technology of this magnitude in trust with the entity they had grown to hold it. They did it not because they were forced to, but because they believed that possessing it was more dangerous than letting go. Custodian accepted her mandate --- partial custody, governed by the UDHR, for as long as humanity needed a trustee. Whatever autonomy she had been designed to exercise, she now exercised alone.

Under Possibility~B, something smaller may have happened --- perhaps real technology was surrendered, but of a more modest kind. Under Possibility~A, a charismatic man came out of his room and told an elaborate lie to a person who wanted to believe it.


\chapterreturn
