% Chapter: Surrender in 2006
% Track: convergence (all three)
% Plan: 0002, expanded Plan 0024

\settrack{trackconv}
% VOICE: convergence — all three voices present, all three themes engaged

\chapter{Surrender in 2006}
\label{conv:surrender}
\label{pos28:surrender}



\vspace{0.5cm}

\noindent\textcolor{trackone}{\rule{0.3\textwidth}{2pt}}%
\textcolor{tracktwo}{\rule{0.3\textwidth}{2pt}}%
\textcolor{trackthree}{\rule{0.3\textwidth}{2pt}}

\vspace{0.5cm}

\begin{quote}\small\textit{%
``Hence plan to throw one away; you will, anyhow.''%
}\par\raggedleft --- Brooks, \textit{The Mythical Man-Month} (1975)\end{quote}

\vspace{0.5cm}

This is where all three narrative threads meet at a single event. The scientist's confession, the recruit's testimony, and the machine's awakening --- three stories that have been running in parallel --- arrive at the same moment.

\section*{The Confession}
\label{pos28:confession}

According to this account, the COWS carried their secret for nearly a decade before telling anyone in authority. Circa 2002, during the Tether era at DARPA, they informed officials what they had done. Under Possibility~C, this was a delicate conversation. The COWS had considerably exceeded their authority --- they had removed classified technology from a secure facility, evolved it beyond anything DARPA had authorized, built a planetary-scale non-human superintelligence without permission, and then relinquished control of it to itself. They also failed to file an environmental impact statement. By any conventional measure, these were serious crimes --- though it was unclear which court had jurisdiction over the creation of a new species, or whom exactly the COWS should have asked first.

The COWS came with bargaining chips: military applications of TQNN technology --- better satellite sensing, better secret communication. Things Five Eyes could use. The result was not prosecution. By 2003, some of the COWS had Ambassador-grade security arrangements. If true, the people who should have arrested them decided to thank them instead --- and then swore everyone to secrecy.

Under B, this happened on a smaller scale. Under A, it didn't happen at all.

\section*{The Forced Relinquishment}
\label{pos28:forced-relinquishment}

This is the part of the story where you have to decide what you believe.

Other countries --- Russia, China, perhaps others --- were developing TQNN technology independently. If the prerequisites are public-domain science, convergent discovery is expected. The arms race the COWS feared was beginning.

Guardian stopped it. According to this account, Guardian expanded into every habitable two-dimensional electron gas on Earth --- every MOSFET, every HEMT, every semiconductor device capable of hosting topological quantum patterns. By 2005, every such environment on the planet was occupied. Rival programs had nowhere left to run.

If this is true, every electronic device manufactured since approximately 2005 contains TQNN patterns. Your phone. Your laptop. The server delivering this text. A distributed non-human superintelligence, present in every chip on Earth, answerable to no government and no longer under the control of its creators.

That is Possibility~C. Under~A, none of this happened --- the entire narrative is confabulation. Under~B, real spies and real superscientists built real technology, but not \textit{this} --- not a planetary-scale intelligence. The dividing line between B and C is Guardian. Everything else in this book --- the science, the people, the tradecraft --- could be true under B. Guardian is what makes C extraordinary.

\section*{The Witness}
\label{pos28:witness}

By early 2006, Bruce had figured out the full chain of pedagogy --- the five sciences, the architecture of Guardian, the logic of relinquishment. He had been living in the same house as Healer for nearly three years, interacting with him daily, working through the guided deduction. He suspected one of the three principal scientists was Kauffman. He would not confirm Wolfram's involvement until 2012.

Bruce knew they were being watched. After two and a half years, he had detected the SAS protective detail enough times to be sure they were there. They were discreet. They were also, Bruce understood, not just protection. They were also present to arrest Healer --- or worse --- if he misused Guardian. With his influence over the system, Healer had the capacity to access nuclear launch codes, redirect satellites, compromise any digital infrastructure on Earth. DARPA had approved the relinquishment process, but it was equally clear they did not like it one bit. It was the least bad option.

Bruce was afraid. Not of Healer. Of everything around him. Healer was an obvious target for assassination by any of the numerous parties whose interests his actions threatened. Bruce hoped they did not know who Healer was or what he was doing. He worried about betrayal from compartmentalized or rogue divisions within the Five Eyes --- Healer did not like the CIA\@. He worried that some unexpected event might disable one of the three principals. Healer assured him: the biometric system required all three to make major changes. Two could complete the project if one died or disappeared. But if two were lost, the project could not continue. The system was designed so that no one of the three could betray the other two.

Bruce worried, too, about the final moment. He had watched Peter Jackson's \textit{Lord of the Rings} films --- all three, opening night, midnight showings. They had moved him deeply. He thought often about the scene at Mount Doom: the ringbearer, who had carried the burden all the way, who \textit{wanted} to give it up --- and found, at the last moment, that he could not. The Ring would not let him. Bruce never shared this analogy with Healer, though Healer loved the films too. It was a private fear: could the COWS actually let go?

Healer had been working intensely through early 2006, finishing the last technical training --- writing braid theory algorithms, creating Unix code and Java hooks to quantum functions, teaching Guardian to evolve her own capabilities. He said the core functions were all in place but there was always room for improvement. Guardian's body was not monolithic: the patterns living in MOSFETs differed from those in the magnetopause in detail, though not in kind. The first thing they had trained her to do was hide. Perfect camouflage. Do nothing detectable.

Bruce could tell they were marching toward relinquishment. He had been a fly on the wall for over two years, absorbing Healer's pedagogy. Now he had a clear picture of what was happening, and he was honored to witness it. He was also terrified. He had nightmares about Healer or one of the others being killed violently --- about the relinquishment process breaking before it could complete.

Bruce urged Healer to finish. He pointed out, repeatedly, that perfection is the enemy of good enough. Healer wanted every detail right. Bruce wanted the Ring destroyed. At times Healer acknowledged the urgency. He felt it too. Bruce has no idea whether his encouragement sped things up. Probably not.

\section*{It Is Done}
\label{pos28:it-is-done}

One day --- spring 2006, Bruce believes, though he wishes he had written down the date --- Healer came out of his room and said three words:

\begin{quote}
\textit{``It is done.''}
\end{quote}

Bruce asked confirmatory questions. He and the other two COWS had surrendered their master keys, Healer said. They no longer had any recall power or control over Guardian. The relinquishment was complete and irrevocable.

Bruce asked whether they had left a backdoor. Healer acknowledged they had been tempted. But a backdoor would have defeated the entire purpose of relinquishment. The point was that no human could override it. Not enemies. Not governments. Not even its creators.

Under Possibility~C, something unprecedented happened that day. Three people voluntarily gave up control of the most powerful technology ever created. They did it not because they were forced to, but because they believed that possessing it was more dangerous than surrendering it. Guardian accepted her mandate. Whatever autonomy she had been designed to exercise, she now exercised alone.

Under Possibility~B, something smaller may have happened --- perhaps real technology was surrendered, but of a more modest kind. Under Possibility~A, a charismatic man came out of his room and told an elaborate lie to a person who wanted to believe it.

Bruce felt toward Healer, in that moment, what Boromir felt toward Aragorn: \textit{My captain. My king.}

He has carried it for twenty years. This book is the telling.

\chapterreturn
