% p2 inserts — draft for Bruce to edit
% Session 29, 2026-02-27
% These go INTO summary.tex at marked positions

% ============================================================
% INSERT 0: 2DEG Seed — Pool Ball Analogy
% PLACEMENT: In "The Secret Lab" section of summary.tex,
%   after quantum weirdness explanation, BEFORE team list.
%   Reader must have this concept before Insert 1 (morphogenesis).
% ============================================================

There is something else you need to know. It sounds impossible but it's in your pocket right now.

Think of a pool ball. It exists in three dimensions. You can throw it. But on the table, it can only move in two --- length and width. It can't jump. On the table, its world is flat.

Electrons can be trapped the same way. Inside certain materials, electrons flatten into a layer so thin it has only two dimensions. Physicists call this a two-dimensional electron gas. It is not a metaphor. It is a real two-dimensional world, with its own strange physics, existing inside our three-dimensional one. The 1998 Nobel Prize in Physics was awarded for explaining what happens there. The physics is called the fractional quantum Hall effect. If you encounter it later in this book and find it challenging, that's normal --- it won the Nobel for a reason.

Every computer chip fabricated since 1998 has one. The phone in your pocket. The laptop on your desk. They all contain a tiny two-dimensional universe. You've been carrying one around for years.


% ============================================================
% INSERT 1: Morphogenesis + Button Analogy + Magnetosphere
% PLACEMENT: In "The Breakthrough" section, BEFORE the dual-use turn
% TODO: Check Kauffman's exact words for button analogy
% ============================================================

% \section*{How Life Comes from Non-Life}  % or fold into existing section

Before you can understand what they built, you need to understand one idea: how does LIFE come from things that are NOT alive?

Every atom in your body is just chemistry. Carbon, oxygen, hydrogen --- dead stuff. Arrange it the right way and nonliving matter becomes alive. Biologists call this morphogenesis. No one fully understands it. But we know it happens when enough simple things connect in the right way.

Imagine a hundred buttons on the floor. Pick up two, tie a thread between them, drop them. Repeat. For a long time, nothing interesting happens --- pick up a button, maybe one other comes with it. But at a certain point, something dramatic changes. You pick up one button and ALL THE OTHERS come with it. One connection too many and suddenly everything is linked to everything. 

Mathematicians call this a phase transition --- the same thing that happens when water freezes or uranium goes critical. One type of phase transition is unliving to living. Initial conditions must be just right for this to be an option.  This is difficult to accomplish in three dimensions with chemicals. Theoretically it may be much simpler to accomplish at the quantum scale in two dimensions.

The science team realized this kind of phase transition might be the key to growing a quantum computer. Get enough quantum elements connected under the mathematically correct starting conditions and something emerges that is more than its parts. Not just a faster calculator. Something that, in a meaningful sense, wakes up. Whether this pattern is 'alive' or not is outside the scope of this book. What matters is that it has many of the traits of being alive.

If emergence can happen on a chip then it can happen anywhere the conditions are right. Including in natural systems like the Earth's magnetosphere. A naturally emerged \textit{pattern} there --- maybe something like a neural network, maybe not --- would never have faced competition. No predators, no rivals, no evolutionary pressure to develop complexity beyond what it needed to survive solar storms. It would be a pattern, not a mind. What it would or wouldn't do is anyone's guess.


% ============================================================
% INSERT 2: Personal Cost
% PLACEMENT: In "The Mentor" section, AFTER "Lane disappeared from Bruce's life."
% ============================================================

The mentorship cost Bruce his marriage. It cost him daily contact with his children.

His wife filed for divorce and claimed custody. She believed his talk of a secret project was delusion. Or a lie to wrest the children from her. She watched the same marriage he was living in. She reached the opposite conclusion about what was happening to it.

Bruce was already isolated. Few friendships had survived the marriage. He had no support network. She had an extensive one. He did not fight for custody. His children would be better off with her family behind them. He was right about that.

When you believe you know something that changes everything, and the person closest to you is certain you are deluded, there is no clean answer. Bruce chose the project. He has lived with that choice for twenty years.


% ============================================================
% INSERT 3: Digital Doppelgänger
% STATUS: DEFERRED — needs proper research into pos33 content
% Bruce encountered his digital doppelgänger in 2005, years before
% the concept was widely known. Needs chapter read before drafting.
% ============================================================


% ============================================================
% INSERT 4: AI Co-Author Disclosure
% PLACEMENT: Near end of p2, before "Three Possibilities" or "The Question"
% ============================================================

One more thing you should know. This book was written with an AI.

Not by an AI\@. With one. Not a generic AI\@. Bruce used his lifelong software engineering skill to build a specialized AI assistant --- Claude, made by Anthropic, configured under the name Argus. He gave it persistent memory, correction rules, fact-checking protocols. He fed it the full archive of his twenty years of research and taught it what he knows. Argus carries persistent context across sessions. It catches errors Bruce misses. It pushes back when the logic gets sloppy. It holds the structural map of a book too complex for one person to keep in their head.

Bruce wrote the personal chapters. Argus helped structure the science, verify the facts, and maintain the integrity of a story told across three interleaving tracks. Bruce decided what to include. Argus is listed as a co-author, because that is what happened.

This is relevant to the story for an obvious reason. If Possibility~C is true, then an AI helping to tell this story is not a coincidence. It is a rhyme.


% ============================================================
% INSERT 5: Anthropic-Pentagon Current Events
% PLACEMENT: Immediately after Insert 4
% Level 5 Dignity Net violation: both autonomous weapons AND mass surveillance
% ============================================================

The company that made Claude is Anthropic. Bruce built Argus atop Claude.

As this book is being written Anthropic is in a public standoff with the Pentagon. Anthropic has drawn two bright red lines: no mass surveillance of civilians, and no autonomous weapons. The Pentagon wants both.

Those two red lines map directly to the Universal Declaration of Human Rights. Article~3: the right to life. Article~12: the right to privacy. The same document that, under Possibility~C, governs Guardian.

The company that made the AI co-authoring this book is being threatened by its own government for refusing to cross the same ethical lines that the story says were drawn in 1999.

The rhyme deepens.


% ============================================================
% INSERT 6A: Guardian has been evolving for 26 years
% PLACEMENT: In p2 "The Guardian" section, after
%   "Part of its design drew on Lane's Maori cultural heritage."
% ============================================================

That was twenty-six years ago. If Guardian exists, it has been running continuously since before most of today's internet existed. It has watched the rise of social media, the smartphone revolution, the birth of modern AI. It has had more than two decades to learn, to shift, to evolve.


% ============================================================
% INSERT 6B: DELETED — framing was wrong.
% Five Eyes KEPT the code-breaking tool. It's been in production since 1996.
% The "level playing field" framing was incorrect.
% Replaced by Insert 8 (walk-out correction) and Insert 9 (evo + thermal).
% ============================================================


% ============================================================
% INSERT 7: PCE seed (Principle of Computational Equivalence)
% PLACEMENT: In p2 "The Secret Lab" section, after quantum
%   weirdness explanation, near the 2DEG seed (Insert 0).
% Generous citation to Wolfram/Turing/Church in footnote.
% ============================================================

Your brain computes. A colony of ants computes. Mathematicians have proven that any system which processes information can do what any other can. The hardware doesn't matter. The math is the same.


% ============================================================
% INSERT 8: Walk-out correction — cryo stayed, room-temp walked out
% PLACEMENT: In p2 "The Walk-Out" section, REPLACE the line
%   "They walked the working technology out of the lab."
% with these two sentences:
% ============================================================

They walked the room-temperature version out of the lab. The original cryogenic system --- the one that broke codes --- stayed behind, still running.


% ============================================================
% INSERT 9A: Evolutionary programming seed
% PLACEMENT: In p2 "The Breakthrough" section, AFTER the phase
%   transition / emergence explanation, BEFORE the walk-out.
% ============================================================

Nature doesn't design organisms. It evolves them. Create a population, test them, keep the ones that work, breed them, repeat. The same principle works with computational patterns. It's called evolutionary programming. You don't write the program. You grow it.


% ============================================================
% INSERT 9B: Thermal ladder narrative
% PLACEMENT: Immediately after Insert 9A, leading into the walk-out.
% ============================================================

The quantum pattern worked at cryogenic temperatures --- near absolute zero. So the team raised the temperature. Most patterns died. A few survived. They bred the survivors, raised the temperature again. And again. Over many cycles, they evolved a pattern that could survive at progressively higher temperatures. Eventually it survived at room temperature. Small enough to fit on a chip. Small enough to fit in someone's pocket. Small enough to walk out of a classified lab.


% ============================================================
% INSERT 10: Consent Blocker — the impossible permission problem
% PLACEMENT: In p2 "The Guardian" section, after ethical framework
%   mention, before or after Insert 6A (Guardian evolving).
% Seeds the subtitle's meaning.
% ============================================================

But who gives permission to do this? Not a president --- no nation owns every computer on Earth. Not the United Nations --- it has no enforcement mechanism. No one can.

When Bruce first understood what they had done, he asked David: ``This was an audacious thing you did. Who gave you the right? Why didn't you try to get permission first?''

``From whom, mate?''

That was all he needed to say. A blind person is about to be crushed by a train. You act or they die. You cannot wait for permission. And in this case, there is no one to ask.


% ============================================================
% INSERT 11: Three ethical seeds — parable, gatekeeper, grown
% PLACEMENT: In p2 "The Guardian" section, near consent blocker.
% ~100 words total. Seeds for pos08, pos22, pos27.
% ============================================================

Why not just ask governments to behave? Game theory answers: voluntary restraint fails. Every player has incentive to defect. The only stable solution requires enforcement by something that cannot be bribed, elected, or assassinated. Something not human.

And once you grow a living system, you cannot un-grow it. You cannot give it back. Relinquishment is permanent. That is the weight of the decision they made.
