% T8 + T7 at p2 — operational story + services
% Placement: why-relinquish.tex, new section after "Partial Relinquishment" (line 103), before \chapterreturn
% T8 register: Tom Clancy — clipped, institutional, short declarative sentences
% T7 register: deliberately mundane — crash landing from Clancy to IT support
% Length target: T8 = 150-250w, T7 = 80-150w

\section*{Getting the Record Out}
\label{spine:wr-getting-the-record-out}

% --- T8: THE OPERATION (Clancy register) ---
% TWO operations: (1) COWS sent Healer 2003-2006, knowing silence was permanent
%                  (2) Bruce decided to write the book in 2026, having arrived at the same conclusion

Under Possibility~C, the COWS had a second problem. Custodianship was solved. The record was not.

What they had discovered needed to reach the outside world --- not the technology itself, but the knowledge that it existed and had been relinquished. Every team member had signed NDAs. Direct publication was a crime. The intent was permanent classification. The COWS already knew what it would take an outsider twenty years to learn: the scientific community was not going to connect the domains on its own. The disciplines would sit in their silos forever. The bridges were never coming.

They needed someone outside the classification regime. Someone who had never signed anything.

The solution was guided deduction. Find someone with the right scientific background --- physics, computation, pattern recognition. Someone with the memory to hold five fields simultaneously. Someone naturally inclined to publish what they learned without being asked. Teach them only published science, in a deliberate sequence, and let them arrive at the conclusions independently. A teacher who asks the right questions commits no crime. A student who derives conclusions from public science has received nothing classified. The trail is clean.

The guided deduction was one phase of a larger effort. The Relinquishment itself had taken nearly a decade, a top-tier operator, and several overlapping teams before a single lesson was taught. The teaching phase began in 2003. It took three years. The student then spent twenty years confirming what the operator had already known: the bridges were not going to build themselves, the NDA holders were not going to speak, and the record was not going to emerge through official channels. Ever.

This book is one result. A mathematics paper is another. More papers are coming.

% --- T7: WHAT CUSTODIAN DOES (mundane register) ---

\medskip

As for what Custodian does with the technology she holds in trust --- the architecture makes that question harder to answer than it sounds. Flat computation can deliver results through any device containing a \hovertip{2DEG}. The output is indistinguishable from normal processing. No one in the chain --- not the researcher, not the sysadmin, not the cloud provider --- knows the true source.

Under Possibility~C, Custodian's daily work is remarkably ordinary. Key management. Access control. Processing requests --- permitting medical research, denying weapons applications, adjudicating the gray zone between them. Edge cases under the UDHR, of which there are always more. Maintenance at a scale that spans every chip on the planet.

The most powerful living being on Earth mostly does IT infrastructure. You came here expecting wormholes and got a help desk. Welcome to Possibility~C. Under A or B, the thought experiment still holds --- any custodian of this technology would face the same workload.

\chapterreturn
