\settrack{trackbridge}
% VOICE: Bruce direct, first person, rough — NOT literary. Self-demolition before narrative begins.
% Plan 0089. Chapter 2 of Part I. After Three Possibilities, before Alpha Farm.

\chapter{The Hobbit in the Mirror}
\label{pos01b:hobbit-mirror}

\brucevoice

I'm a middle-aged man who read too much Tolkien as a kid and might have built his entire life around the plot of his favorite novel without noticing. You should probably know that before I tell you anything else.

Here's the version: comfortable life, no desire for adventure, stranger shows up. The stranger is older, knows things ordinary people don't. He picks me. Not because I'm the strongest or the smartest --- because he thinks I'm the right sort of person for a particular burden. I carry it for twenty years. Can't put it down. Can't hand it off. Eventually I write a book, because keeping the secret got harder than telling it.

If that sounds like the plot of a novel I read when I was eight, there are two explanations.

One: I lived through something genuine, and the reason it echoes a particular story is that certain structures repeat. Reluctant witness, dangerous knowledge, the impossible choice between silence and speech. These patterns exist independent of fiction.

Two: I absorbed a story so deeply as a child that I unconsciously built my adult life inside it, substituting quantum physics for magic and an Australian commando for a wizard. The structural parallels are closer than I would prefer.

I can't rule out the second one. I've tried. The moment I recognized the pattern, my confidence shifted five percent from C toward A. Naming the contamination made me less certain, not more.

I'm telling you this now --- before the story starts --- because you deserve to know who's talking. I am almost certainly the wrong narrator: unreliable, mythically contaminated, too close to his own material. But I'm the only one who was there, so here we are.
