\settrack{trackbridge}
% VOICE: expository/pedagogical — teaches concepts, no narrative events, no "the COWS did X"

\chapter{Dual-Use: A Brief History of Dangerous Ideas}
\label{pos08:dual-use}
\aidraftchapter

% Source: gpt-history.md + unintended-consequences.md (imported Plan 0011)

A general purpose technology is a technology that generates other technologies. Stone tools are an early example. Artificial Intelligence is a recent example.

One new technology can sometimes create whole new classes of technology. Control of fire gave our species many extra hours each day and thereby made possible boating, agriculture, metallurgy, simple machines, and writing. After thousands of years these, in turn, led to mathematics, engineering, chemistry, steam power, electricity and magnetism, atomic power, information technology, and genetic engineering. Early GPTs came thousands, then hundreds of years apart. More recently, new general technologies have been arriving every few decades.

\section*{General Technologies in Prehistory}

For many hundreds of thousands of years Humans lived in hunter-gatherer bands with good Control of Fire. Humans have had ample time to evolve since discovering fire. Human control of fire provided humans with multiple benefits. One large benefit from controlling fire is probably cooked food. Cooked food drastically reduces time spent chewing, freeing up many hours each day for other purposes. Control of fire also has many other benefits, such as improved security, light, heat, improved tool building, et cetera. Cooked food gave us time to pursue activities besides chewing food. Humans have been eating cooked food for so long that our bodies have evolved to expect it.

Boating probably qualifies as a general technology, and has certainly served humanity well. The earliest circumstantial evidence of boating is 130,000 years ago. The oldest surviving boat is about 10,000 years old. Boating enabled increased travel and trade. Boating qualifies as a Dual Use Technology, with both peaceful and warlike applications.

Agriculture came next. Agriculture probably started with forest gardening in prehistoric times. Pastoralism, the raising of livestock, probably followed. Agriculture greatly increases the human carrying capacity of a region. Farmers spend more time finding food than do hunter-gatherers, but many more farmers can inhabit a given region. Irrigation technology drastically improved the effectiveness of agriculture. Permanent villages, towns, and cities became possible.

Metallurgy literally defines the transitions between the stone age, the bronze age, and the iron age. The earliest evidence of humans smelting metal is approximately 7,000 years old. Tools of copper and bronze were routine for several thousand years. The Iron Age began about 3,000 years ago. Metalworking was seemingly never discovered in Australia. In the Americas, sophisticated metallurgy developed independently in South America but did not reach most of North America before European contact. Metalworking improved the efficiency of farming and introduced the potential for large scale military conquest. Metalworking is thus a dual-use technology, suitable for both peaceful and warlike activities.

The wheel and other simple machines were developed independently in several parts of the world. The first evidence for wheeled vehicles is in 5,300+ year old pottery. The water wheel came much later, with the earliest evidence dating to roughly 2,300 years ago. Chariots and wagons have been in common use for about 4,000 years. Wheeled vehicles need roads. Wheeled vehicles allowed for trading more stuff over longer distances, improved farming efficiency, war chariots, et cetera. Wheeled vehicles were not developed in the Americas, nor in Australia, probably because of the absence of large domesticable animals like cows and horses that could pull heavy loads. Simple machines have both peaceful uses, such as grinding grain with a water wheel, and military uses, such as war chariots.

\vspace{0.5cm}
\begin{center}
\textbf{Timeline of General Purpose Technologies in Human History}
\end{center}
\vspace{0.3cm}

\noindent\begin{tabular}{p{3.2cm}p{4.5cm}p{6.3cm}}
\textbf{Technology} & \textbf{When} & \textbf{Who} \\
\hline
Stone toolmaking & about 3 million years ago & Unknown \\
Control of Fire & about 500,000 years ago & Unknown \\
Boating & certainly 10,000 years ago, possibly 130,000 & Unknown \\
Agriculture & about 12,000 years ago & Unknown \\
Wheel & about 7,000 years ago & Unknown \\
Metallurgy & about 5,000 years ago & Unknown \\
Written language & about 5,000 years ago & Mesopotamians and Mesoamericans \\
Fossil energy & about 600 years ago & Unknown \\
Chemistry & about 500 years ago & Many scientists \\
Combustion Engine & about 300 years ago & Thomas Newcomen \\
Electricity & about 200 years ago & Franklin, Faraday, Maxwell \\
Germ Theory & about 140 years ago & Pasteur, Lister \\
Radio & $\sim$1890 & Marconi, Tesla, and others \\
Atomic power & $\sim$1898 & Curie, Einstein, Fermi, Szilard \\
Computer & $\sim$1941 & Alan Turing \& the Ultra project \\
Robotics & 1920 & Many scientists \\
Internet & 1969 & DARPA \\
Genetic Engineering & 1973 & Many scientists \\
Nanotechnology & 1959 & Many scientists, starting with Feynman \\
\end{tabular}

\section*{General Technologies in Early History}

Written language began about 5,000 years ago. Writing was useful for recording increasingly complicated transactions that came with more cities and more trade. Early writing and counting systems often involved taxation. Writing can reliably transfer information between individuals who are far apart in space and time. Sustained scientific inquiry became possible with the advent of writing, but mathematics was still hindered by poor writing methods. Geometry was discovered and written down, pretty much as we learn it today. Engineering became a profession. Writing was discovered independently in Mesopotamia and, later, in Mesoamerica.

Humanity now passes several thousand years. About a dozen solar-powered complex civilizations emerge, grow, and collapse in Africa, Asia, Europe, and the Americas. Existing technologies are refined during this time. Agricultural humans become dominant, although some hunter-gatherer cultures hold on in marginal land. The destruction of the Library of Alexandria --- through multiple events over several centuries --- was a great loss of accumulated knowledge.

About 1200 years ago a Persian mathematician and scholar named al-Khwarizmi transmitted to the Islamic world what we now call the Hindu-Arabic number system, which Indian mathematicians had invented centuries earlier. His treatise helped spread the Indian concept of zero as a useful number throughout the Middle East and eventually to Europe. The Islamic Golden Age --- centered in Baghdad under the Abbasid Caliphate, with many Persian scholars --- developed and spread techniques for algebra, geometry, advanced medicine including surgery, early chemistry, and other arts and sciences. While some of these technologies certainly have military applications they were not generally thought of that way.

The practice of burning small amounts of fossil fuel deposits began thousands of years ago. Coal use became common 600 years ago in parts of Europe where the forests had been cut, as a replacement for wood. People observed that coal released much more heat than wood when it was burned. Burning of fossil fuel reserves began exponential human population growth to today's 7.4+ billion. Oil contains even more concentrated energy than does coal. There is no known replacement for the quantity and quality of how human industrial civilization currently uses oil. Human industrial civilization is almost entirely fossil fuel powered.

\section*{General Technologies in the Industrial Age}

The pace of advancement in science and technology greatly increases during the industrial age. The Industrial Age has seen a new general purpose technology invented every 50 years or so: combustion engines, electricity, radio, computer, atomic power, the Internet, genetic engineering, nanotechnology, and robotics. The steam engine heralded the Industrial Age. The combustion engine eventually led to planes, trains, and automobiles. Good metallurgy, the practice of burning coal, and the problem of water infiltration of underground mines resulted in the invention of the steam engine. In 1712 Thomas Newcomen built the first practical combustion engine. Humans had learned to power machines by burning Earth's accumulated bounty of ancient sunlight. Advanced metallurgy was required, as was a steady supply of coal. The first steam engines, while inefficient, successfully pumped water out of mines to allow mining below the water table. One need only consider steam-powered warships to see that the combustion engine is clearly a dual use technology.

Electrical technology is an especially important general purpose technology. Electrical technology has spawned many other new technologies. Some aspects of electricity, such as static charge and magnets, were known in ancient times. The first careful scientific investigation of electricity began in 1600. In 1752 Benjamin Franklin discovered that lightning is an electric phenomenon. In 1831 Michael Faraday began his great experiments with electricity, leading to his experimental discovery of electromagnetic induction. In 1865 James Maxwell published a mathematical explanation for what Faraday had discovered experimentally. In 1879 a young scientist from Johns Hopkins University named Edwin Hall discovered the Hall Effect. The Hall Effect provided early evidence that electric current in metals is carried by negatively charged particles --- evidence that would converge with other experiments leading to J.J.\ Thomson's identification of the electron in 1897. Some specific technologies derived from control of electricity include the light bulb, telegraph, telephone, radio, electric motor, electronics, and many others.

The 1890s saw electricity and magnetism precisely controlled over distance in the form of radio. Guglielmo Marconi, building on work by Nikola Tesla and others, is generally credited with practical radio. Some important technologies derived from radio are radio communication, radio astronomy, x-rays, and modern cellular and wifi communication. Radio makes it possible for a human in, say, China to instantly and directly communicate with a human in, say, Mexico. Radio technology is obviously Dual Use, with clear civilian and military applications. Atomic power was the first new general technology to manifest during the 20th century. It began with Physicists exploring the subatomic world of electrons, protons, and neutrons. The discovery of Radioactivity was announced by Marie Curie and others in 1898. In 1905 Albert Einstein published his most famous equation, which gave the first hint of possible military applications for Atomic Power. Atomic Power is the classic Dual Use Technology.

Massive government investment was essential to the development of atomic power and all other 20th century General Technologies. Gone were the days when one brilliant human could develop a new GPT\@.

\section*{General Technologies in the Information Age}

The later part of the 20th century has been called the Information Age. Computer \& Information technology become very influential.

The Information Age arguably began when Lady Ada Lovelace wrote what is considered the first computer program, in 1843, to calculate Bernoulli numbers on Charles Babbage's theoretical Analytical Engine. Computer hardware took 100 years to catch up. Scientific investigation of biological neural networks began circa 1880 and reached its fruition more than 100 years later. Information technology was arguably invented in 1936 by British mathematician Alan Turing, now considered the father of both Computer Science and Artificial Intelligence.

Computer Science became a thing during World War Two via the Ultra Secret. Early in WW2 British Prime Minister Winston Churchill decided to spend precious wartime resources on code breaking. British code breakers succeeded in reliably breaking critical wartime Axis military codes. They did this by building binary electric computers programmed to break codes. Wartime information gained from reading the opposition's most secret messages was designated Ultra. Historians generally credit the Ultra Secret with shortening WWII by several years. Despite many thousands of participants, Ultra was kept secret until 1974, about 35 years after it was begun. The Ultra Secret is probably the best historical example in which thousands of people conspired to keep an important secret for several decades. After WWII computer science became the domain of private corporations, especially International Business Machines (IBM). Computer science is clearly a dual use technology, although military applications are perhaps less obvious than peaceful applications.

The invention of the Transistor heralded the Information Age and is the origin of Moore's Law. The Transistor is probably the most influential invention of the 20th century. The transistor was invented in Bell Labs shortly after WW2. A transistor can amplify an electric signal. Before transistors, for a few decades, vacuum tubes did the same job, but they were big, clunky, and prone to failure. Nearly all electronic devices contain transistors. Modern transistors are called Metal Oxide Semiconductor Field Effect Transistors, or MOSFETs. Every MOSFET contains a Two Dimensional Electron Gas (2DEG) at its silicon-oxide interface --- hundreds of billions of them, in every electronic device on the planet. It goes without saying that Transistor technology is a dual use technology.

The Internet is a recent GPT\@. A team of scientists recruited by ARPA (later renamed DARPA) invented the internet in 1969. DARPA stands for Defense Advanced Research Projects Agency. The Internet was originally intended for military communication in the event of apocalyptic nuclear war. Only later was the Internet deemed useful for civilian purposes.

DARPA is a US government agency specialized in scientific research for military purposes. DARPA focuses on short-term (two to four-year) projects run by small, cross-disciplinary, purpose-built teams. DARPA informally bills itself as ``100 geniuses connected by a travel agency.'' As DARPA describes itself:

\begin{quote}
\textit{DARPA's original mission, established in 1958, was to prevent technological surprise like the launch of Sputnik, which signaled that the Soviets had beaten the U.S.\ into space. The mission statement has evolved over time. Today, DARPA's mission is still to prevent technological surprise to the US, but also to create technological surprise for our enemies.}
\end{quote}

Since its founding DARPA has conducted many projects. Most DARPA projects remain secret and we never hear of them. Some few are declassified and become public, like the Internet. DARPA is a hub for the development of new technologies and is likely associated with any major new technology breakthroughs.

Artificial Intelligence (AI) is a relatively new GPT, closely related to robotics. Alan Turing is known as ``the father of artificial intelligence.'' AI is sometimes divided into ``Narrow AI'' and ``General AI\@.'' AI can now outperform humans in a wide variety of fields. The origins of the AI revolution are shrouded in mystery and complexity, although it's a safe bet that DARPA is somehow involved.

Genetic engineering is a young general purpose technology. Exponential self-replication is a key distinguishing characteristic of Genetic Engineering. Numerous genetically engineered organisms have been released into the biosphere. We are still learning the benefits and potential pitfalls of Genetic Engineering.

The concept of Nanotechnology was introduced in 1959 by Richard Feynman. Some aspects of nanotechnology are just extensions of existing technologies. Other aspects of nanotechnology, particularly the ability to self reproduce, have more in common with Genetic Engineering. Nanotechnology is still in its infancy. It's possible that more advanced forms of nanotechnology have been invented but are not yet public knowledge.

\section*{Unintended Consequences}

Many of these technologies are quite new and we are still learning what will be their consequences, intended or otherwise. It seems likely that any new GPT will pose risks in proportion to its potential. Here are just a few examples of unintended consequences in science and technology.

\subsection*{Eli Whitney and the Cotton Gin}

Eli Whitney (1765--1825) was an American inventor known for inventing the cotton gin and also for successfully promoting the concept of manufacturing standardized identical parts. Before the cotton gin one worker could produce about one pound of seed-free cotton in one day. With the cotton gin that same worker could produce about 50 pounds of seed-free cotton. He had observed slaves toiling in the cotton fields.

Instead, the cotton gin revolutionized the way cotton was harvested and reinvigorated the institution of slavery. The cotton gin changed the economics of slavery, making slavery much more profitable. This was an unintended consequence of Eli Whitney's invention.

\subsection*{Alfred Nobel, Military Explosives, and the Nobel Prize}

\begin{quote}
``I intend to leave after my death a large fund for the promotion of the peace idea, but I am skeptical as to its results.'' --- Alfred Nobel
\end{quote}

Alfred Nobel was a Swedish chemist, engineer, innovator, and armaments manufacturer. He was the inventor of dynamite and several other types of explosive. His family businesses gathered great wealth making arms and military explosives.

In 1888 Alfred's brother Ludvig died while visiting Cannes and a French newspaper erroneously published Alfred's obituary. It condemned him for his invention of dynamite and is said to have brought about his decision to leave a better legacy after his death. The obituary stated ``Le marchand de la mort est mort'' (``The merchant of death is dead'') and went on to say, ``Dr.\ Alfred Nobel, who became rich by finding ways to kill more people faster than ever before, died yesterday.'' Alfred was disappointed with what he read and concerned with how he would be remembered.

On 27 November 1895, at the Swedish-Norwegian Club in Paris, Nobel signed his last will and testament and set aside the bulk of his estate to establish the Nobel Prizes, to be awarded annually without distinction of nationality. After taxes and bequests to individuals, Nobel's will allocated 94\% of his total assets, 31,225,000 Swedish kronor, to establish the five Nobel Prizes.

\subsection*{Albert Einstein and $E = mc^2$}

In 1905 Albert Einstein, known as the ``father of modern physics,'' proposed Mass-Energy equivalence, which led to the most famous equation of the 20th century: $E = mc^2$. In the 1930s, the discovery of nuclear fission revealed that Einstein's equation explained the enormous energy released when atoms split. This led to the discovery of atomic weapons, the atomic bombing of Hiroshima and Nagasaki, the nuclear arsenals of the Cold War, the concept of Mutually Assured Destruction, and the ongoing threat of global thermonuclear war.

Einstein was a committed pacifist for most of his life, though he made a pragmatic exception when confronted with the existential threat of Nazi Germany. He had this to say about war and soldiering:

\begin{quote}
``He who joyfully marches to music rank and file, has already earned my contempt. He has been given a large brain by mistake, since for him the spinal cord would surely suffice. This disgrace to civilization should be done away with at once. Heroism at command, how violently I hate all this, how despicable and ignoble war is; I would rather be torn to shreds than be a part of so base an action. It is my conviction that killing under the cloak of war is nothing but an act of murder.'' --- Albert Einstein
\end{quote}

Einstein was deeply opposed to anything to do with nuclear weapons. Although his 1939 letter to President Roosevelt helped initiate the Manhattan Project, Einstein himself was denied a security clearance to participate in the program, because his pacifist views and associations made him a security risk in the eyes of the military. Einstein had this to say about nuclear war:

\begin{quote}
``I know not with what weapons World War III will be fought, but World War IV will be fought with sticks and stones.'' --- Albert Einstein
\end{quote}

Einstein was horrified at how his most famous scientific discovery, $E = mc^2$, led to the creation of atomic weapons:

\begin{quote}
``The release of atom power has changed everything except our way of thinking\ldots{} the solution to this problem lies in the heart of mankind.'' --- Albert Einstein, 1946
\end{quote}

\noindent A year later, he was more blunt:

\begin{quote}
``Had I known that the Germans would not succeed in producing an atomic bomb, I never would have lifted a finger.'' --- Albert Einstein, \textit{Newsweek}, March 10, 1947
\end{quote}

Einstein tells us that, had he known the consequences, he would not have done the work.

\section*{Sources}

\begin{enumerate}
\item Robert Wehrli, ``Essays in General Purpose Technologies,'' Doctoral dissertation, ETH Zurich, 2009.
\item ``50 Years of Bridging the Gap,'' DARPA\@.
\item ``A Brief History of Cooking with Fire,'' National Geographic.
\item ``Why Fire Makes Us Human,'' Smithsonian Magazine.
\end{enumerate}

\chapterreturn
