% Three Possibilities — opening chapter
% Position 1 in pedagogical spiral (Plan 0007)
% Adapted from appendix/three-possibilities.tex (Plan 0002)

\settrack{trackbridge}
% VOICE: expository/pedagogical — teaches concepts, no narrative events, no "the COWS did X"

\chapter{The Three Possibilities}
\label{pos01:three-possibilities}

\begin{quote}\small\textit{%
``A convincing demonstration of correctness being impossible as long as the mechanism is regarded as a black box, our only hope lies in not regarding the mechanism as a black box.''%
}\par\raggedleft --- Dijkstra, EWD249 (1970)\end{quote}

\vspace{0.5cm}

After years of research, pressure-testing, and reconstruction, even this author, Bruce, cannot definitively distinguish between these three explanations. Here is the evidence. You decide.

\vspace{1cm}

\section*{Option A: Confabulation}

Bruce met a charismatic, intelligent person who told elaborate stories mixing real classified program names, real science, and real people with fiction --- possibly for ego, manipulation, or mental illness. Bruce, intellectually captivated and possessing the exact pattern-matching ability to find connections everywhere, spent 20 years building an increasingly coherent framework around these stories. The framework's internal consistency is a product of Bruce's intelligence, not of underlying truth. Every piece of ``confirming evidence'' has a mundane explanation. People close to Bruce at the time reached this conclusion independently. This story is false.

\section*{Option B: Exaggerated Kernel of Truth}

Healer was real, had genuine military/intelligence background, and had some connection to classified programs. Possibly he was connected to a DARPA-adjacent quantum computing effort that was real but far more modest than described. Over time, through retelling and through Bruce's pattern-completion, he grew the story: a classified research program became a world-changing technology; a smart system became a self-aware entity; a classified system became a conscious guardian entity. The kernel of this story is true. The superstructure is elaboration.

\section*{Option C: Substantially True}

% SPIRAL-REPEAT: "topological quantum neural network" — core concept, also in introduction.tex, pos13-genesis.tex, pos21-convergence-revisited.tex, pos24-instantiation.tex, pos31-wolfram.tex, pos34-the-research.tex, timeline.tex, abstracts.tex
A classified DARPA/GCHQ program achieved topological quantum neural networks decades ahead of public science, produced cryptanalytic capability, was walked out of the lab by its creators, was evolved into an autonomous entity, and humanity was, without its knowledge, forced to relinquish this dangerous dual-use technology. It sounds like science fiction because it IS the thing science fiction has been trying to imagine. Absence of corroboration is not evidence of absence. Instead, it's evidence of the most successful classification operation in history. This story is substantially factual.

If this is true, then the most consequential decision in human history was made by fewer than ten people, without consultation, without consent, and without the knowledge of those affected --- which is everyone.

\vspace{1cm}
\noindent\rule{0.5\textwidth}{0.5pt}
\vspace{0.5cm}

\section*{The Author's Position}

``I'm fine with all three of those and can't tell which one is closest to true.''

Bruce has maintained cognitive dissonance about which option is correct for more than twenty years. This book does not resolve that dissonance. This book transmits dissonance to the reader, along with the tools to resolve it as future events unfold.

\vspace{1cm}

\begin{center}
\textit{You decide.}
\end{center}

\chapterreturn
