\settrack{trackbridge}

\chapter{The Experiment}
\label{pos11:the-experiment}

% DMS MVP import from staging/raw/ — not final prose

\srcnote{CH3 - Practical Cryptography \& project ULTRA II.txt | staging/raw/pos11-the-experiment.md | 2026-02-15 | DARPA project inception, five scientists, COWS formation, walk-out}

\section*{Project Inception}

In 1991 DARPA gathered a team of scientists, each a leading specialist in a particular field, to develop a new code breaking device.  The intent was to combine several very different scientific fields in pursuit of a particular new technology.  Specifically, the project would exploit the newly discovered phenomenon of experimental quantum teleportation to produce a useful cryptography tool. The ULTRA~II project was completely successful, and produced the sought after ``Magic Decoder Box'' in 1995.  In fact, it was the most successful DARPA project since DARPA invented the Internet. Project ULTRA~II generated another new General Purpose Technology.  The best name for this new technology is ``Quantum Neural Network'', or QNN, technology.

\section*{The Team}

The team leader had previous experience leading large successful cryptographic science projects, and has a personal body of work that is both deeply counterintuitive and demonstrably true.  One team member was at the forefront of the new scientific discipline now called Complex System Biology.  Another team member, a Solid State Physicist, contributed understanding of the FQHE\@.  One mathematician specialized in the sort of quantum algorithms pioneered by Deutsch and Josza.  A young wildcard with exceptional raw mathematical and mechanical aptitude rounded out the team.  Several team members had practical working knowledge of neural networks.

\section*{The Unintended Consequence}

At first the ULTRA~II team was exhilarated. They were working on a very exciting project, and things were going very well. The basic approach, generating an emergent quantum neural net in the FQHE environment, had worked. At first it did nothing but exist. The project scientists gradually guided its basic form and function to evolve in specific ways, and trained it to do basic tasks. Stuart Kauffman describes one such technique, although it may have been combined with other methods of evolutionary programming. Once they had adequate control over it, and had caused it to evolve the required capabilities, they trained it to perform the needed computations. Only then would it be possible to begin to implement quantum algorithms.

\section*{The COWS Formation}

In 1994 some of the ULTRA~II project scientists began to worry about the implications of QNN technology. They had been at it for several years but were far from finished.  They realized they had invented a major new technology that seemed to qualify as a new general technology.  They had already used it to create one new technology, a working quantum computer, and they had ideas for several others.  They knew that any new general technology can profoundly influence the human condition, for good or ill. They did not want to make the same errors as the atomic scientists of the Manhattan project, and permit the proliferation of a new technology that might pose unknown risks to humanity. They realized that QNN technology could be used for terrible purposes. They did not wish to see those purposes attempted, and they did not wish to see a new arms race.

They decided that they must take drastic action to prevent the newly discovered QNN technology from being turned into a terrible weapon or, possibly worse, an Orwellian spy machine.

\section*{Replacing Gell-Mann}

\srcnote{HEALER-RECONSTRUCTION.md | staging/raw/pos11-the-experiment.md | 2026-02-15 | Gell-Mann medical leave, Healer selection, Angerman on-the-record source}

Murray Gell-Mann --- Nobel Laureate (quark theory, 1969), co-founder of SFI, age sixty in 1990 --- took a leave of absence ``for medical reasons'' from the Santa Fe Institute that year. Michael Angerman, who was at SFI from 1989 to 1995, confirmed the leave and was surprised Bruce knew about it. Angerman stated it was not really a medical issue but ``something else'' --- personal circumstances that rendered him unable to continue on a demanding project. The ``medical reasons'' were a cover story.

You cannot replace a Gell-Mann one-for-one. But by 1990, the theoretical framework was presumably already in place. What the project needed was someone to operationalize the theory. The replacement criteria were not ``find another theorist'' but: TS/SCI clearance minimum, young enough for a long-duration classified project, intellectually capable of absorbing the theory with mentorship, military/intelligence background, and an operational mindset to bridge theory and implementation.

The proposition: Healer was selected around 1990--1991. Australian SAS, age approximately twenty-five, identified through the Five Eyes special forces pipeline. Not a peer theorist --- an operational lead who would be brought up to speed. Wolfram mentored Healer from approximately 1991 to 1993, with Wolfram's computation-as-physics framework serving as the bridge.

Angerman was working for Kauffman on Neural Networks at SFI during the entire period. He went directly from SFI to Sun Microsystems (1996--1997) --- Bill Joy's company. His willingness to speak on the record, combined with no NDA, makes him the most accessible living witness.

\chapterreturn
