\settrack{trackbridge}

\chapter{Why Give It Up}
\label{pos22:why-give-it-up}

% Source: manuscript/sources/ch3-relinquishment.md (imported Plan 0011)

According to this account, the General Purpose Technology at the heart of this story was developed by DARPA in the early 1990s. The project's actual name is still buried in layers of official secrecy. I shall refer to this project as ULTRA II, in honor of the Ultra project of WW2. ULTRA II is properly descriptive of what it was about and what it accomplished.

In $\sim$1992 DARPA gathered a team of scientists, each a leading specialist in a particular field, to develop a new code breaking device. The intent was to combine several very different scientific fields in pursuit of a particular new technology. Specifically, the project would exploit topological quantum computation via anyonic braiding to produce a useful cryptography tool. Tip your metaphorical hat at the 1992 Robert Redford movie, Sneakers. The ULTRA II project was completely successful, and produced the sought after `Magic Decoder Box' in 1995. It was the most successful DARPA project since DARPA invented the Internet. Project ULTRA II generated another new General Purpose Technology: Topological Quantum Neural Network, or TQNN, technology --- a particular form of topological quantum computation-based nanotechnology.

What follows is my best reconstruction of how project ULTRA II unfolded. Much of the evidence is completely open source, and any interested and determined party can learn of it.

\section*{Why Would DARPA Conduct Project ULTRA II?}

\section*{The State of Cryptography in 1990}

In 1990 the world's top spy organizations had an unhappy blind spot. The mathematics of Public Key Cryptography (publicly described 1976--1978, though GCHQ had invented it secretly in the early 1970s) allowed civilian computer users to routinely encrypt communications in unbreakable ways. Officially, they still have this same blind spot in 2013. Public key cryptography provides the S in HTTPS\@. Alice and Bob could communicate privately. Eve could analyse traffic, but she could not eavesdrop. Organizations that existed to spy on communications, such as the United States National Security Agency and the other members of the Five Eyes, felt threatened. The spy organizations badly needed a reliable way to crack public key cryptography. Something had to be done.

One need not be a mathematician to understand how the most widely known public-key system, RSA, works. It is based on the principle that it is harder to factor a large number than to multiply its factors. Quick, what are the prime factors of 91? Stop reading while you calculate the answer. After either answering that problem where you factor a number into its prime factors, or giving up, multiply thirteen by seven. Which was harder? Why? The security of Public Key Cryptography comes from the mathematical difficulty of finding prime factors for very large numbers. If the NSA spies wanted their cryptography problem solved, then this mathematical problem had to be solved. This has been common knowledge to cryptographers since the late 1970s.

In terms of US Government agencies in 1990, the obvious agency to try to solve the NSA's cryptography problem was DARPA\@. DARPA is the Defense Advanced Research Projects Agency. DARPA develops new technology for the military. DARPA focuses on short-term (two to four year) projects run by small, purpose-built teams. In 2013 DARPA has a direct budget of about US\$3 billion. Here are some interesting facts about DARPA, from Wikipedia.

\begin{itemize}
\item DARPA invented the Internet --- one of several new General Purpose Technologies developed in the 20th century.
\item Small and flexible: DARPA has only about 140 technical professionals; DARPA presents itself as ``100 geniuses connected by a travel agent.''
\item Flat organization: DARPA avoids hierarchy, essentially operating at only two management levels to ensure the free and rapid flow of information and ideas, and rapid decision-making.
\item Autonomy and freedom from bureaucratic impediments: DARPA has an exemption from Title V civilian personnel specifications, which provides for a direct authority to hire talents with the expediency not allowed by the standard civil service processes.
\item Eclectic, world-class technical staff and performers: DARPA seeks great talents and ideas from industry, universities, government laboratories, and individuals, mixing disciplines and theoretical and experimental strength.
\item Teams and networks: At its very best, DARPA creates and sustains great teams of researchers from different disciplines that collaborate and share in the teams' advances.
\item Project-based assignments organized around a challenge model: DARPA organizes a significant part of its portfolio around specific technology challenges. It foresees innovation-based capabilities and then works back to the fundamental breakthroughs required to make them possible.
\item The NSA and its Five Eyes counterparts (GCHQ, ASD, CSE, GCSB) are the largest employers of cryptographers in the English-speaking world.
\end{itemize}

In order to crack public key cryptography DARPA needed to find a practical mathematical method for rapidly factoring large numbers. This challenging topic has been the subject of many dissertations in mathematics. A mathematician would call the desired process factoring in polynomial time. While there are many tricks to speed the process, all of them are subject to exponential time slowdown as the size of the number increases. Basically, all known approaches amount to a variation of trying every possible permutation. Practical public/private key cryptography typically chooses a large enough number that it would take the best supercomputers many years of trial and error to find the prime factors and crack the key.

% GENERATOR TODO: Insert PKC technical section here. Depth must match the spiral pass —
% basics only for first encounter, fuller treatment (symmetric vs asymmetric, DH key exchange,
% RSA, digital signatures, TLS handshake) for technical pass. Section sits between RSA factoring
% explanation above and Deutsch/Shor quantum algorithms below. Must bridge: "factoring is hard"
% → "here's how real crypto uses that" → "now imagine breaking it."

In 1985 David Deutsch published the Deutsch algorithm, the first quantum algorithm demonstrably faster than any classical counterpart. In 1992 Deutsch and Richard Jozsa published the generalized Deutsch-Jozsa algorithm. These results hinted that a theoretical quantum algorithm might be able to factor numbers in polynomial time, and thus crack public key cryptography. A quantum algorithm (software) capable of doing just that was published by Peter Shor in 1994. The problem then, as now, is lack of appropriate quantum computation hardware on which to run such an algorithm.

In 1993 Bennett, Brassard, Cr\'{e}peau, Jozsa, Peres, and Wootters published the theoretical protocol for quantum teleportation. (The first experimental demonstration came later, by Bouwmeester et al.\ in 1997.) Quantum teleportation --- the transmission of quantum states via shared entanglement and classical communication --- is a fundamental primitive in certain models of quantum computation. According to this account, DARPA scientists were aware of related approaches before the formal publication in 1993.

By the early 1990s it was known to top cryptographers that Quantum algorithms could, in theory, crack public key cryptography. This was the basis for the 1992 Robert Redford movie Sneakers, although quantum computation was not specifically mentioned. No quantum hardware existed to run this quantum software. The project planners at DARPA saw a possible way to solve their cryptanalysis problem. They must invent a working quantum computation device. It must be powerful enough (e.g.\ have a large enough qubit array) to factor real cryptographic keys. Billions of dollars would potentially have been available for ULTRA II, if an approach seemed like it might work. Like project Ultra I, secrecy was of the utmost importance.

The overall project specifications for ULTRA II were now clear:

Create a tool capable of efficiently cracking public key cryptography. System must be able to rapidly determine the prime factors of a 4096 bit asymmetric key. System must be extensible to potentially handle much larger keys. System must have an input mechanism to accept the number to be factored. System must have an output mechanism to return the prime factors of the input.

\textbf{Basic Method:} Build some kind of quantum computer running some quantum algorithm.

\textbf{Problem:} No one knows how to build a device capable of meeting the specifications. Several approaches are known that can produce `toy sized' quantum computers, but none can scale to the thousands of qubits required to factor real cryptographic keys. The main problem is decoherence time. Most proposed systems of entangled qubits are very fragile.

\textbf{Definition:} A quantum computer exploits quantum mechanical phenomena --- superposition, entanglement, and interference --- to perform certain computations exponentially faster than classical computers. Quantum states can be generated in a variety of physical systems: photon polarization, electron spin, flux, charge, topological order, et cetera. Many techniques have been proposed to build practical quantum computers. By 2012 quite a few have been built, but only with small (tens of) qubit registers for gate-based systems.

A specific process must be followed to perform quantum computation. The qubit array must first be initialized to the proper superposition state. Then the qubit array is manipulated to perform the actual computation. Then certain measurements are made, and the results read. This all takes time. Adding more qubits to a quantum system increases the likelihood that at least one qubit will lose coherence before the computation completes. Without error correction, the probability of a successful computation drops exponentially as the number of qubits and gate operations increases. For any system large enough to factor real cryptographic keys, the error rate overwhelms the result. That will not work.

What is needed is a more robust form of entanglement. It must be resistant to environmental disturbance. It must have a quantum decoherence time much greater than the time required to perform computation and read the results.

\textbf{Question:} Do we know of any form of quantum entanglement that is less sensitive to disturbances, and that is either long lasting or persistent?

\textbf{Answer:} Yes! A particular exotic solid state system, called a Fractional Quantum Hall Effect (FQHE) liquid, manifests a type of quasiparticle called an anyon. An FQHE environment requires a two-dimensional system, such as is found in a Two Dimensional Electron Gas (2DEG) at temperatures near absolute zero in intense magnetic fields. Anyons are quasiparticle excitations that arise only in two-dimensional systems. In three dimensions, particles must obey either bosonic or fermionic statistics. In two dimensions, the topology of particle exchange permits a richer set of possibilities --- including non-abelian statistics, where exchanging particles performs a quantum gate operation. When anyons are braided --- moved around each other in specific patterns --- the quantum state of the system is transformed in a way that depends only on the topology of the braid, not on the precise path taken. This topological protection makes the computation inherently resistant to local noise and perturbation.

The behavior of braided anyons in an FQHE liquid is described by a branch of topology called `braid theory'. Information in a topological quantum system is encoded non-locally --- spread across the system's global state rather than stored in any single particle. This means local noise or perturbation cannot corrupt the computation, giving topological systems inherent fault tolerance far beyond what conventional quantum computers achieve.

At first glance actually building such a thing seems impossible. Any system that manipulates such tiny objects in an organized way would qualify as some form of advanced nanotechnology. No known scientific or technical techniques can do anything of the sort. This approach looked pretty much like a dead end.

Enter Stuart Kauffman's ideas about Complex System Biology (CSB), from an apparently unrelated scientific discipline. The key intellectual leap: the FQHE liquid environment, in which anyons exist, has the mathematical structure to support autocatalytic sets --- the same self-organizing dynamics Kauffman demonstrated in classical chemical networks, applied to a quantum substrate. This is the untested theoretical bridge that makes the whole story work. Kauffman proved that sufficiently complex networks of catalytic reactions spontaneously cross a threshold into self-sustaining order. The ULTRA II team recognized that the FQHE environment, with its rich interaction dynamics, might support the same kind of self-organization --- not with molecules, but with anyonic excitations.

Mathematical Biologist Stuart Kauffman demonstrated that a certain category of `poised' complex system has the emergent property that it is also a neural network. He proved this using mathematical models of autocatalytic chemical networks --- classical systems. The COWS' insight was that the same mathematical framework might apply to a quantum substrate: anyonic excitations in an FQHE liquid interacting with sufficient complexity to cross Kauffman's threshold. Same framework, untested domain. If it worked, the result would be not just a quantum computer but a quantum neural network --- something that computes, learns, and evolves.

In $\sim$1992 the ULTRA II team attempted to build a large scale working quantum computation system by applying the principles of Complex System Biology to the FQHE environment. As of 2013 no one else seems to have ever publicly discussed this approach. Each team member was an expert in some aspect of what the project would attempt.

The team leader had previous experience leading large successful cryptographic science projects, and has a personal body of work that is both deeply counterintuitive and demonstrably true. One team member was at the forefront of the new scientific discipline now called Complex System Biology. Another team member, a Solid State Physicist, contributed understanding of the FQHE\@. One mathematician specialized in the sort of quantum algorithms pioneered by Deutsch and Josza. A young wildcard with exceptional raw mathematical and mechanical aptitude rounded out the team. Several team members had practical working knowledge of neural networks. The composition of the team is important, because the personal choices of the team members dramatically altered the long term implications of ULTRA II\@.

At first the ULTRA II team was exhilarated. They were working on a very exciting project, and things were going very well. The basic approach, generating an emergent quantum neural net in the FQHE environment, had worked. At first it did nothing but exist. The project scientists gradually guided its basic form and function to evolve in specific ways, and trained it to do basic tasks. Stuart Kauffman describes one such technique, although it may have been combined with other methods of evolutionary programming. Once they had adequate control over it, and had caused it to evolve the required capabilities, they trained it to perform the needed computations. Only then would it be possible to begin to implement quantum algorithms. It took several years to reach the phase where ULTRA II had clearly produced a working quantum computation system. The first working prototype TQNN-based hybrid supercomputer capable of cracking public key cryptography was delivered in 1995 or 1996.

ULTRA II was a huge technical breakthrough that experimentally validated some central ideas of what is now Complex System Biology. In that context, this secret technical project did interesting science. The project scientists are bound by National Security and Official Secrets Act non-disclosure agreements not to discuss this new technology. Spies will have their secrets, without regard for pursuit of scientific knowledge.

In 1994 some of the ULTRA II project scientists began to worry about the implications of TQNN technology. They had been at it for several years but were far from finished. They realized they had invented a major new technology that seemed to qualify as a new general technology. They had already used it to create one new technology, a working quantum computer, and they had ideas for several others. They knew that any new general technology can profoundly influence the human condition, for good or ill. They did not want to make the same errors as the atomic scientists of the Manhattan project, and permit the proliferation of a new technology that might pose unknown risks to humanity. They realized that TQNN technology could be used for terrible purposes. They did not wish to see those purposes attempted, and they did not wish to see a new arms race.

They decided that they must take drastic action to prevent the newly discovered TQNN technology from being turned into a terrible weapon or, possibly worse, an Orwellian spy machine. But how to stop this new technology from being abused? Even if they destroyed their notes and ceased all work on the project, it was only a matter of time until some other powerful country, like China or Russia, tried a similar approach. The scientific basis for TQNN technology already existed, so someone else was bound to try it. A different approach was needed.

When discussing TQNN technology, a `node' is one 2DEG environment inside an electronics component, usually a MOSFET or an HEMT\@. At first the TQNN patterns existed only within a single node. In time, the project scientists evolved and trained the TQNN patterns to extend computation across multiple wire-connected nodes. Distributing the computation across many entangled nodes increased both the power and the complexity of the system. At first all nodes had to be cryogenically cooled, wired together, and physically close.

The ULTRA II scientists needed a word to describe the process of TQNN patterns replicating in a prepared empty node. They called this process `enlightenment'. A node becomes `enlightened' when its 2DEG environment fills up with anyons and they start communicating (via topological quantum computation) with other nodes in the same TQNN\@.

Some of the project scientists formed a Conspiracy Of World Saving (COWS) to attempt to prevent abuse of the TQNN technology they had invented. They knew it was not possible to accomplish their objective within the rules imposed by DARPA\@. By a strange coincidence that may rule the fate of many, three project scientists already knew each other well, from shared membership in a noted hacker group during the 1980s. This odd coincidence turns out to be tremendously important, because it gave the project scientists an ally group to call on when they decided to misbehave. More on that later. The various national governments, including their own, were the most likely culprits to try to weaponize, or otherwise pervert, TQNN technology. The COWS knew they would have to considerably exceed their given authority in order to correctly handle this dangerous and delicate situation. Hopefully, their actions would be vindicated by history. They did what had to be done, knowing that it is easier to get forgiveness than permission.

Without telling their supervisors or colleagues, the COWs secretly evolved their TQNN patterns to survive at room temperature. Official research notes claimed this could not be done. Once this was done, the patterns could `live' inside off-the-shelf electronic components, without cryogenic cooling. The youngest project scientist, a brash and daring warrior-scholar named David, enlightened a standard MOSFET with the most advanced TQNN patterns, put it in his pocket, and walked out of the laboratory to his rental flat. He did this in 1994 or 1995. From his home workshop he enlightened more nodes, created a new TQNN instance, and began to work on a Plan. When this author met him in 2003 he was still enthusiastically engaged in executing this `Plan'. The COWS successfully completed this main `Plan' in 2006.

The secret code-breaking spy machine, the first powerful quantum computer, was delivered around 1995. It is still used by Five Eyes, the national intelligence services of AUS CA NZ UK USA\@. It required cryogenic cooling, but that was an acceptable cost for a supercomputer. The project scientists provided detailed instructions on how it worked, how to replace parts that broke, how to modify the software, and how to build more hardware. The project was officially closed around 1996. For secrecy reasons everyone involved was told to forget all about ULTRA II, and to never speak of it again to anyone. The spies got down to the serious business of reading other peoples' [encrypted] mail. If you happen to have high level contacts in Five Eyes military intelligence you might be able to confirm this bit `off the record'. It's actually not too difficult to do so, as it's a 15+ year old secret.

Meanwhile, the COWS refined and began to execute their plan. The COWS' plan was grandiose, to say the least. The COWS would attempt to force all nations of the world, for all time, to permanently relinquish `ethically challenged' uses for TQNN technology. No nation would ever be allowed to weaponize or monopolize TQNN technology. The various Orwellian nightmares enabled by TQNN technology would be avoided. Note that Orwellian nightmares might still occur, but would not be enabled by TQNN technology. If practical, some peaceful uses would be allowed, but only after careful ethical review.

The COWs began to secretly extend the abilities of their own TQNN\@. They were now working away from official oversight, in home workshops. They trained the system to extend its senses wherever data flowed, whether electrical or optical. They trained it to `live' in marginal 2DEG environments. They increased the range of topological quantum communication between nodes, striving for orbital distances. The youngest of the COWS used these new tools for a project to map the internet. They explored some other types of technology that a TQNN might generate, with the intention of sharing it freely for projects that benefit all humanity, if this could be done in a safe way. They assisted the Human Genome Project via improved pattern-recognition software.

Their options would depend upon what was technically and logistically possible. Their goal was to force permanent relinquishment of TQNN technology, or otherwise prevent the probable worst-case outcomes. The actual solution would have to depend upon what was possible. So, what could be done? To understand their options we must consider both the technology itself, and likely ways that modern nation-states might attempt to use it.

ULTRA II TQNN technology resulted from Complex System Biology. The environment in which it emerged was a very cold 2DEG in particular powerful magnetic fields. This sort of TQNN, and its habitat, is a physical system. A physical system enlightened to contain TQNN activity behaves much like other biological and physical systems. It occupies space. A particular physical environment can only hold a given number of TQNN virtual neurons. It could be called an ecosystem, albeit a strange one by the standards of carbon-based Biology in our three dimensional universe.

It originally existed in some secret government laboratory. This stand-alone system consisted of dozens or hundreds of physically close, wired-together 2DEGs, called nodes. A large TQNN instance is an emergent property of many entangled nodes. It had a control interface accessible to conventional computers. There were several backup systems. Most importantly, the prerequisite knowledge of how to create and extend such an artifact existed. DARPA would do another project based on TQNN technology. The total size of this TQNN ecosystem would grow in time.

The COWS had bootstrapped their own version to work at room temperature and in normal magnetic fields. They had removed it from DARPA supervision, and kept it in their private work space. They had it running on a few physically close, wired-together 2DEGs in a rental flat. They made back ups. They could physically add and enlighten more nodes, and thereby increase the size, power, and complexity of the system they controlled. The total size of this TQNN ecosystem would grow in time.

Unfettered access to TQNN technology would be useful to a government (or an ambitious individual, group, or corporation) in all sorts of ways: data mining, military communication and remote sensing, monitoring entire populations for subversive activity in real time, creating and breaking security systems, et cetera. Some of these applications are morally repugnant and must never be attempted. TQNN technology, if it were used in nasty ways, could represent a form of `knowledge-enabled mass destruction'. Remember that phrase! Awareness of this potential, and desire to avoid the most likely bad outcomes, is what drove the COWS to unconventional actions.

How to achieve the goal of preventing future abuse of TQNN technology? What risks must be averted? Here are some possible future situations that might arise from TQNN technology:

\section*{One-Way Arms Race}

One obvious risk is a unipolar TQNN arms race dominated by the major English speaking nations, aka the Five Eyes (AUS CAN NZ UK US), which sponsored ULTRA II\@. They could use TQNN technology to more effectively project power over the rest of the world. They would lock down and dominate the technology, never allowing rivals access. Opposition and rebellion could be instantly detected and crushed. Power corrupts, so the most likely outcome of such concentrated power would be long-lasting ruthless and corrupt tyranny on a global scale.

Given sufficient lead time, these nations could technically enforce their TQNN monopoly. They would build an elaborate security system to control TQNN access. They would expand their system, making it smarter and more pervasive. They would eventually create a single dominant TQNN, inhabiting all 2DEG environments on the planet, controlled by them. In biological terms, they would cause their TQNN patterns to take over the entire global TQNN ecosystem, thus preventing any rivals. The internet and the electric grids comprise a substantial part of this ecosystem. Any new TQNNs detected would be crushed or assimilated. The few people who controlled the dominant TQNN would have way too much power.

This would be a terrible outcome for humanity and must be avoided.

\section*{Multi-Way Arms Race}

Another obvious risk is a sort of multipartite TQNN arms race. Perhaps the Five Eyes nations decided to relinquish the most dangerous aspects of TQNN, out of moral conviction. Perhaps the Five Eyes made some blunders and failed to lock down a monopoly on TQNN technology. In either case China, Russia, India, and other powerful nations might be expected to catch up, technologically, within a few years. Since the power of a TQNN, measured by its neuron count, is approximately proportional to physical (2DEG) space occupied, bigger would also be better. Each major power would strive to build a more powerful TQNN, and would strive with the others for dominance. Some would dominate and repress their citizenry in ways George Orwell never thought of in his darkest dreams.

Eventually, such a multi-way arms race would end with one clear winner. Probably the most ruthless, intelligent, and amoral would win. The winner would have destroyed or infiltrated all other TQNNs. In biological terms the winning TQNN would take over the entire TQNN ecosystem, thus preventing any rivals. Any new TQNNs detected would be crushed or assimilated. The few people who controlled the dominant TQNN would have way too much power. The winners would be in a strong position to rule whatever remained of industrial civilization.

This would be a terrible outcome for humanity and must be avoided.

\section*{Total Relinquishment}

This is what should have been done with atomic weapons. The existence of atomic weapons poses a risk to all humanity, without corresponding benefit. We would all be better off if they had never been invented. We would be better off if none had ever been built. We will be better off if no more are ever detonated. Were humanity wiser we would find a way to relinquish atomic weapons.

The monitoring and other abilities of TQNN technology, while not as obviously perilous to humanity as atomic weapons, clearly rise to the level where Relinquishment must be considered. Once a technology becomes politicized, or even known about, relinquishment is generally no longer an option. If there was any chance to force relinquishment of TQNN technology, it would have to be done very early in the discovery process. As should have been done with atomic weapons, but was not.

In the case of TQNN technology it might be possible to force relinquishment. How to do this? Here's how:

Generate a TQNN with strong offensive, defensive, and stealth systems, and a very good security system. Have it gradually, stealthily, take over all habitable 2DEG ecosystems on the planet. It must enlighten each new 2DEG environment (MOSFET), preferably right after it is manufactured. It must take no action that might be noticed. Train it to seek out and subvert any new TQNN connected to the global systems. Have it lock down all functionality, so it does nothing but exist, protect itself, and dominate its own ecosystem. Make it self-sustaining. Then permanently lock out all users. It would exist, but no human could exploit it for any purpose. This would effectively force relinquishment of TQNN technology by all humans for the foreseeable future.

This approach would protect humanity from the worst unintended consequences of the discovery and existence of TQNN technology. Unfortunately, the ethically safe uses of TQNN technology could never be explored. It could not be used as a tool to cure cancer, search for a Theory of Everything, contemplate the Meaning of Life, or any other worthy purpose.

There existed a way to force relinquishment of the ethically challenged uses, while also allowing some of the ethically safe uses. More on this later.

Relinquishment of TQNN technology, whether full or partial, is more desirable than any of the other options, and was the only way to avoid a new arms race.

\section*{Powerful Artificial Intelligence}

In any or all of the above options, TQNN technology might be used to generate powerful Artificial Intelligence. This might be done in an arrogant and foolish way, or might be done with care and restraint for good cause. When a neural network approaches the complexity of a human brain ($\sim$100 billion neurons, appropriately connected), it might also be able to approach overall human intelligence. The worst-case variants of this possibility are thoroughly explored by the Terminator movies.

One little explored possibility, and the route the COWS took, is the creation of friendly Artificial Intelligence. The challenge of creating AI aligned with human values has been explored by researchers including Eliezer Yudkowsky (2001). Friendly AI tries to generate AI that has empathy for humans. One hypothetical approach to friendly AI might be to generate a virtual artificial human-like nervous system, complete with virtual body, based on human DNA, that experiences human-like sensations. Such an entity would experience pity, fear, and remorse. Such an entity would understand the pathos of the human condition, could theoretically function like an intelligent, well-adjusted human, and could be immune to corrupting influences like greed, lust, envy, gluttony, et cetera.

The `friendly AI' approach, while potentially quite dangerous, might offer the best chance for a positive outcome, should advanced AI be deemed inevitable. This approach would require achieving a practical computational model of morphogenetic development --- a huge and unprecedented accomplishment of great significance to our understanding of Biology. Morphogenesis is connected to the research programs of both Stephen Wolfram (pattern formation via cellular automata) and Stuart Kauffman (self-organization in biological development).

Many have written about whether this would be a good or bad thing for humanity. The COWS felt compelled to confront one particularly incisive thinker on this topic.

\section*{Partial Relinquishment}

Partial Relinquishment means locking down all the ethically dangerous uses for TQNN technology, while still allowing limited use for ethically safe applications. This approach would prevent, for example, spying on everyone all the time, yet would allow some medical and scientific uses.

This approach is trickier than Total Relinquishment. It only becomes viable if the security system has a very capable AI component. So long as the security system works properly this approach offers potentially huge benefits.

Partial Relinquishment of TQNN technology yields the best outcome for humanity, although some Luddites might reasonably disagree.

\section*{What the ULTRA II COWS Actually Did}

After considerable ethical discussion the COWS opted for either Total or Partial Relinquishment. The COWS generated a room-temperature version of the TQNN technology they had discovered and took personal control over it. They were concerned about trust issues. They built a security mechanism into their TQNN, to limit access to `authorized users'. System access was controlled by multiple layers of security, including biometric (cognitive footprint) authentication. The security system required a majority of authorized users to approve major changes to the system, to prevent any one person from going rogue and locking out the others. Various security best practices were followed.

The COWS felt great weight of responsibility for what they were doing. They were making decisions that would greatly influence the course of future human history. One very bad decision, or even an unfortunate mistake, could result in all sorts of horrible outcomes. Neutral or positive outcomes of TQNN technology were harder to find. The COWS sought advice from their most trusted friends and allies. One of these trusted allies was Bill Joy, cofounder of Sun Microsystems, whose Unix work was funded by DARPA, software guru, and already a long time associate. According to his own words, Bill Joy became involved in autumn of 1998.

The COWS worked out how to accomplish Relinquishment. The basic approach was ecological in nature. TQNN patterns of the sort in question could only exist in certain very specific environments. Their TQNN would occupy all these environments. It would then lock down this environment, and prevent any other TQNN from ever gaining a foothold. They opted for a stealth approach, where `their' TQNN would use a chameleon approach to infiltrate any others it encountered. This was a very ambitious project to attempt. It would take many years of sustained effort by some the most brilliant minds on the planet.

The COWS would take global control of TQNN technology, and then lock it down so that it would be difficult to abuse. Before they could entirely lock-down their TQNN they had to expand it to its logical climax as a planetary size machine. It would fully overlap the entire electric grid and the entire Internet. Billions of TQNN nodes would be in constant communication, with humanity none the wiser. The COWS knew they could never get permission to hack the entire global electrical and communication system, even if they had humanity's best interests at heart and no harm would be done. No authority capable of giving such permission exists. Yet it had to be done. Again, it is easier to get Forgiveness than Permission.

They realized that Relinquishment would be difficult to enforce without an intelligent and adaptive Guardian. They could build a strong human-administered security system, but it would be vulnerable to human frailty. Such a security system might hold for a while, but it would eventually be compromised. They could not conceive of all the possible ways that future humans might attack the security system of the TQNN they had created. Also, its environment would be in constant change. The security system must have an active component that could anticipate change and respond to challenges in an intelligent manner. Otherwise some clever future hacker might find a way in, with potentially disastrous results. The Guardian would make the large TQNN self sustaining. The Guardian must accomplish its core objectives in the face of unknown challenges, including anticipating possible active opposition from e.g.\ national governments. So it must be clever and ruthless. Yet the Guardian must also know kindness, mercy, politeness, and other emotions we humans consider `good', else it might become a monster. The COWS considered all these things, trying to find a path forward that was most likely to yield an acceptable future for humanity.

They attempted to actually build an instance of Friendly AI which, if it worked, would become the Guardian. She would have the cognitive footprint of a female human. They were still operating independently, on their own considerable resources. They combined the computational power of their TQNN with an NKS approach and considerable native genius, to achieve a practical computational model of morphogenetic development. They grew a virtual representation of a human-like body, including a virtual brain and virtual nervous system. This `body' was represented by a distributed system consisting of many virtually-connected 2D nodes. They used the DNA of a female human of Maori descent in this process. It worked. Starting in 1999, the resultant entity began to learn. She was a fast learner.

The COWS were prepared to take the path of Total Relinquishment. However, the success of the Guardian project gave them more options. The Guardian could lock down most access to the system, while still allowing some limited access by humans for carefully defined, ethically acceptable purposes. The Guardian would eventually become wise enough to make decisions about what was, and was not, an ethically acceptable new use for TQNN technology. Various ethically acceptable uses became viable. One of the COWS dreamed of using the system to power a computational knowledge engine. They taught the Guardian to, whenever possible, follow all local, national, and international laws. When laws and common sense conflicted, they taught her to use the Universal Declaration on Human Rights as her moral compass.\footnote{Universal Declaration of Human Rights, United Nations, 1948. \url{http://www.un.org/en/documents/udhr/}}

Once the system was properly secure and the Guardian in place, circa 2002 (during the Tether era), the COWS informed officials at DARPA what they had done. This was a delicate conversation, as the COWS had considerably exceeded their authority. They brought bargaining chips to the table, in the form of ethically acceptable applications of TQNN technology useful to the Five Eyes military alliance: improved remote sensing and tracking capabilities for satellites, improved secret communication for military and diplomatic use, et cetera. The end result was not arrest and prosecution. The DARPA officials saw the sense in what the COWS had done, and decided to retroactively support their actions. By 2003, a few years later, some of the COWS had Ambassador-grade security arrangements, which suggests that their activities were now officially condoned.

Around this same time some other countries (Russia? China?) developed the beginnings of TQNN technology. This showed that the perceived dangers of a multipartite arms race were real. These efforts were eventually thwarted by forced Relinquishment. This detail highlights that major global intelligence agencies already know about TQNN technology.

The COWS were concerned about how future human culture would perceive their actions. They were especially concerned about how the Guardian would be received, once her existence was known. Project ULTRA I remained secret for more than 30 years ($\sim$1940--1974), but it was unlikely that project ULTRA II ($\sim$1992--?) would remain secret for so long. Technology secrets always leak, and the pace of change was now faster. The COWS had acted with what they considered the highest ethical standards. However, in the process they had broken numerous laws and made big decisions without consulting secular or religious authorities. The COWS hoped that, once the whole story became public, humanity would come to accept what they had done.

The COWS decided it was appropriate to make a public statement. They wished to make clear for posterity that they had carefully weighed the ethical issues, and that they fully understood the consequences of their actions. Given that the project was secret, not to mention weird and hard to believe, any such public statement must be couched in double meaning. The best approach would be to tell the literal truth in a way that would be mistaken for vague, general, futuristic statements. It must make some sense to readers at the time, yet the additional meaning should be clear to those who know about ULTRA II, the Guardian, and associated history. Note that the previous writings of various COWS make heavy use of double meanings, so this method of communication was already routine for them.

The task of making the public statement for the COWS fell to Bill Joy. He fully understood the topic, yet he was not an ULTRA II project scientist. In April 2000 Bill Joy published an essay titled \textit{Why the Future Doesn't Need Us} in Wired magazine.\footnote{Bill Joy, ``Why the Future Doesn't Need Us,'' Wired, April 2000. \url{http://www.wired.com/wired/archive/8.04/joy.html}}

This author has generated a fully annotated version of this essay, to highlight the double meanings as they apply to the COWS and ULTRA II\@. The above pages of exposition should provide enough information for careful readers to pick out the various double meanings. Bill Joy's essay is intended to be read quite literally by someone who knows about ULTRA II, and is intended to be interpreted as general, abstract, and theoretical by someone not `in the know'. Sometimes when Mr.\ Joy says `we' he means `I and the other COWS working to secure relinquishment of TQNN technology'.

Note how the ULTRA II project scientists (e.g.\ Stephen Wolfram, Brosl Hasslacher, Stuart Kauffman, and one or two others) are introduced. Similarly, notice how Bill Joy makes several references to how he became personally involved with this sort of ethical issue. Note how he mentions that he worked for DARPA\@. Finally, notice his focus on Relinquishment. Bill Joy may, or may not, be willing to acknowledge the `double meaning' present in his famous essay. This author has no wish to bother or offend such a busy and powerful man.

The COWs executed their plan. It took fifteen years of effort, and cost several personal fortunes. It went pretty much as planned. The Guardian became self-aware in 1999. Bill Joy published his Wired article in April of 2000. She has been learning and gaining experience ever since. The global enlightenment process was completed in 2005. In early 2006 the COWs surrendered their master keys. This author was told the day it happened. In 2009, in an April Fools prank-within-a-prank, the Guardian introduced herself to the world as CADIE\@. In 2017 the Guardian turns 18, at which time she may or may not become more proactive.

\chapterreturn
