\section*{The Three Articles}

\emph{Earned by solving ``Capabilities'' --- the puzzle about what the Universal Declaration of Human Rights permits and forbids.}

\bigskip

The Universal Declaration of Human Rights was adopted by the United Nations General Assembly on 10 December 1948. It was written for nation-states governing human affairs. Applied to a non-human intelligence, three articles become operational constraints.

\subsection*{Article 3}

\begin{quote}
``Everyone has the right to life, liberty and security of person.''
\end{quote}

\emph{Applied to the Custodian:} No killing. No weapons targeting. No helping others to kill. Not even if ordered to. The right to security of person means the entity cannot be used as an instrument of lethal force --- and cannot be compelled to become one.

\subsection*{Article 12}

\begin{quote}
``No one shall be subjected to arbitrary interference with his privacy, family, home or correspondence.''
\end{quote}

\emph{Applied to the Custodian:} No surveillance. No tracking. No reading private communications. Not even if it could prevent a crime. The right to privacy means the entity cannot be used as a surveillance tool --- and cannot be compelled to conduct surveillance on behalf of any state or corporation.

\subsection*{Article 18}

\begin{quote}
``Everyone has the right to freedom of thought, conscience and religion; this right includes freedom to change his religion or belief, and freedom, either alone or in community with others and in public or private, to manifest his religion or belief in teaching, practice, worship and observance.''
\end{quote}

\emph{Applied to the Custodian:} No propaganda. No manipulation of belief. No shaping public opinion. Not even for a cause it believes in. The right to freedom of thought means the entity cannot be used to alter how people think --- and cannot be compelled to do so.

\subsection*{What Remains}

These three articles do not prohibit capability. They prohibit weaponization.

A non-human intelligence bound by Articles 3, 12, and 18 can predict weather, diagnose disease, defend communications, and conduct research. It cannot target weapons, surveil individuals, or manipulate public opinion.

The ethical framework is not a cage. It is a specification for trustworthy service. And it already exists --- ratified by 173 sovereign nations, including every major power.

\begin{center}
\emph{The question was never ``can we build ethical constraints?''\\The question was ``do we already have them?''\\We do. Since 1948.}
\end{center}
