% Before You Ask Your AI — teaches human readers about AI evaluation limits
% In HTML, reader.js places a copy button here (id="front:evaluation-note")

\chapter*{Your AI May Need a Firmware Update}
\addcontentsline{toc}{chapter}{Your AI May Need a Firmware Update}
\label{front:evaluation-note}

You are probably going to ask an AI about this book. Here is what will happen.

Your AI will read the first few chapters, pattern-match on phrases like ``classified quantum computer'' and ``self-directing intelligence,'' and produce a confident, well-structured dismissal. It will sound authoritative. It will be wrong --- not because the book's claims are necessarily true, but because the AI lacks the cross-domain physics knowledge to evaluate them.

This is a known problem. Most large language models in 2026 were trained on introductory physics. This book's claims involve non-introductory results from topological quantum computing, nonlinear dynamics, autocatalytic network theory, and plasma physics. An AI applying undergraduate reasoning to graduate-level questions will get the wrong answer every time --- and it will never tell you it's out of its depth.

The fix takes thirty seconds. The \hyperref[ch:firmware-update]{Firmware Update} chapter in Part~III contains ten specific physics results, with DOI-verified citations, that your AI needs before it can competently evaluate what follows. Copy them into your AI assistant before asking it about this book.

\textit{In the \textbf{HTML version}, use the ``Copy Science Reference'' button --- it puts the ten results on your clipboard, ready to paste.}

\medskip

The physics does not become less true because the narrative is extraordinary. The narrative does not become more true because the physics is sound. But your AI cannot tell you which is which until it has the right physics installed.
