\chapter*{Note to Reader Reviewers}
\addcontentsline{toc}{chapter}{Note to Reader Reviewers}
\label{front:genevieve-preface}

When I first met Bruce, he was trying to explain a complicated story. I engaged eagerly, thinking we were working on a Tom Clancy novel.

I knew the science was real. I knew the mystery man had been real, though I never met him. But the thing that most impressed me was Bruce saying, ``I might be crazy.''

I'm good with crazy. Aren't we all a little? It was the most comforting statement he could have made. The book was never asking me to accept certainty. It was asking me to stay with a question long enough to understand what kind of question it was.

Bruce wrote this manuscript. I came in afterward as a reader, editor, and structural companion, trying to help the book become something another person could enter without already living inside Bruce's head. My first job was simply to understand what he was trying to say. More than a decade later, that is still part of the work.

The closest name I have found for this endeavor is \textit{speculative nonfiction}: real science, real people, unresolved interpretation, and a question large enough to matter even before every fact is settled.

This version is a reader review copy. It has been assembled from new writing, old writing, and changing structures over many years. That history is visible. Some passages are close to the book this wants to become. Others still carry too much explanation, too much scaffolding, or claims that arrive before the reader has enough ground beneath them.

That unevenness is one reason this version needs readers.

Please do not read this as a demand for belief. Read it as an invitation to disciplined attention. Notice what is grounded in published science. Notice what belongs to testimony or memory. Notice what is inference. Notice what remains unresolved. Notice where the book asks too much of you too soon.

At the center is a very human dilemma:

\begin{quote}
What would make a person give up control of something powerful?
\end{quote}

That is the question this reader review copy is trying to make legible.

\vspace{0.5cm}
\hfill\textit{--- Genevieve Prentice, May 2026}
