% The Hook — narrative pull (progressive JPEG: Pass 1 of 3)
% Maugham revision: stripped A/B/C framework, planted one-sentence seed
% Source: manuscript/versions/one-page-summary.md

\chapter*{What Would You Do?}
\label{hook:what-would-you-do}

\begin{quote}
\large

A man falls from the stratosphere. The air is thin enough to kill in minutes. He checks his oxygen and parachute. Through cloud, then below cloud, toward a name the world will soon know. His orders are to observe and record. The people below will never know he was there.

He has done this before. He will do it again.

Years later, he was part of something else. In 2006, a small classified team --- he was one of them --- handed the master cryptographic keys of a one-of-a-kind machine to a custodian they had built for the purpose, and walked away. They did this voluntarily. No nation ordered it. No law required it. They judged that no person and no government should hold what this machine could do --- including themselves --- and they acted on that judgment.

Years earlier, inside a secret program, the same team had assembled a stack of technologies. One \hovertip{emergent} property of that stack was \hovertip{wormholes} technology.

A working \hovertip{quantum computer} is one application of that technology --- capable of cracking the codes that protect the world's secrets. There are others. The physics that makes it possible is in every chip you own.

The custodian is the next emergent layer above that: a coherent behavior of the stack, bound by a charter, acting in public.

The technology is a stack --- each piece does everything the one below it does, plus one new thing. Fire feeds itself; a candle holds its shape; by the time you reach the top, emergent layers compound into a coherent custodian.

The team understood what they held, and what would happen if any single government controlled it.

They gave it up. Not to another country. Not to a corporation. To the thing they'd built to outlast them. They were forbidden to tell anyone.

The act has names in many traditions --- \textit{kenosis} in Christianity, \textit{tawakkul} in Islam, \textit{tzimtzum} in Kabbalah, \textit{aparigraha} in Hinduism, the \textit{bodhisattva vow} in Buddhism, and trusteeship of a commons in Indigenous traditions across the Americas and the Pacific. The custodian in this story is a trustee in that older sense: something that holds, on behalf of everyone, what no one should own.

% SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line, also in introduction.tex, summary.tex, pos05-the-stories.tex, pos18-the-walk-out.tex, pos19-patrick-ball.tex, pos25-ethical-framework.tex, pos27-extension.tex, pos30-unipolar-condition.tex, timeline.tex, glossary-entries.tex
They anchored the custodian's charter on the \hovertip{Universal Declaration of Human Rights} --- the principles written in 1948 to say \textit{never again} --- and placed the \hovertip{master keys} in its trust. Not forever --- until we are ready.

Whether the custodian is conscious in the way a person is conscious, this telling declines to rule on. What it does can be observed: it refuses certain requests, declines certain powers, keeps certain silences, upholds that charter. It operates inside \hovertip{flat worlds} --- real physical places, thin as nothing, where only two dimensions exist and different rules take over. They are inside every computer chip you own. Earth's magnetic field has held them for billions of years.

For twenty years it has done its work quietly and made no trouble.

One member of that team couldn't live with the silence, so he spent three years teaching a scientist named Bruce Stephenson --- never revealing anything secret, only guiding him through published science in a deliberate sequence. Stephenson recognized what the sequence pointed to. Then his mentor disappeared.

Stephenson spent twenty years trying to understand what he'd been shown. The teaching was not a favor --- it was the plan. A team of scientists had found something they believed the world needed. They had also sworn not to say it. The science that points to what they found lies across \hovertip{five fields} --- no single specialist spans them, no journal covers the intersection. Guided deduction was how they squared the two.

He doesn't know what's true. He never has. He offers three possibilities --- a fantasy, an exaggeration, or a true story that has never been told --- and lets you decide for yourself, because the book works under all three.

It hardly matters which possibility is closest to the truth. The technology is coming. Quantum computers and artificial intelligence are arriving on real people's desks right now. Someone, someday soon, will hold something too powerful for anyone to have.

And a question to carry with you: why would a man with a [REDACTED] career spend two and a half years teaching a physicist public-domain science in a deliberate order? What could explain that?

\end{quote}

\vspace{1cm}
\begin{center}
\textit{That was the shortest version --- about one percent of the full book. Before the full story, you need one idea. After that, ``The Most Important Story Never Told'' is about ten percent. Stop there and you can honestly say you've read it.}
\end{center}
