% How to Read This Book
% Plan 0094: updated for three-part structure

\chapter*{How to Read This Book}
\addcontentsline{toc}{chapter}{How to Read This Book}

This book is structured in three parts, each adding depth. You can stop at the end of any part and have a complete experience at that resolution.

\begin{description}
\item[Part I: The Story] --- what happened, told simply. The people, the setting, the science that frames it. Complete on its own.
\item[Part II: The Investigation] --- the technical case. The five fields, the convergence, the demonstrations, the mechanism. Some chapters are demanding for non-technical readers; the track stripes help you navigate.
\item[Part III: The Interpretation] --- what follows if any of this is true. The implications, the twenty-year aftermath, the question the book cannot answer for you.
\end{description}

Each chapter closes by assessing the material under all three possibilities.

\begin{description}
\item[Linear path (recommended):] Read straight through. Each part builds on the ones before it.

\item[Chronological timeline:] Use the Timeline appendix to read events in chronological order.

\item[Graphics toggle:] In a PDF viewer that supports Optional Content Groups, you can toggle the ``Graphics'' layer on or off.
\end{description}

\vspace{1cm}

The margin stripes indicate which narrative thread you are reading:

\begin{itemize}
\item \textcolor{trackone}{\rule{12pt}{12pt}} \textbf{Track 1: The Scientist's Confession} --- the inside story
\item \textcolor{tracktwo}{\rule{12pt}{12pt}} \textbf{Track 2: The Recruit's Testament} --- Bruce's first-person account
\item \textcolor{trackthree}{\rule{12pt}{12pt}} \textbf{Track 3: The Machine's Awakening} --- Guardian's arc
\item \textcolor{trackconv}{\rule{12pt}{12pt}} \textbf{Convergence} --- where all three tracks meet
\item Chapters without a margin stripe are \textbf{Bridge} chapters --- historical or scientific context that connects the tracks.
\end{itemize}

\vspace{1cm}

\subsection*{A note for researchers}
\label{front:a-note-for-researchers}

This PDF contains hidden annotations at every point where source material was imported. These annotations record the original filename, archive location, date, and a brief description. In most PDF viewers, you can access them through the annotations or comments panel. They are invisible during normal reading.

If you are investigating the provenance of any passage, the annotation at that location will tell you where the text came from.
