\chapter*{A Note on Names and Sources}
\addcontentsline{toc}{chapter}{A Note on Names and Sources}

\textbf{Real names.} This book uses real names because anonymized claims cannot be checked. Every person named here is either a public figure discussed in their public capacity, or a person whose involvement is described under the three-possibilities framework presented in Chapter~\ref{pos01:three-possibilities}. Under Possibility~A, none of the classified activities described occurred --- the names attach to public-domain facts only. Under Possibilities~B and~C, some or all of the described activities occurred as presented.

\textbf{Reported speech.} All quoted dialogue is reconstructed from the author's memory, not transcribed from recordings. The author's memory is twenty years old and unreliable on exact wording. When the text says ``Healer said X,'' it means the author's best recollection of approximately what was communicated. The emotional register and substance are more reliable than exact words.

\textbf{Fair use.} This book excerpts published works --- including Bill Joy's ``Why the Future Doesn't Need Us'' (\textit{Wired}, 2000), Stuart Kauffman's published research, and other materials identified in the text --- under the fair use doctrine (17~U.S.C.\ \S~107) for commentary, criticism, and scholarship.

\textbf{AI-assisted drafting.} This book was written with structural assistance from Claude (Anthropic). The methodology is described in the Engine chapter. Sections marked with a dagger (\dag) are AI structural drafts based on the author's source material. All claims, judgments, and editorial decisions are the authors'.
