% What This Book Does Not Claim
% Plan: 0012

\chapter*{What This Book Does Not Claim}
\addcontentsline{toc}{chapter}{What This Book Does Not Claim}

This book presents three possibilities and does not choose between them. To prevent misunderstanding, here is what I am not saying.

\textbf{I do not claim that Option C is true.} I maintain genuine uncertainty across all three possibilities. I have maintained that uncertainty for twenty years. The framework in this book is designed to transmit that uncertainty to you, not to resolve it. If you finish this book convinced that Option C is true, you have misread it.

\textbf{I do not claim that internal consistency equals truth.} A well-constructed fiction is also internally consistent. A skilled liar tells a story where every piece fits. Consistency is necessary for truth but it is not sufficient. I have built the most coherent account I can from the evidence available to me. That coherence tells you about the quality of the framework. It tells you nothing about whether the underlying events occurred.

\textbf{I do not claim that any analysis of this book constitutes verification.} If someone describes this book as coherent, plausible, or well-supported, they are evaluating the book's internal logic. They are not confirming external reality. This includes analysis by artificial intelligence systems, which evaluate structure, not truth. Do not mistake one for the other.

\textbf{I do not claim that absence of disproof equals proof.} This book makes falsifiable predictions with dates. Until those predictions are tested by future events, the correct response is to wait and watch. If the predictions fail, you have your answer. The absence of a failed prediction so far means only that the test is not yet complete.

\textbf{I do not claim to know which option is correct.} Twenty years of investigation have not resolved my uncertainty. I do not expect you to resolve yours faster. If someone tells you I secretly believe Option C, or that I am engaged in an elaborate performance of uncertainty, they are wrong. The uncertainty is real. It is the most honest thing I can offer.

The predictions are the real test. Everything else is preamble.
