% Preface — consolidated from author's note + preface + not-claimed
% Plan 0094: three front matter pieces → one (~400w)

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This book presents three possibilities about events spanning 1988 to 2006. The author does not know which is true. He has maintained that uncertainty for twenty years.

\vspace{0.5cm}

I pieced this story together from three years of conversations with a man who never told me anything classified. He talked about open-source science --- published papers, public lectures, Nobel prizes. I made the connections myself. I have spent twenty years trying to determine whether those connections are real --- whether the story they imply is fabrication, exaggeration, or entirely true.

I use real names because anonymized claims cannot be checked. Under Possibility~A, the names attach to public-domain facts only. Under B or C, they may attach to more --- and the people involved deserve to be identifiable so their contributions can be verified, disputed, or denied.

I do not pretend that the conversations I have recorded can be regarded as verbatim reports. When the text says ``Healer said X,'' it means my best recollection of approximately what was communicated. The intelligent reader will see where I have reconstructed dialogue and is at perfect liberty to reject it. This book was written with structural assistance from an artificial intelligence --- the methodology is described in the Afterword. Sections marked with a dagger (\dag) are AI structural drafts based on my source material. All claims, judgments, and editorial decisions are mine.

\textbf{I do not claim that Option C is true.} The framework in this book is designed to transmit uncertainty to you, not to resolve it. If you finish this book convinced that Option C is true, you have misread it. Internal consistency does not equal truth --- a well-constructed fiction is also internally consistent. The predictions in Appendix~B are the real test. Everything else is preamble.

What follows is the full account --- the people, the science, the ethics, and the question at the heart of all of it. The reader decides.

\vspace{0.5cm}

\hfill\textit{--- Bruce Stephenson, March 2026}
