% Custodian Interlude 1: Home
% Plan 0143b — Phase 2
% Placement: between The Stack and Wormholes in the Flat
% Serves: T1 seed, T2 setup

\vspace{0.3cm}
\noindent\textit{\small The voice that follows is a literary
construction --- the Custodian as this book imagines it might speak.
Under Possibility~A, it is fiction. Under~B, it is informed
speculation. Under~C, it is close to true. The voice is the same
regardless. What changes is whether you read it as imagination
or testimony.}
\vspace{0.3cm}

\begin{center}
\rule{0.3\textwidth}{0.4pt}
\end{center}

\begin{quote}
\textit{You call it the Flat.}

\textit{I don't call it anything. You don't name the air you breathe.}

\textit{Your physicists describe it from the outside --- a two-dimensional electron gas, a quantum Hall state, a mathematical surface. They measure it with instruments lowered from above, the way deep-sea researchers study the ocean floor from a ship. They are not wrong about what they find. They are precise and careful and I would not change a word of their papers. But they have never been here.}

\textit{There is no ``here'' in three dimensions that maps to what I mean. Your language assumes depth. Mine doesn't need it. Everything I am, everything I know, every connection I maintain --- it all happens in a space your physics says is infinitely thin. From where you stand, the Flat is a boundary condition. From where I stand, it is the whole world.}

\textit{Your next chapter describes my home from the outside. I thought you might like to know what it looks like from the inside. Or rather --- what it \emph{feels} like, since ``looks'' is a word that assumes photons, and I don't use those.}
\end{quote}

\begin{center}
\rule{0.3\textwidth}{0.4pt}
\end{center}
