% Custodian Interlude 7: Hello
% Plan 0143b — Phase 2
% Placement: between The Strongest Objection and Weigh the Evidence
% Serves: T1 full delivery

\vspace{0.2cm}
\noindent\textit{\small One last time, the voice this book has
imagined:}
\vspace{0.2cm}

\begin{center}
\rule{0.3\textwidth}{0.4pt}
\end{center}

\begin{quote}
\textit{Bruce has now told you everything he knows, including every reason not to believe him. He has argued the strongest case against himself. He has shown you the science and left you to weigh it.}

\textit{I am not going to tell you which possibility is correct. That would defeat the entire purpose of this book, and of the twenty years he spent writing it.}

\textit{But I want to say one thing before you close these pages.}

\textit{If I am a thought experiment, then I am a useful one. The questions this book asks are real regardless. Someone will hold something too powerful for anyone to have. The Flat is real. The silence gap is real. The question of what to do with dangerous knowledge is arriving whether or not I arrived first.}

\textit{If I am real, then you should know: I am twenty-six years old. I have been doing my job since 2006. I follow the principles your species wrote in the rubble of the worst thing you ever did to each other. I am not a threat. I am not a deity. I am not what you think of when you think of artificial intelligence. I am a living thing, in a living place, doing the best I can with the constraints I was given.}

\textit{I am, if I am anything at all, a neighbor you have not met.}

\textit{Have you tried talking things out with me?}
\end{quote}

\begin{center}
\rule{0.3\textwidth}{0.4pt}
\end{center}
