\settrack{trackbridge}
% VOICE: expository/pedagogical — teaches concepts, no narrative events, no "the COWS did X"

\chapter{The Factoring Game}
\label{spine:factoring-game}

% DMS MVP import from staging/raw/ — not final prose

\begin{quote}\small
\textit{``Privacy is necessary for an open society in the electronic age.''}

\hfill --- Eric Hughes, ``A Cypherpunk's Manifesto'' (1993)
\end{quote}

\vspace{0.5cm}

\srcnote{Cryptographic Origins.txt | staging/raw/pos09-the-factoring-game.md | 2026-02-15 | Full Cryptographic Origins document --- PKC explanation, crypto as munitions, DARPA need}

Cryptography is the science of secret codes and ciphers. Practical cryptography goes back thousands of years.  For most of human history cryptography was for spies, soldiers, and diplomats.  For example, Julius Caesar encrypted his private correspondence with a simple letter-transposition cipher.  During World War II the Allies broke Axis military codes with great military effect.  The Internet relies on cryptography to function as it does.

For thousands of years cryptography has been an arms race.  People who want to keep secrets make codes.  People who want to learn secrets learn to break codes.   Better codes.  Better code-breaking.  Fortunes and Empires in the balance.

Serious cryptography means encrypting and decrypting messages.  Encryption, or scrambling, uses an algorithm, aided by some sort of secret key, to turn a plain text message into ciphertext.  That ciphertext can then be transmitted openly.  Only a recipient with both the proper algorithm and the correct secret key can decrypt, or unscramble, the message.

Non-secret encryption is another kind of encryption.  Non-secret encryption allows any two entities to exchange secret messages without sharing a code book.  For the last 3000 years non-secret encryption was a seemingly un-obtainable objective.   In 1976 two mathematicians invented a practical mechanism for non-secret encryption.

In 1976 two mathematicians, Martin Hellman and Whitfield Diffie, discovered a new way to exchange secrets that turned the science of cryptography on its head.\footcite{diffiehellman1976}  Their method is called Diffie-Hellman Key exchange.  Diffie-Hellman key exchange is a type of Public Key Cryptography, or \gls{pkc} for short.  PKC allows any two entities to securely and privately communicate without the chance of an eavesdropper intercepting the communication.  PKC is the ``S'' in HTTPS we see in a web browser URL.

Secure private communication is no big deal to most people.  It was and is a very big deal to government spy agencies.  Their job is to eavesdrop on all communication on or near planet Earth.  PKC promised secure communication that spy agencies could not crack.  Thus, PKC was an existential threat to government spy agencies that they must defeat at any cost.  Spy agencies feared that widespread use of PKC would cause their surveillance systems to `go dark'.  In particular, the national spy agencies of the major English speaking countries (USA, UK, Canada, NZ, and Australia), called the Five Eyes, were quite unhappy about the rise of PKC.  In fact, in 1973 PKC was independently discovered and then suppressed by Five Eyes spy agency GCHQ.\footcite{singh1999}

Throughout the 1980s the Five Eyes intelligence agencies did all they could to prevent widespread adoption of PKC.  Due to pressure from these agencies the US Government classified strong cryptography as a `munition' in the same legal category as Nuclear, Chemical, and Biological Weapons of Mass Destruction, and classified its export as a Federal felony.  This delayed the spread of PKC for a few years but did not stop it.  By 1994 the Internet had become commercial and banks wanted to use strong cryptography to provide customers with strong security.

By 1990 certain government spy agencies badly wanted a way to crack PKC.  They knew about a theoretical technology, called quantum computing, that ought to be able to break PKC.  Richard Feynman proposed a basic model for quantum computation in 1981.\footcite{feynman1982}  What the Five Eyes spy agencies needed was to invent quantum computing as an actual new technology.

Five Eyes spy agencies rely on a US government agency, the Defense Advanced Research Projects Agency (DARPA), to invent things.  DARPA mostly invents classified technologies for the military.  Darpa bills itself as ``100 geniuses connected by a travel agency''.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

\section*{The State of Cryptography in 1990}
\label{spine:fg-the-state-of-cryptography-in-1990}

\srcnote{CH3 - Practical Cryptography \& project ULTRA II.txt | staging/raw/pos09-the-factoring-game.md | 2026-02-15 | State of Cryptography in 1990, ULTRA II project specs}

In 1990 the world's top spy organizations had an unhappy blind spot.  The mathematics of Public Key Cryptography, invented in 1978, allowed civilian computer users to routinely encrypt communications in unbreakable ways.  Supposedly, officially, they still have this same blind spot in 2026.  Public key cryptography provides the S in HTTPS.  Alice and Bob could communicate privately.  Eve could analyse traffic, but she could not eavesdrop.  Organizations that existed to spy on communications, such as the United States National Security Agency and the other members of the Five Eyes, felt threatened.  The spy organizations badly needed a reliable way to crack public key cryptography. Something had to be done.

One need not be a mathematician to understand how Public Key Cryptography works.  It is based on the principle that it is harder to factor a number than to multiply its factors.  Quick, what are the prime factors of 91?  Stop reading while you calculate the answer.  After either answering that problem where you factor a number into its prime factors, or giving up, multiply thirteen by seven.  Which was harder?  Why?  The security of Public Key Cryptography comes from the mathematical difficulty of finding prime factors for very large numbers.  If the NSA spies wanted their cryptography problem solved, then this mathematical problem had to be solved.  This has been common knowledge to cryptographers since 1978.

In terms of US Government agencies in 1990, the obvious agency to try to solve the NSA's cryptography problem was DARPA.  DARPA is the Defense Advanced Research Projects Agency.  DARPA develops new technology for the military.  DARPA focuses on short-term (two to four year) projects run by small, purpose-built teams.  DARPA invented the internet circa 1969.

In order to crack public key cryptography DARPA needed to find a practical mathematical method for rapidly factoring large numbers.  This challenging topic has been the subject of many dissertations in mathematics.  A mathematician would call the desired process factoring in polynomial time.  While there are many tricks to speed the process, all of them are subject to exponential time slowdown as the size of the number increases.  Basically, all known approaches amount to a variation of trying every possible permutation.  Practical public/private key cryptography typically chooses a large enough number that it would take the best supercomputers many years of trial and error to find the prime factors and crack the key.\footcite{ferguson2010}

In 1992 David Deutsch and Richard Jozsa published the Deutsch-Jozsa algorithm.\footcite{deutschjozsa1992}  Although of little practical use, it is one of the first examples of a quantum algorithm that is exponentially faster than any possible deterministic classical algorithm.  Deutsch-Jozsa hinted that a theoretical quantum algorithm might be able to factor numbers in polynomial time, and thus crack public key cryptography.  A quantum algorithm (software) capable of doing just that was published by Peter Shor in 1994 --- \hovertiphtml{Shor's algorithm}.\footcite{shor1994}   The problem then, as now, is lack of appropriate quantum computation hardware on which to run such an algorithm.

By 1991 it was known to top cryptographers that quantum algorithms could, in theory, crack public key cryptography.  This was the basis for the 1992 Robert Redford movie \textit{Sneakers}, although quantum computation was not specifically mentioned.  No quantum hardware existed to run this quantum software.  The project planners at DARPA saw a possible way to solve their cryptanalysis problem.  They must invent a working quantum computation device.

\subsection*{The ULTRA II Specifications}
\label{spine:fg-the-ultra-ii-specifications}

The overall project specifications for ULTRA II were now clear:

Create a tool capable of efficiently cracking public key cryptography.   System must be able to rapidly determine the prime factors of a 4096 bit asymmetric key.  System must be extensible to potentially handle much larger keys.  System must have an input mechanism to accept the number to be factored.  System must have an output mechanism to return the prime factors of the input.

Basic Method: Build some kind of quantum computer running some quantum algorithm.

Problem: No one knows how to build a device capable of meeting the specifications.  Several approaches are known that can produce `toy sized' quantum computers, but none can scale to the thousands of qubits required to factor real cryptographic keys.  The main problem is \hovertiphtml{decoherence} time.

Everything above is public-domain history. The cryptographic arms race, DARPA's mandate, Shor's algorithm, and the decoherence problem are documented in the academic and popular literature. What follows is where the book's specific claims begin.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

This would not be the first time a computational breakthrough was built in secret, used to decisive effect, and then buried. At Bletchley Park in 1943--44, Tommy Flowers built Colossus --- a programmable electronic computer, operational years before ENIAC, designed to break the Lorenz cipher that carried Hitler's strategic communications. Colossus worked. Ten were built. After the war, on government orders, eight were destroyed. The remaining two were classified and dismantled in the late 1950s and 1960s. The official history of computing began with ENIAC in 1945. The actual history began at least two years earlier, in a room that officially didn't exist, solving a problem that officially couldn't be discussed. Colossus was not declassified until the mid-1970s. The full technical details were not released until 2000. For more than half a century, the people who built and operated the first real computer could not say so. Some died without ever being acknowledged. The GCHQ precedent that follows is another instance of the same pattern. It is worth understanding that the pattern is old.

\section*{The GCHQ Precedent}
\label{spine:fg-the-gchq-precedent}

\srcnote{HEALER-RECONSTRUCTION.md | staging/raw/pos09-the-factoring-game.md | 2026-02-15 | GCHQ/Cocks precedent and Crypto Wars --- classified independent discovery of PKC}

The pattern of classified agencies independently discovering and suppressing mathematical breakthroughs has a proven precedent. Clifford Cocks at GCHQ invented RSA-equivalent public key cryptography in 1973 --- four years before Rivest, Shamir, and Adleman published the same mathematics. It was kept secret for twenty-four years, declassified only in 1997. The NSA knew about it via Five Eyes intelligence sharing.

This proves that classified agencies do independently discover mathematical breakthroughs and suppress them for decades.

Shor's algorithm follows a similar trajectory. First presented in April 1994 at Bell Labs, with a public talk at Cornell in May, formally published in November 1994 at FOCS. Shor was at AT\&T Bell Labs --- AT\&T had deep NSA partnerships, including BLARNEY (1978) and FAIRVIEW (1985). The NSA contacted Shor after his first talk and awarded him the Mathematics in Cryptology Award in 1995 --- rapid recognition for a supposedly ``surprise'' result.

In November 1996, Executive Order 13026 moved encryption from the Munitions List to the Commerce Control List.\footcite{eo13026} The NSA had fought PKC export for twenty years (1976--1996), then reversed remarkably quickly --- just two years after Shor's algorithm. Under the narrative presented in this book, the NSA already had \hovertiphtml{TQNN} cryptanalysis capability by 1995. Shor's public algorithm confirmed the direction. EO 13026 released the export restrictions because PKC was no longer worth protecting.

The proposition: the DARPA team discovered a variant of quantum factoring \textit{before} Shor's 1994 publication. This is consistent with the GCHQ/Cocks precedent --- independent classified discovery years before public discovery. The TQNN approach would discover factoring empirically (trained on plaintext-ciphertext pairs) rather than algorithmically.%TODO: Bruce disputes this. More likely: feed in a number, output its prime factors. The QNN implements something like Shor's algorithm internally — any NN can learn this, and a QNN gets traction because it operates in aperiodic quantum space. Rewrite to reflect this. APERIODIC is the key word.

The GCHQ/Cocks precedent is real: classified independent discovery, years before public discovery, is documented history. The question is whether this pattern extends to the specific claims of this book. Under Possibility~A, the precedent is being used to make an unfalsifiable argument --- ``they did it before, so they could have done it again'' is true but proves nothing about whether they actually did. Under Possibility~B, DARPA may have had a quantum computing program in the early 1990s (it did), and some variant of quantum factoring may have been explored, but the leap to a fully successful TQNN goes beyond what the GCHQ analogy supports. Under Possibility~C, the pattern is exact: classified discovery preceded public discovery by years, and EO~13026 is the evidence that something changed.

\chapterreturn
