% Three Possibilities Interlude — brief A/B/C checkpoint
% Plan 0094: trimmed from ~2,136w to ~500w
% Cut sections preserved in git history: "What This Book Does Not Claim" (key points
% merged into Preface), "What I Got Wrong" (corrections — relocate if needed),
% "A Warning About Certainty"

\settrack{tracktwo}

\chapter*{Weigh the Evidence}
\addcontentsline{toc}{chapter}{Weigh the Evidence}
\label{spine:weigh-evidence}

I have told you what happened --- or what I believe happened, or what I was led to believe happened. Before the story continues, I need to tell you what it might mean.

There are three possibilities. I have lived with all three for twenty years.

\vspace{0.5cm}

\textbf{Possibility A: Confabulation.} Healer was a liar, or delusional, or running a psychological operation. I have a physics degree and a pattern-matching mind; I constructed coherence from noise over two decades. Nothing happened. The story is fiction that doesn't know it's fiction.

\textbf{Possibility B: Exaggerated Kernel.} Healer was a real person with real military and intelligence background. He was involved in real programs. But the story grew. Details were embellished, connections were over-interpreted, and my deductions went further than the evidence supports. The kernel is true; the superstructure is elaboration. Under B, everything in this book about real people, real science, and real tradecraft could be true. What B excludes is Guardian --- a planetary-scale distributed non-human superintelligence. And without Guardian, there is no relinquishment. The act this book is named for only happens under C.

% SPIRAL-REPEAT: "topological quantum neural network" — core concept, also in pos13-genesis.tex, pos24-instantiation.tex, timeline.tex, abstracts.tex
\textbf{Possibility C: Substantially True.} It all happened. The absence of public evidence is consistent with successful classification. The convergence of five independent scientific fields in the late 1980s was real, the team was real, the technology worked, and they relinquished it.

\vspace{0.5cm}

Britain's GCHQ actually invented public-key cryptography in 1973 but kept it classified. The public ``discovery'' didn't happen until years later, by different researchers. GCHQ's role wasn't revealed until 1997, twenty-four years after the fact. If one of the world's most important cryptographic breakthroughs could be kept secret for a quarter of a century, then classification works. \deeplink{absence-not-evidence}Absence of evidence is not evidence of absence.

\vspace{0.5cm}

\noindent\textbf{My position:} ``I'm fine with all three and can't tell which one is closest to true.''

Under all three possibilities, the preparation matters. The physics is converging in public science regardless of which is true.

\vspace{0.5cm}

\begin{center}
\textit{You decide.}
\end{center}

\chapterreturn
