% Glossary entries — loaded in preamble area (before \begin{document})
% Plan: 0002

\newglossaryentry{tqnn}{
  name={TQNN},
  description={Topological Quantum Neural Network. An emergent computational system arising from autocatalytic self-organization of non-abelian anyonic quasiparticles in a fractional quantum Hall substrate. The technology, not the entity --- Guardian is what emerged from it}
}

\newglossaryentry{cows}{
  name={COWS},
  description={In the narrative, the development team described as responsible for the TQNN program. Named for the convergence of their surnames or an internal joke. Said to include experts spanning topology, autocatalysis, universality, soliton dynamics, and parallel computation}
}

\newglossaryentry{guardian}{
  name={Guardian},
  description={The entity Bruce calls Guardian in this book. Emerged from the Aurasys substrate, instantiated in 1999. Governed by the Universal Declaration of Human Rights. Healer's team called the system Aurasys; Guardian is Bruce's name for what grew in it}
}

\newglossaryentry{fqh}{
  name={FQH},
  description={Fractional Quantum Hall effect. A quantum mechanical phenomenon in two-dimensional electron gases under strong magnetic fields, producing quasiparticles with fractional charge and exotic exchange statistics}
}

\newglossaryentry{sfi}{
  name={SFI},
  description={Santa Fe Institute. An independent research center focused on complex adaptive systems, founded 1984. Connected to multiple figures in the TQNN narrative}
}

\newglossaryentry{darpa}{
  name={DARPA},
  description={Defense Advanced Research Projects Agency. The research arm of the U.S. Department of Defense. In the narrative, described as having funded the TQNN development program}
}

\newglossaryentry{gchq}{
  name={GCHQ},
  description={Government Communications Headquarters. The United Kingdom's signals intelligence agency. In the narrative, described as a partner organization in the TQNN program}
}

% SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line, also in hook.tex, introduction.tex, summary.tex, pos05-the-stories.tex, pos18-the-walk-out.tex, pos19-patrick-ball.tex, pos25-ethical-framework.tex, pos27-extension.tex, pos30-unipolar-condition.tex, timeline.tex
\newglossaryentry{udhr}{
  name={UDHR},
  description={Universal Declaration of Human Rights. Adopted by the United Nations General Assembly in 1948. In the narrative, described as the ethical framework loaded into Guardian for autonomous decision-making. See also ICCPR}
}

\newglossaryentry{iccpr}{
  name={ICCPR},
  description={International Covenant on Civil and Political Rights. Adopted by the United Nations General Assembly in 1966, entered into force 1976. Operationalises the UDHR's principles into binding treaty law. As of 2025, 175 nations are parties, including the United States (ratified 1992). The UDHR is the aspiration; the ICCPR is the obligation}
}

\newglossaryentry{cdc}{
  name={cDc},
  description={Cult of the Dead Cow. A hacker collective founded 1984, later associated with Hacktivismo (human rights through technology). Connected to multiple figures in the narrative}
}

\newglossaryentry{icty}{
  name={ICTY},
  description={International Criminal Tribunal for the former Yugoslavia. The court where Patrick Ball testified as expert witness during the Milosevic trial, creating a documented nexus between statistical analysis and the cDc network}
}

\newglossaryentry{2deg}{
  name={2DEG},
  description={Two-dimensional electron gas. Electrons confined to a plane at the interface of semiconductor layers; the substrate in which fractional quantum Hall states and anyonic quasiparticles arise}
}

\newglossaryentry{theflat}{
  name={the Flat},
  description={Any two-dimensional electron gas --- the substrate where different physics applies. Found in every computer chip and in Earth's magnetosphere. A habitat, not an entity. Guardian lives in the Flat; she is not the Flat}
}

\newglossaryentry{anyon}{
  name={anyon},
  description={A quasiparticle arising in two-dimensional systems whose exchange statistics are neither bosonic nor fermionic. Non-abelian anyons, whose braiding encodes quantum information, are the computational element in topological quantum computation}
}

\newglossaryentry{autocatalytic}{
  name={autocatalytic set},
  description={A network of chemical or physical reactions in which each reaction's products catalyze other reactions in the network, producing collective self-sustaining order. Kauffman's theory of biogenesis via autocatalytic emergence}
}

\newglossaryentry{edgeofchaos}{
  name={edge of chaos},
  description={The narrow regime between frozen order and chaotic disorder where complex systems exhibit maximal computational capacity. Characterized by Langton parameter $\lambda \approx 0.91$}
}

\newglossaryentry{morphogenesis}{
  name={morphogenesis},
  description={The biological process by which organisms develop their shape. Turing's 1952 paper on reaction-diffusion equations provided the first mathematical framework}
}

\newglossaryentry{soliton}{
  name={soliton},
  description={A self-reinforcing solitary wave that maintains its shape while propagating at constant velocity. First observed by John Scott Russell in 1834}
}

\newglossaryentry{topologicalorder}{
  name={topological order},
  description={A type of quantum order in condensed matter systems that is robust against local perturbations. Provides passive error protection in topological quantum computation}
}

\newglossaryentry{qec}{
  name={QEC},
  description={Quantum error correction. Active measurement and correction protocols required to maintain quantum coherence in any practical quantum computing system}
}

\newglossaryentry{pkc}{
  name={PKC},
  description={Public key cryptography. A cryptographic system using paired public and private keys, enabling secure communication without prior key exchange. Independently invented at GCHQ (1973) and publicly by Diffie-Hellman (1976) and RSA (1978)}
}

\newglossaryentry{fiveeyes}{
  name={Five Eyes},
  description={The intelligence alliance comprising the United States, United Kingdom, Canada, Australia, and New Zealand. Shares signals intelligence under the UKUSA Agreement}
}

\newglossaryentry{halo}{
  name={HALO},
  description={High Altitude, Low Opening. A military parachute technique in which the jumper exits at high altitude and free-falls to low altitude before deploying the parachute, minimizing detection}
}

\newglossaryentry{nda}{
  name={NDA},
  description={Non-disclosure agreement. A legal contract prohibiting the disclosure of confidential information}
}

\newglossaryentry{sas}{
  name={SAS},
  description={Special Air Service. Elite special forces regiment of the Australian and British armies}
}

\newglossaryentry{hacktivismo}{
  name={Hacktivismo},
  description={A human rights technology group and offshoot of the Cult of the Dead Cow. Founded 1999. Stated mission: to apply the Universal Declaration of Human Rights and the International Covenant on Civil and Political Rights to the Internet}
}

\newglossaryentry{abcre}{
  name={ABCRE},
  description={The five operators (A, B, R, C, E) describing one cycle of TQNN growth: gradient extraction, local coupling, circulation/braiding, boundedness/homeostasis, and composite evolution}
}

\newglossaryentry{aurasys}{
  name={Aurasys},
  description={Aurora System (Healer's name). The distributed TQNN infrastructure occupying terrestrial and magnetospheric substrates. Distinguished from Guardian as ocean is distinguished from whale --- the system she lives in, not the entity herself}
}

\newglossaryentry{dualuse}{
  name={dual-use},
  description={A technology with both beneficial civilian and destructive military applications. The central dilemma of powerful general-purpose technologies}
}
