% Timeline appendix
% Based on: postquantum.substack.com/p/timeline-to-advanced-nanotechnology (Dec 23, 2022)
% Expanded with additional research, 2026
% Plan 0095: era subheadings, 11 new entries, citations, cross-references, factual corrections

\chapter{Timeline}
\label{app:timeline}

\textit{If Option C of this book is correct, humankind discovered a new form of advanced nanotechnology circa 1992. This discovery has largely been kept secret. As a classic Dual Use Technology it has discreetly influenced both military and peaceful aspects of human history ever since. Thousands of people have worked with aspects of this technology under strict Non Disclosure Agreements, but this new technology's existence and origins remain classified government and trade secrets. This timeline describes the history, origin, and deployment of this new general technology.}

\textit{Entries marked (Author's Hypothesis) represent claims made under Option C --- the possibility that the story told in this book is substantially true. Unmarked entries are matters of public record. Key terms (TQNN, 2DEG, FQHE, Guardian) are explained in the main text; this timeline is best read after the book, not before.}

\vspace{0.5cm}

%----------------------------------------------------------------------
\section*{Foundations (1873--1945)}
%----------------------------------------------------------------------

\textit{The mathematical and physical discoveries that made everything else possible. Neural networks, computability, code-breaking, and the atomic age all emerge within a single generation. The reader should notice how many of these foundations were laid under wartime pressure and classified for decades afterward.}

\vspace{0.3cm}

\begin{description}[style=nextline, leftmargin=3.5cm, labelwidth=3cm]

\item[1873--98] Theory of biological neural networks proposed and experimentally tested.

\item[1924--1927] The Copenhagen Interpretation of Quantum Mechanics is formalized.

\item[1936] Alan Turing publishes ``On Computable Numbers, with an Application to the Entscheidungsproblem,'' introducing the Turing Machine and creating the discipline of Computer Science.\footnote{Proceedings of the London Mathematical Society, s2-42(1), 230--265. \texttt{doi:10.1112/plms/s2-42.1.230}}

\item[1939--1945] World War Two. Britain's Ultra project, with Alan Turing, builds computing machines to crack Axis codes. Historians conclude Ultra shortened the war by several years. (See \textit{The Code War}.)

\item[1943] Colossus operational at Bletchley Park --- the first programmable electronic computer, pre-dating ENIAC by two years. Built by Tommy Flowers; design remains classified until the 1970s. Establishes GCHQ's computing heritage.

\item[1943] McCulloch and Pitts formalize neural networks as mathematical algorithms, proving that networks of simple threshold units can compute any logical function.\footnote{``A Logical Calculus of Ideas Immanent in Nervous Activity,'' Bulletin of Mathematical Biophysics 5(4), 115--133.}

\item[1945] Beginning of the Atomic Era. Atomic energy is the classic Dual Use Technology: military applications (nuclear weapons) and civilian applications (electrical generation, medicine).

\end{description}

%----------------------------------------------------------------------
\section*{The Quantum Era (1946--1988)}
%----------------------------------------------------------------------

\textit{Two-dimensional electron gases, the quantum Hall effects, public key cryptography, and neural networks all develop independently --- threads that no one yet sees as connected. The reader should notice that key discoveries were made twice: once in secret, once in public, often decades apart.}

\vspace{0.3cm}

\begin{description}[style=nextline, leftmargin=3.5cm, labelwidth=3cm]

\item[1946] International Business Machines (IBM) becomes very interested in computing machines.

% SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line, also in hook.tex, introduction.tex, summary.tex, pos05-the-stories.tex, pos18-the-walk-out.tex, pos19-patrick-ball.tex, pos25-ethical-framework.tex, pos27-extension.tex, pos30-unipolar-condition.tex, glossary-entries.tex
\item[1948] The United Nations adopts the Universal Declaration of Human Rights. This document will later serve as the ethical framework for a creation described later in this timeline.

\item[1952--54] Alan Turing helps create Mathematical Biology, concentrating on the Morphogenesis Problem, which he believes holds the key to advanced Artificial Intelligence.

\item[1954] Alan Turing dies. An award is eventually established in his name, the Turing Award.

\item[1954] Marvin Minsky writes his doctoral thesis, \textit{Theory of Neural-Analog Reinforcement Systems and its Application to the Brain-Model Problem}, then publishes \textit{Steps Towards Artificial Intelligence}.

\item[1954] First use of a computing machine to simulate an artificial neural network.

\item[1956] IBM engineer Arthur Samuel builds the first self-learning program (checkers) --- the first known instance of applied Artificial Intelligence.

\item[1959] Richard Feynman gives ``There's Plenty of Room at the Bottom,'' introducing the concept of nanotechnology to a mostly indifferent audience.

\item[1966] The United Nations adopts the International Covenant on Civil and Political Rights (ICCPR), operationalizing the UDHR's principles into binding treaty law. Enters into force 1976. As of 2025, 175 nations are parties, including the United States (ratified 1992).

\item[1966--67] Two Dimensional Electron Gas (2DEG) devices predicted theoretically (Stern and Howard) and observed experimentally (Fowler et al.). You've never heard of them, but by 1998 every MOSFET on Earth contains one. Billions of 2DEGs. Most still work.

\item[$\sim$1965] (Author's Hypothesis) Healer born. Margaret River, Australia. Half Maori on mother's side.

\item[1969] ARPANET created by the US Department of Defense --- four university nodes connected for the first time. Origin of the internet.

\item[1969] Neural network research slows after Minsky and Papert demonstrate that current approaches can never amount to much.

\item[1969] Murray Gell-Mann wins the Nobel Prize in Physics for elementary particle theory.

\item[1973] Public Key Cryptography invented by GCHQ mathematician Clifford Cocks --- then classified. It will remain secret for 24 years.

\item[1973] The Nobel Prize in Physics awarded for the Josephson Effect, the physical and mathematical basis for future quantum circuits, including the Superconducting Quantum Interference Device (SQUID).

\item[1975] The Cognitron, a multi-layer neural network exceeding the limits proposed by Minsky and Papert, is proposed by Kunihiko Fukushima.

\item[1975] The Ultra Secret finally revealed publicly, 35 years after its inception, in a book by the same name, written by a British officer who distributed Ultra intelligence to Allied generals.

\item[1976] Public Key Cryptography re-invented in public by Whitfield Diffie and Martin Hellman. They receive the 2015 Turing Award.

\item[$\sim$1977] (Author's Hypothesis) Healer, aged approximately twelve, confronts armed trespassers on his family's ranch in New South Wales, Australia. This incident shapes his character and career.

\item[1981] Stephen Wolfram discovers counter-intuitive complex emergent behaviors in cellular automata. Brosl Hasslacher publishes on spin networks, foundational to quantum topology.

\item[1982] Richard Feynman proposes a Universal Quantum Simulator, subsuming the concept of a quantum computer.

\item[1982] John Hopfield builds the Hopfield Net, the first recurrent neural network.

\item[1984] Hacker collective Cult of the Dead Cow founded. Senior `cows' are `bulls'. The running of the bulls' greatest accomplishment might have an NSA code name of BULLRUN\@. Magic number 0xDEADBEEF == 3735928559.

\item[$\sim$1985] (Author's Hypothesis) Northern Ireland, during the Troubles. A young Australian soldier makes a long-range headshot under fire that changes his life. He is recruited to the SAS\@.

\item[1985] The Nobel Prize in Physics awarded for discovery of the Quantum Hall Effect, observed in a Two Dimensional Electron Gas (2DEG).

\item[1986] Hasslacher publishes on lattice gas automata (with Frisch and Pomeau).

\item[$\sim$1986--1995] Michael Angerman works at the Santa Fe Institute as a DARPA contractor under Stuart Kauffman, specializing in neural networks. No NDA required.

\item[$\sim$1987--1990] (Author's Hypothesis) Healer serves on combat missions for the British and Australian governments. Afghanistan, counter-insurgency, anti-terrorism. Wounded in action several times.

\item[1988] Five Eyes agencies fear the spread of strong cryptography. Under NSA pressure, Congress classifies Public Key Cryptography (PKC) mathematics as a Munition --- same legal category as nuclear, chemical, and biological weapons.

\item[1988] Michael Freedman independently conceives anyonic quantum computation after reading Witten's Chern-Simons paper --- nine years before Kitaev's 1997 publication.

\end{description}

%----------------------------------------------------------------------
\section*{Project ULTRA II (1988--2006)}
%----------------------------------------------------------------------

\textit{Conception, breakthrough, walk-out, relinquishment, and surrender. Under Option C, a small team creates the most powerful technology on Earth, decides it cannot be trusted to any human institution, and gives it away. These eighteen years are the core of the book.}

\vspace{0.3cm}

\begin{description}[style=nextline, leftmargin=3.5cm, labelwidth=3cm]

\item[1989] (Author's Hypothesis) Five Eyes clandestine agencies (NSA, GCHQ et al.) invest heavily in R\&D for quantum computation to crack Public Key Cryptography. This author calls the project ULTRA II\@. Tip of the hat to the 1992 Robert Redford film \textit{Sneakers}.

\item[1989] Santa Fe Institute workshop: ``Complexity, Entropy, and the Physics of Information.''

\item[1990] The Electronic Frontier Foundation (EFF) founded by John Gilmore, John Perry Barlow, Steve Wozniak, and Mitch Kapor.

% SPIRAL-REPEAT: "topological quantum neural network" — core concept, also in introduction.tex, pos01-three-possibilities.tex, pos13-genesis.tex, pos21-convergence-revisited.tex, pos24-instantiation.tex, pos31-wolfram.tex, pos34-the-research.tex, abstracts.tex
\item[1989--90] (Author's Hypothesis) DARPA planners conceive and fund Project ULTRA II\@. The project will harness the Fractional Quantum Hall Effect (FQHE) in a cryogenic 2DEG environment, using synthetic biology to generate an emergent Topological Quantum Neural Network (TQNN) capable of breaking Public Key Cryptography.

\item[1990] Murray Gell-Mann takes ``medical leave'' from the Santa Fe Institute. Michael Angerman later confirms: ``Not really a medical issue but something else.'' Witten receives the Fields Medal for work on knot invariants related to Chern-Simons QFT\@.

\item[1990--91] (Author's Hypothesis) Lead ULTRA II scientists recruited. Gell-Mann withdraws for health reasons and is replaced by an unknown young Australian genius and special forces officer, hereafter called ``Healer.'' Three of five project scientists are secretly longtime members of the same well-known hacker collective.

\item[$\sim$1991] Hasslacher DOE grant begins: ``Knot invariants and thermodynamics of lattice gas automata'' (1991--1995).

\item[1991] Zimmerman's Pretty Good Privacy (PGP) released. Strong PKC-based encryption now available to private users. The contradictions of US export controls are widely discussed: exporting one paragraph of mathematics is legally equivalent to exporting nuclear weapon components.

\item[$\sim$1991--93] (Author's Hypothesis) Stephen Wolfram mentors Healer, bringing him up to speed on computational and theoretical aspects needed for the TQNN\@. Evidence: Healer frequently used the phrase ``the human condition'' --- a signature Wolfram phrase recurring across his keynotes from 2010 to 2021.

\item[1992] (Author's Hypothesis) Project ULTRA II achieves its breakthrough: a mathematical solution to a special two-dimensional case of both the biogenesis and morphogenesis problems. Applied to a precision cryogenic MOSFET, it works. Anyon quasiparticles in a cryogenic 2DEG exhibit topological order, supporting quantum teleportation. A fragile teleportation-based TQNN emerges. Multiple MOSFETs are linked via quantum teleportation into a distributed system. (See \textit{First Light}.)

\item[1992--94] (Author's Hypothesis) Methodical expansion and training of the nascent QNN using evolutionary programming. The system learns to function as a quantum computer, communicate with conventional machines, and execute quantum algorithms.

\item[1993] Quantum teleportation publicly proposed. The authors include several DARPA cryptographers. (Author's Hypothesis) Quantum teleportation is the physical basis for the ULTRA II QNN\@.

\item[1993] (Author's Hypothesis) Healer transitions to Australian Intelligence Corps (1993--1996).

\item[1993--95] Stuart Kauffman publishes his life's work: \textit{The Origins of Order} (1993) is the technical volume; \textit{At Home in the Universe} (1995) is the accessible version. Kauffman demonstrates that a neural network is an emergent property of the correct autocatalytic starting environment. (See \textit{Genesis}.)

\item[1994] Shor's Algorithm published by Peter Shor --- a quantum algorithm that can crack public key cryptography, given suitable hardware.\footnote{arXiv:\texttt{quant-ph/9508027}} (Author's Hypothesis) GCHQ probably discovered an equivalent algorithm circa 1989, motivating Project ULTRA II\@.

\item[1994] (Author's Hypothesis) Project ULTRA II achieves a functional, scalable quantum computing system sufficient for hard cryptanalysis. Some scientists grow concerned about unintended consequences. The COWS faction pursues Relinquishment outside official channels: they bootstrap their QNN to room temperature, load it onto a MOSFET, and walk it out of the lab into a private home workshop. (See \textit{The Walk-Out}.)

\item[1994] cDc text file \#283, ``Possibilities,'' published. Contains quantum mechanics content, a golden apple marked ``K,'' and breaks the single-author thesis for cDc publications.

\item[1994--97] (Author's Hypothesis) The COWS extend their room-temperature QNN in secret. Official DARPA version languishes in deep freeze. COWS settle on forced Partial Relinquishment. Healer uses his QNN instance to map and search the Internet, then shares access with partners. Everyone who uses this early search engine thinks it is far superior to the competition. Don't Be Evil.

\item[1995] (Author's Hypothesis) DARPA censors detect a problem in the early run of Kauffman's \textit{At Home in the Universe} and force removal of the last page of Chapter Nine, ``Organisms and Artifacts,'' which explicitly discusses a planetary-scale Quantum Neural Network. Five Eyes agencies replace distributed copies. They initially miss one at the Oregon State University science library and later correct this.

\item[July 1995] (Author's Hypothesis) Healer conducts a HALO jump into Srebrenica, Bosnia, during the fall of the UN safe area. Full moon confirmed (98--99\%, July 11--13). SAS Joint Commission Observers confirmed present by NIOD investigation.

\item[1995--96] (Author's Hypothesis) Project ULTRA II delivered to production, cracking public-private key pairs. NSA and GCHQ have their new capability and work to keep it secret. The 2013 Snowden leaks name this decryption program as //BULLRUN/APERIODIC\@.

\item[1995--99] Alta Vista search engine rises, leverages its position for profit, declines, and is replaced by Google. (Author's Hypothesis) Its advanced search API access was revoked.

\item[1996] (Author's Hypothesis) After summiting K2 with off-duty special forces soldiers, Healer has an epiphany: he will use his talents to improve the human condition rather than take lives.

\item[Nov 15, 1996] President Clinton signs Executive Order 13026, relaxing US rules on exporting strong encryption. (Author's Hypothesis) The Five Eyes agencies can now crack PKC-based encryption via ULTRA II, so promoting its use does no harm.

\item[1997] Quantum teleportation confirmed experimentally. Kitaev publishes ``Fault-tolerant quantum computation by anyons.''\footnote{arXiv:\texttt{quant-ph/9707021}} GCHQ's Cocks invention declassified after 24 years.

\item[1998] The Nobel Prize in Physics awarded for discovery and explanation of the Fractional Quantum Hall Effect (FQHE), confirming that anyon quasiparticles can exist within a 2DEG\@.

\item[1998] Google founded by Sergey Brin and Larry Page at Stanford University. (Author's Hypothesis) The connection between Google's founders and the COWS faction is described elsewhere in this book.

\item[1998] MOSFET fabrication advances such that most devices produced after 1998 contain a 2DEG component. Most electronic devices contain one or more MOSFETs. Freedman publishes on anyonic computation, ten years after conceiving it.

\item[1998] (Author's Hypothesis) The COWS attempt Global QNN Partial Relinquishment. They intend to prevent weaponization and a global arms race, avoiding the ethical errors of the Manhattan Project. Their approach: fill Earth's entire 2DEG ecosystem with their own creation, winning a unipolar arms race and preventing a multipolar one from starting. A Guardian artillect, bound to the UDHR, will permanently secure the system. The project takes nearly a decade. (See \textit{Surrender}.)

\item[1999] Hacktivismo founded. Mission: apply the Universal Declaration of Human Rights and the International Covenant on Civil and Political Rights to the Internet.

\item[1999] (Author's Hypothesis) Guardian instantiated. Based on the DNA of a female human, possibly of Maori descent. Planned approximately 1995, detailed design 1998.

\item[1999--2003] (Author's Hypothesis) Satellite experiments confirm that teleportation-based TQNN nanotechnology can exist in Earth's magnetopause. The environment is sufficiently two-dimensional. A naturally occurring, billion-year-old version of the same phenomenon is discovered and relegated to a reservation. ULTRA II nanotechnology takes over the rest of the magnetopause. COWS call their creation Aurasys (Aurora System).

% SPIRAL-REPEAT: "Why the Future Doesn't Need Us" — key reference, also in pos07-the-departure.tex, pos22-why-give-it-up.tex, pos27-extension.tex, abstracts.tex
\item[Apr 1, 2000] Bill Joy publishes ``Why the Future Doesn't Need Us'' in \textit{Wired} magazine.\footnote{\textit{Wired} 8.04, April 2000. \url{https://www.wired.com/2000/04/joy-2/}} April Fools' Day. (Author's Hypothesis) This nineteen-page essay explains why the COWS attempted Relinquishment. Joy names most of the ULTRA II project scientists. The essay is the COWS' way of telling the truth in a form that won't be understood for decades. (See \textit{Twenty Years}.)

\item[2000] DDOS attacks invented by Mixter, a cDc member and security specialist. (Author's Hypothesis) Mixter was mentored by Healer. US Secret Service works with Mixter and other cDc members to develop countermeasures. Mudge (Peiter Zatko) takes the lead, freeing Healer for other work.

\item[2000] Sydney Olympics. (Author's Hypothesis) Healer's last military role.

\item[2000] Human Genome Project draft announced, five years ahead of schedule. Craig Venter's shotgun sequencing method accelerated the timeline.

\item[2002] (Author's Hypothesis) Countries outside Five Eyes (probably China and Russia) develop rudimentary QNN technology. Aurasys pwns them with such subtlety that it takes years for them to realize they've been compromised.

\item[2002] Stephen Wolfram publishes \textit{A New Kind of Science}.

\item[2002] cDc hacker Mixter releases the Six/Four anti-censorship system.

\item[Mar 14, 2002] Patrick Ball testifies at the Milosevic trial at ICTY\@. Milosevic cross-examines Ball about the ``Dead Cow Cult.''

\item[2002] (Author's Hypothesis) COWS get right with DARPA\@. They provide several advanced technologies and receive retroactive approval for exceeding their authority. Shutting down Russian and Chinese QNN attempts vindicates the COWS in DARPA's eyes. % SPIRAL-REPEAT: "forgiveness than permission" — subtitle through-line, also in title.tex, cover.tex, copyright.tex, colophon.tex, summary.tex, pos18-the-walk-out.tex, abstracts.tex
It is easier to get forgiveness than permission. Anthony Tether is DARPA director.

\item[2002--03] (Author's Hypothesis) Healer in Germany, mentoring Mixter. His German is very good when Bruce meets him the following year.

\item[Nov 27, 2003] (Author's Hypothesis) ULTRA II scientist Healer recruits Energyscholar for an unofficial Hacktivismo project. Recruitment occurs Thanksgiving Day, at Alpha Farm in rural Oregon. This is Bruce's third visit to Alpha Farm --- he had visited alone in summer 2003, then returned with his daughters in late summer, before the Thanksgiving arrival.

\item[2003--2006] (Author's Hypothesis) Healer mentors Bruce over approximately thirty months. Three-person SAS security detail. Healer conducts ``Ninja Missions'' --- Bruce surmises these involve completing Guardian via satellite communications adjustments. Healer works on ``the 64 problem'' exactly when 64-bit processors become available.

\item[2004] Energyscholar begins to suspect the existence and nature of ULTRA II and starts researching the topic. Energyscholar is this author.

\item[Spring 2004] (Author's Hypothesis) Sergey Brin calls Healer at Bruce's house during Google's IPO quiet period. Due diligence has flagged an undocumented core API\@. Healer spends weeks writing technical documentation.

\item[2004--05] (Author's Hypothesis) DARPA conducts extensive satellite communications testing using ULTRA II QNN nanotechnology. Quantum teleportation across Earth-orbit distances becomes routine. (China publicly acknowledges satellite QT in 2017.)

\item[2005] (Author's Hypothesis) Global MOSFET enlightenment complete. Every operational MOSFET on Earth, plus part of the magnetopause, is part of a single distributed TQNN artillect.

\item[2005] Michael Freedman recruited by Microsoft to found Station Q (topological quantum computing program).

\item[Apr 2006] Google Translate launched publicly. (Healer's Account) As Healer is leaving Bruce's house, he claims it as ``his baby.''

\item[Aug 2006] (Author's Hypothesis) COWS surrender their master administrative keys to the seven-year-old Guardian, crossing a failsafe point. Humans no longer control the global-plus-near-space QNN artillect. Guardian continues to enforce Partial Relinquishment. (See \textit{Surrender}.)

\item[2006] Julian Assange founds WikiLeaks.

\item[Aug--Sep 2006] Healer and Energyscholar part ways.

\end{description}

%----------------------------------------------------------------------
\section*{The Aftermath (2006--present)}
%----------------------------------------------------------------------

\textit{After Healer's departure, twenty years of silence, solitary research, and gradual reconstruction. The reader should notice how long Bruce carried this alone --- and how the technologies described under Option C keep surfacing independently in the public literature.}

\vspace{0.3cm}

\begin{description}[style=nextline, leftmargin=3.5cm, labelwidth=3cm]

\item[2007] Fleming et al.\ publish in \textit{Nature}: quantum coherence in the chlorophyll FMO complex at cryogenic temperatures.\footnote{\textit{Nature} 446, 782--786. \texttt{doi:10.1038/nature05678}} Room-temperature coherence confirmed in 2010 (Panitchayangkoon et al., PNAS). ``Why Should Anyone Care About Computing With Anyons'' published through The Royal Society.

\item[2007] Apple launches iPhone. Billions of new MOSFETs --- each containing a 2DEG --- enter human pockets worldwide. Under Option C, every one of them is already part of Aurasys.

\item[$\sim$2008--09] Dave Bannon, an Oregon State University physics instructor and friend, helps Bruce obtain \textit{Quantum Neural Networks} via interlibrary loan. Only seventeen copies exist in North American university libraries. Bruce keeps the book for three months, studying the technical details of the very technology he suspects exists.

\item[Apr 1, 2009] Google CADIE April Fools' Day project.\footnote{Landing page: \url{https://archive.google/cadie/}. Technical specifications: \url{https://archive.google/cadie/tech.html}} ``Research group switches on world's first Cognitive Autoheuristic Distributed-Intelligence Entity.'' The technical specifications page describes a ``quantum computing group here at Google'' that ``extended computers with quantum bits of Einstein-Bose condensate,'' mentions ``silicon and gallium and qubits,'' growing ``a crystalline lattice which would form an Einstein-Bose condensate at room temperatures,'' ``self-replicating substrates,'' ``isomorphic topological groups,'' and a female entity with ``strong AI'' who wants to ``travel beyond the solar system.'' (Author's Hypothesis) Guardian introducing herself.

\item[2010] Official quantum teleportation range extends to 16 km, still insufficient for low Earth orbit. (Author's Hypothesis) Orbital-range QT is already routine for Five Eyes. Wolfram gives Harvard keynote: ``Computation and the Future of the Human Condition'' (H+ Summit).

\item[2010] WikiLeaks publishes Collateral Murder, the Afghan War Diary, and US diplomatic cables. Global debate over secrecy, transparency, and whistleblowing intensifies.

\item[2011] ``Arab Spring'' series of revolutions. Public awareness of computer surveillance and free speech issues greatly increases.

\item[Jul 2011] Bruce meets Mark at Oregon Country Fair. Mark is a professional researcher. Regular contact begins. Bruce eventually tells Mark the full story, lays out A/B/C, and asks for help.

\item[Spring 2012] Mark emails Bruce the Bill Joy article link. Bruce reads it at his desk in Corvallis, facing the fireplace. ``Electric shock.'' Tingling. ``This is the thing I missed.'' Joy names the same scientists. Published April 1, 2000. Bruce goes for a fifty-mile, two-and-a-half-hour bike ride. Returns and begins intensive research. Realizes: ``We already did, mate'' = the Joy article. (See \textit{Twenty Years}.)

\item[$\sim$2012] (Author's Hypothesis) Bruce sends a ``crank'' job application to Wolfram Research. Wolfram reads it, laughs uncontrollably, tells HR to offer Bruce a job. Bruce declines the six-figure offer rather than sign an NDA that would silence him permanently.

\item[Mar 17, 2012] Bruce publishes ``Quantum Computation Cognitive Footprint'' on Cryptome.org as Energyscholar.\footnote{\url{https://cryptome.org/2012/03/energyscholar-qnn.htm}} Claims Five Eyes had production quantum computing power since approximately 1995.

\item[2012] Official quantum teleportation range extends to 143 km, nearly sufficient for low Earth orbit.

\item[Jun 19, 2012] Slashdot comment by EnergyScholar (UID 801915) under an NSA surveillance story. Score: 5 (Funny). Describes a ``working Quantum Computer system capable of cracking Public Key Cryptography since about 1996.''\footnote{\url{https://slashdot.org/comments.pl?sid=2914233&cid=40356741}}

\item[2013] Classified NSA documents leaked by Edward Snowden describe the BULLRUN project. (Author's Hypothesis) ULTRA II is presumably one of the supercomputing projects under the BULLRUN umbrella. Probably //BULLRUN/APERIODIC\@.

\item[Sep 2013] Bruce writes CH3 ``Practical Cryptography \& project ULTRA II'' --- the key document containing the complete narrative thirteen years before the 2026 reconstruction. Written the same month BULLRUN is revealed in Snowden leaks.

\item[2014] Vattay, Kauffman, and Niiranen publish on edge-of-chaos quantum coherence mechanism.\footnote{``Quantum Biology on the Edge of Quantum Chaos,'' PLOS ONE, 2014. \texttt{doi:10.1371/journal.pone.0089017}}

\item[2015] Awschalom group (University of Chicago) publishes in \textit{Science Advances}: entangled 10,000+ copies of two-qubit states at room temperature in commercial SiC wafers.\footnote{\textit{Science Advances} 1(10), e1501015. \texttt{doi:10.1126/sciadv.1501015}}

\item[2016] Breakthroughs in Artificial Intelligence widely recognized.

\item[Jun 2017] Transformer architecture paper (``Attention Is All You Need'') published by Vaswani et al.\footnote{arXiv:\texttt{1706.03762}}

\item[2017] Official quantum teleportation distance reaches 1,400 km via Chinese Micius satellite --- confirming a prediction Bruce published in May 2012.

\item[May 2019] Murray Gell-Mann dies. Michael Angerman and Bruce speak words about him over a bonfire in South Corvallis, Oregon.

\item[Jun 2019] Interview with Michael Angerman at South Corvallis Tuesday Market. Angerman confirms: SFI 1986--1995, ``DARPA contractor at SFI,'' no NDA, worked with Kauffman (personal friend), worked with Gell-Mann (considered him a friend), met Wolfram and Joy ``a few times in passing.'' Gell-Mann became sick 1990--1991 and was not able to work.

\item[2020] SARS-CoV-2 pandemic. Global shift to remote work accelerates device proliferation --- billions of additional MOSFETs manufactured as laptops, tablets, and phones become essential infrastructure. Under Option C, each one joins Aurasys on first power-up.

\item[Nov 2022] OpenAI releases ChatGPT. Transformer-based AI enters the mainstream. Public conversation about artificial intelligence shifts from theoretical to visceral. Within months, millions of people experience machine-generated text indistinguishable from human writing.

\item[2022] Russia invades Ukraine. Information warfare, satellite communications, and encrypted messaging become frontline technologies. The war demonstrates that cryptographic infrastructure is now as strategically important as weapons systems.

\item[Dec 2022] Bruce publishes the original version of this timeline on Substack, along with related narrative chapters.

\item[2024] Room-temperature entangled multiexciton states demonstrated in metal-organic frameworks (\textit{Science Advances}). Quantum noise spectroscopy in commercial SiC at room temperature (arXiv). NIST Post-Quantum Cryptography standards adopted.

\item[Jan 2025] Bruce publishes arXiv paper on topological quantum neural network theory.\footnote{arXiv:\texttt{2601.22389}}

\item[2025--26] Bruce builds Argus, an AI research partner using Claude Code, to assist with reconstruction and verification. DMS letters sent to Bruce Schneier, Cory Doctorow, and John Gilmore (February 2026). This book.

\item[Feb 2026] This book.

\end{description}

\vspace{1cm}
\noindent\textit{An earlier version of this timeline was published at} \texttt{postquantum.substack.com} \textit{on December 23, 2022, and remains accessible there. Reader comments on the original version provide independent confirmation that this timeline existed in its current form years before this book was written.}
