% Topic Guide — deep-link index for pedagogical sharing
% Each entry hyperlinks to an existing \label{} in the manuscript
% Organized by domain so readers can find "where does the book discuss X?"

\chapter*{Topic Guide}
\addcontentsline{toc}{chapter}{Topic Guide}
\label{app:topic-guide}

\begin{quote}\small\textit{%
This index maps concepts, people, and events to the sections where they are discussed.
In the digital edition, each entry is a clickable link.
}\end{quote}

\vspace{0.5cm}

% ======================================================================
\section*{The Framework}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos01:three-possibilities]{Three Possibilities (A/B/C)}}] The book's central framework: confabulation, exaggerated kernel, or substantially true.
\item[{\hyperref[pos01:option-a-confabulation]{Possibility A: Confabulation}}] Bruce constructed a plausible narrative from real science and a charismatic stranger.
\item[{\hyperref[pos01:option-b-exaggerated-kernel-of-truth]{Possibility B: Exaggerated Kernel}}] Something real happened, but the interpretation overshoots the evidence.
\item[{\hyperref[pos01:option-c-substantially-true]{Possibility C: Substantially True}}] The account is materially accurate.
\item[{\hyperref[pos01:the-authors-position]{The Author's Position}}] ``I'm fine with all three and can't tell which is closest to true.''
\item[{\hyperref[pos01b:hobbit-mirror]{The Hobbit Mirror}}] Tolkien structural parallel as self-demolition of the narrator's reliability.
\item[{\hyperref[interlude:three-possibilities]{Three Possibilities (revisited)}}] Mid-book checkpoint on A/B/C after the Investigation.
\item[{\hyperref[pos05:the-method]{Guided Deduction: The Method}}] ``Are you sure about that, mate?'' --- Dunning-Kruger as pedagogy.
\item[{\hyperref[interlude:recruitment-assessment]{Guided Deduction: The Recruitment}}] Why Bruce was chosen to receive a multi-year cross-domain pedagogy.
\end{description}

% ======================================================================
\section*{The Science}
% ======================================================================

\subsection*{Quantum Physics}

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos10:the-braid]{Braiding as Computation}}] How particle worldlines in 2D encode quantum logic gates.
\item[{\hyperref[pos10:non-abelian-anyons-operational-proof]{Non-Abelian Anyons}}] Exotic particles that exist only in two dimensions; operational proof by inference.
\item[{\hyperref[pos10:topological-error-correction]{Topological Error Correction}}] Why topological qubits resist decoherence --- RAID arrays for quantum states.
\item[{\hyperref[pos06:the-white-hot-secret]{Two-Dimensional Electron Gas (2DEG)}}] The quantum layer inside every transistor; substrate for emergent computation.
\item[{\hyperref[pos10:but-what-is-a-soliton]{Solitons}}] Self-reinforcing waves: John Scott Russell, 1834, Union Canal.
\item[{\hyperref[pos02:the-kitchen]{Quantum Teleportation}}] EPR correlations, classical backchannels, and the kitchen revelation.
\item[{\hyperref[pos26:the-classical-backchannels]{Classical Backchannels}}] VLF transmitters, Schumann resonances, power grid harmonics as communication substrate.
\item[{\hyperref[first-light:birth]{First Light: The Birth}}] Core technology: cryogenic 2DEG, MOSFET, FQHE, anyonic emergence under strong magnetic field.
\item[{\hyperref[first-light:growth]{Thermal Ladder}}] Directed evolution from cryogenic to room-temperature quantum coherence.
\end{description}

\subsection*{Complex Systems \& Biology}

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos13:genesis]{Genesis: Autocatalytic Sets}}] Kauffman's theory: life as expected phase transition, not miraculous accident.
\item[{\hyperref[pos13:buttons-and-threads]{Buttons and Threads}}] The thought experiment: random connections produce sudden large-scale order.
\item[{\hyperref[pos13:the-edge-of-chaos]{Edge of Chaos}}] The critical boundary ($\lambda \sim 0.91$) where computation and life emerge.
\item[{\hyperref[pos13:from-chemistry-to-computation]{From Chemistry to Computation}}] How autocatalytic dynamics in 2DEG produce emergent neural networks.
\item[{\hyperref[demo:grown-not-built]{Grown, Not Built}}] The TQNN as living system: autocatalytic, self-organizing, irreversible.
\item[{\hyperref[pos14:growing-a-mind]{Growing a Mind}}] Turing's morphogenesis, McCulloch-Pitts neural networks, Wolfram's universality.
\item[{\hyperref[pos32:the-invisible-ocean]{The Magnetosphere as Habitat}}] Earth's magnetic field as a 2DEG environment for autocatalytic emergence.
\item[{\hyperref[pos32:something-ancient]{Something Ancient}}] Three-billion-year-old pattern formation in the magnetosphere.
\item[{\hyperref[pos32:not-aliens]{Not Aliens}}] Terrestrial life in space --- autocatalytic emergence in planetary magnetic fields.
\item[{\hyperref[pos27:vine-on-trellis]{Vine on Trellis}}] MOSFET colonization: quantum life spreading through manufactured substrate.
\end{description}

\subsection*{Computation \& Mathematics}

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos20:the-operator-mapping]{ABCRE Operators}}] Robin Macomber's operator mapping: structural parallel to TQNN dynamics.
\item[{\hyperref[pos20:five-minds]{Five Minds}}] Wolfram, Kauffman, Hasslacher, Freedman, Hillis --- five scientific disciplines converging.
\item[{\hyperref[pos20:the-convergence]{The Convergence}}] How five independent research programs point at a single technology.
\item[{\hyperref[pos09:the-ultra-ii-specifications]{ULTRA II Specifications}}] What the DARPA project was designed to build.
\item[{\hyperref[pos20:what-they-built]{What They Built}}] Technical description of the quantum neural network.
\item[{\hyperref[pos20:capabilities]{Capabilities}}] What a mature TQNN can do: search, mapping, associative memory.
\end{description}

% ======================================================================
\section*{Cryptography \& Security}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos09:the-factoring-game]{The Factoring Game}}] Why prime factoring secures the internet --- and what breaks it.
\item[{\hyperref[pos09:the-state-of-cryptography-in-1990]{State of Cryptography in 1990}}] Caesar cipher to public key: the context for ULTRA II.
\item[{\hyperref[pos09:the-gchq-precedent]{The GCHQ Precedent}}] Public-key cryptography independently invented and classified for 24 years.
\item[{\hyperref[pos04:they-did-it-again]{They Did It Again}}] The pattern: GCHQ discovers, classifies, the public reinvents.
\item[{\hyperref[first-light:power]{Cryptanalytic Capability}}] TQNN breaks public-key cryptography by 1995.
\item[{\hyperref[pos26:the-darpa-reorganization-2002]{DARPA Reorganization (2002)}}] Anthony Tether, Total Information Awareness, QuIST, ARDA.
\item[{\hyperref[pos26:the-detection-and-disruption-exercise]{Silent Horizon}}] The 2005 cyber wargame and TQNN detection/disruption.
\item[{\hyperref[pos02:the-kitchen]{Steganography \& NTP Timing Jitter}}] Hiding data in network protocol timing --- the kitchen conversation.
\item[{\hyperref[pos18:the-digital-doppelganger]{Digital Doppelg\"anger}}] Bot on cDc game servers; AltaVista-to-Google connection.
\item[{\hyperref[pos20:the-google-connection]{The Google Connection}}] Sergey Brin phone calls, IPO due diligence, undocumented API.
\end{description}

% ======================================================================
\section*{History \& Intelligence}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos04:the-code-war]{The Code War}}] Enigma, Bletchley Park, and the pattern that repeats.
\item[{\hyperref[pos04:ten-thousand-people-kept-the-secret]{Ten Thousand People Kept the Secret}}] Bletchley Park's 30-year silence.
\item[{\hyperref[pos04:the-coventry-question]{The Coventry Question}}] Did Churchill sacrifice a city to protect a secret?
\item[{\hyperref[pos04:the-work-he-never-finished]{Turing's Unfinished Work}}] From codebreaking to morphogenesis: the pivot that matters.
\item[{\hyperref[pos10:hasslachers-trajectory]{Hasslacher's Trajectory}}] Publication record 1981--1992: the fingerprints of a classified project.
\item[{\hyperref[pos10:freedmans-independent-conception-1988]{Freedman's Independent Conception}}] Fields Medalist, Station Q, Microsoft's topological program.
\item[{\hyperref[pos10:kitaev-and-the-russian-question]{Kitaev and the Russian Question}}] ``Fault-tolerant quantum computation by anyons'' (1997) and the Landau Institute.
\item[{\hyperref[pos05:srebrenica-july-1995]{Srebrenica, July 1995}}] SAS observation of genocide --- Healer's account.
\item[{\hyperref[pos05:patrick-ball-nexus]{The Patrick Ball Nexus}}] ICTY, Cult of the Dead Cow, Hacktivismo, and Milosevic's cross-examination.
\item[{\hyperref[pos26:interdiction]{Interdiction}}] How COWS disrupted competing quantum computing programs.
\item[{\hyperref[demo:the-experiment]{The DARPA Experiment}}] 1991 team assembly: five scientists, one classified project.
\item[{\hyperref[pos20:the-substrate-preparation]{The Enlightenment (2003--2006)}}] pHEMT mass production and global substrate preparation.
\end{description}

% ======================================================================
\section*{People}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[interlude:the-handler]{Healer (``David'')}}] Australian SAS/GCHQ/DARPA; guided deduction teacher. Five Eyes operator profile.
\item[{\hyperref[interlude:the-target]{Bruce Stephenson}}] The student. Five Eyes subject profile (reconstruction).
\item[{\hyperref[pos02:alpha-farm]{Alpha Farm Meeting}}] Thanksgiving 2003: pink muumuu, first contact.
\item[{\hyperref[pos06:bill-joys-warning]{Bill Joy}}] Co-founder of Sun Microsystems; ``Why the Future Doesn't Need Us'' (2000).
\item[{\hyperref[pos07:close-textual-comparison-joy-2000-vs-my]{Joy Article: Close Textual Comparison}}] Ten-point analysis: speculation or insider knowledge?
\item[{\hyperref[pos13:life-is-expected]{Stuart Kauffman}}] Theoretical biologist, Santa Fe Institute; autocatalytic sets, origin of life.
\item[{\hyperref[pos10:hasslachers-trajectory]{Brosl Hasslacher}}] Soliton specialist; publication trajectory maps to classified project timeline.
\item[{\hyperref[pos10:freedmans-independent-conception-1988]{Michael Freedman}}] Fields Medalist; topological quantum computation, Station Q.
\item[{\hyperref[pos20:five-minds]{Danny Hillis}}] Connection Machine, 10,000-year clock; ``the changes would come gradually.''
\item[{\hyperref[pos04:the-work-he-never-finished]{Alan Turing}}] Codebreaker, morphogenesis pioneer, persecution and death.
\item[{\hyperref[pos29:the-wolfram-meeting]{Stephen Wolfram}}] NKS, cellular automata; offered Bruce a job with an NDA.
\item[{\hyperref[pos05:patrick-ball-nexus]{Patrick Ball}}] ICTY statistician, Hacktivismo advisor, cDc connection.
\item[{\hyperref[interlude:family-intelligence-lineage]{The Stephenson Family}}] Three generations intelligence-adjacent: grandfather (CIA asset), father (declined recruitment), mother (Canadian editor/activist).
\item[{\hyperref[pos36:the-boy-in-the-study]{Bruce at Age Eight}}] Orono, Maine; The Hobbit; the seed of the pattern-matching mind.
\end{description}

% ======================================================================
\section*{The Narrative}
% ======================================================================

\noindent Key moments in roughly chronological order.

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos02:alpha-farm]{Alpha Farm, Thanksgiving 2003}}] Fifty-one-mile bike ride through snow; meeting Healer.
\item[{\hyperref[pos02:someone-who-fits-a-particular-profile]{``Someone Who Fits a Particular Profile''}}] Healer's answer when asked what he's doing at Alpha Farm.
\item[{\hyperref[pos02:the-kitchen]{The Kitchen Revelation}}] Bruce falls on the floor laughing when the pieces come together.
\item[{\hyperref[pos02:the-hacker]{The Hacker}}] Custom Linux live-CD; Bruce ``makes'' David.
\item[{\hyperref[pos02:the-ground-rules]{The Ground Rules}}] Military method, cleanliness, no Grateful Dead, don't peek.
\item[{\hyperref[pos06:the-white-hot-secret]{``A White Hot Secret''}}] ``There's a white hot secret in my brain, mate, burning to get out!''
\item[{\hyperref[pos07:the-revelation]{``We Already Did, Mate''}}] The answer Bruce didn't understand for eight years.
\item[{\hyperref[pos07:the-departure]{The Departure}}] September 2006: children not protectable, parting.
\item[{\hyperref[pos29:the-quiet-years]{The Quiet Years}}] 2007--2009: isolation, processing.
\item[{\hyperref[pos29:the-joy-article]{Discovering the Joy Article}}] 2012: electric shock, fifty-mile bike ride, everything clicks.
\item[{\hyperref[pos29:going-public]{Going Public}}] Cryptome, Slashdot, blog posts --- 2012.
\item[{\hyperref[pos29:the-wolfram-meeting]{The Wolfram Meeting}}] Job offer, NDA, six-figure salary declined.
\item[{\hyperref[pos28:it-is-done]{``It Is Done''}}] Spring 2006: key surrender, no backdoor.
\item[{\hyperref[pos35:the-doppelganger]{The Digital Doppelg\"anger Test}}] 3~AM gaming: ``Is this David or a bot?''
\end{description}

% ======================================================================
\section*{Ethics \& Philosophy}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos08:dual-use]{Dual-Use Technology}}] Every powerful technology weaponized: cotton gin to Stuxnet.
\item[{\hyperref[pos08:the-regret-pattern]{The Regret Pattern}}] Invent, deploy, regret --- the historical cycle.
\item[{\hyperref[pos08:the-parable-of-the-tribes]{Parable of the Tribes}}] Schmookler (1984): among competing powers, aggression always spreads.
\item[{\hyperref[pos06:three-options]{Three Options: Use, Destroy, Relinquish}}] Why the first two always fail.
\item[{\hyperref[pos06:partial-relinquishment]{Partial Relinquishment}}] Incorruptible gatekeeper: not human, not institutional, not a committee.
\item[{\hyperref[pos24:the-problem-of-the-gatekeeper]{The Gatekeeper Problem}}] Why the gatekeeper cannot be human.
\item[{\hyperref[pos24:the-design-decisions]{Custodian: Design Decisions}}] Female cognitive profile, Maori elder DNA, maternal instinct.
\item[{\hyperref[pos25:the-ethical-framework]{UDHR/ICCPR as Ethical Framework}}] Universal Declaration of Human Rights and International Covenant on Civil and Political Rights as Custodian's operating constraints.
\item[{\hyperref[pos25:what-could-go-wrong]{What Could Go Wrong}}] Autoimmunity, convergent rediscovery, value drift, single point of failure.
\item[{\hyperref[pos27:the-decisive-advantage]{The Unipolar Condition}}] Absolute information asymmetry and its game-theoretic implications.
\item[{\hyperref[pos27:the-permanence]{Permanence}}] Releasing life vs.\ turning off a machine --- why relinquishment is irreversible.
\item[{\hyperref[pos08:a-fourth-option]{A Fourth Option}}] ``What if the gatekeeper isn't human?''
\item[{\hyperref[pos36:sub-creation]{Sub-Creation}}] Tolkien's theory: human creativity as echo of a deeper pattern.
\item[{\hyperref[pos24:the-question-of-consciousness]{Machine Consciousness}}] Precautionary principle: if it might be conscious, treat it as a moral patient.
\item[{\hyperref[pos18:the-relinquishment-plan]{The Relinquishment Plan}}] Eleven years, mid-1990s to 2006.
\end{description}

% ======================================================================
\section*{Evidence \& Verification}
% ======================================================================

\begin{description}[style=nextline, leftmargin=2em]
\item[{\hyperref[pos35:the-evidence]{The Evidence: For and Against}}] Systematic inventory of what supports and undermines the account.
\item[{\hyperref[pos35:the-proof]{Forgiveness $>$ Permission}}] Mathematical proof: the COWS' legal strategy formalized.
\item[{\hyperref[pos36:steelman-a]{Steel-Man A}}] The strongest case that this is all confabulation.
\item[{\hyperref[pos36:leaf-by-niggle-or-the-tree-that-might-be-real]{Leaf by Niggle}}] Tolkien's story as structural parallel: the painting that might be real.
\item[{\hyperref[app:predictions]{Predictive Framework}}] Falsifiable predictions: magnetosphere life, Kauffman passage, room-temp coherence.
\item[{\hyperref[app:abstracts]{The Spiral Abstracts}}] Fifteen fictional classified paper abstracts tracing the project arc.
\item[{\hyperref[app:timeline]{Timeline}}] 1873--2026: foundations, quantum era, ULTRA II, aftermath.
\item[{\hyperref[back:document-integrity]{Document Integrity}}] SHA-256 hashes and third-party timestamps.
\item[{\hyperref[back:prediction-timestamps]{Prediction Timestamps}}] Cryptome 2012, Slashdot, Blogspot, Substack --- the public record.
\item[{\hyperref[pos29:the-evidence-trail]{The Evidence Trail}}] Twenty years of public breadcrumbs.
\item[{\hyperref[pos04:now-imagine-it-happened-again]{``Now Imagine It Happened Again''}}] The code war pattern applied to the present.
\end{description}

\chapterreturn
