% Chapter: The Stories
% Track: 2
% Content: Kangaroos + The Method (V6) + Srebrenica Witness + Patrick Ball (merged pos19) + Epstein witness
% MERGED: pos05 + pos19 — Plan 0091

\settrack{tracktwo}
% VOICE: 1st-person personal (Bruce) — "I" not "Bruce". Personal/investigative voice.

\chapter{What Healer Said}
\label{ch:t2-stories}
\label{pos05:the-stories}

\begin{quote}\small\textit{%
``The three chief virtues of a programmer are: Laziness, Impatience and Hubris.''%
}\par\raggedleft --- Wall, \textit{Programming Perl} (1991)\end{quote}

\vspace{0.3cm}

\begin{quote}\small\textit{%
``We cultivated our land, but in a way different from the white man. We endeavored to live with the land; they seemed to live off it. I was taught to preserve, never to destroy.'' --- Tom Dystra, Australian Aboriginal Elder
}\end{quote}

\begin{quote}\small\textit{%
The following is the author's fictionalized retelling of a story he heard once, decades ago.
Details including dates may differ from other accounts in this book.
}\end{quote}

\vspace{0.5cm}

\srcnote{Merged pos05+pos19 | Plan 0091 | CH3, HEALER-RECONSTRUCTION.md, HealerStories archive | Kangaroos, The Method (V6), Srebrenica, Patrick Ball ICTY-cDc nexus, SAS corroboration}

The year is 1978.  The place is a large quarter horse ranch in New South Wales, Australia.

A twelve year old boy rides the fence line astride a four year old quarter horse mare.  It's a three day ride around the ranch.  A working ranch never has enough hands.  Securing the fence line is a man's job.  This boy has become a young man.  David rides the fence line with water, food, a knife, a first aid kit, basic field gear, a fence repair kit, and a .303 Lee Enfield rifle.  He knows how to operate a HAM radio but today he has no radio. Any problems along the fence line must be fixed or reported.

Along the way the young man will visit various people who live on the ranch.  Some are retired Australian military men and their families.  Most of the people he will visit are of Kuringai aboriginal ethnicity. Many Kuringai people live on this land as they have for thousands of years.  They maintain good relations with the ranch `owners', who protect them from other white Australians and allow them to live there in peace. In return the Kuringai perform certain maintenance and security responsibilities. These responsibilities include training the children of the ranch owner in aboriginal bushcraft.

David senses trouble before he rides up to the North Gate.  Subtle signs suggest something is amiss.  His friends Koori and Ngemba, both aboriginal adult men, show themselves as he approaches on horseback.  He rides up to them and dismounts.  Friendly greetings are brief. Trouble is in the air.

Ngemba and Koori confirm the boy's suspicions. Ngemba tells David that a ute, or pickup truck, loaded with gubbins, dangerous armed white men, have crashed the gate and are on the ranch making trouble.  They are drinking beer and shooting kangaroos with their rifles. Kangaroos are sacred to the Kuringai.  Killing kangaroos this way is illegal but rarely punished. Ngemba and Koori have already spread the alarm.

These gubbins are known.  They have trespassed on the ranch before.  Each time they leave behind empty bottles of liquor, bullet casings, and dozens of dead 'roos.  The Kuringai people weep at the waste but can do nothing.  The unruly white men have rifles so the Kuringai must avoid a direct confrontation.  None of the Kuringai will show themselves to the gubbins.   Whenever aborigines fight whites the abos are always punished or killed, regardless of who started it, who won, or who has the moral high ground or law on their side.  The white men in the ute will be long gone before the police can arrive.

As the son of the ranch owner David has much higher status with the police than do the aborigines.  If anyone is to do something it's up to him.

Koori tells David that the gubbins are parked about five kilometers away on the north side of Wattle Ridge. Ngemba's friend Darawal, who can effect a very convincing white man's accent, is already on his way to the nearest telephone to alert the police.  This land is big, though, and it will take police at least three hours to arrive.  None of the retired SASR veterans who live on the ranch happen to be nearby.

The police are on their way.   While David is not Kuringai, he and his twin sister grew up around them and learned their ways. Both he and his sister lived among them for months at a time.  Aborinal hunters taught them stealth, tracking, and bushcraft.  David is half Maori on his mother's side \& half European Australian on his father's side.  Most importantly, his parents own the ranch.  He has social and legal authority to defend the ranch against invaders in a way the Kuringai do not.  He can get away with things that would get aborigines arrested or killed.

With Koori and Ngemba's help David forms a daring plan. He will try to delay the gubbins until the police arrive.  Ngemba and Koori wish him luck.  They will stealthily observe from a safe distance.

David checks his gear and saddles up.  He stays off the road and circles towards the intruders, careful to keep hills between them.  Sporadic gunfire tells him where they are.  David has hunted before but this is his first time hunting men.  Their rifles make them very dangerous.  Their probable drunkenness makes them no less dangerous.

The boy knows this hilly land.  He walked it last year with Koori.  He rides up a rocky hill to a ridgeline.  The trespassers are on the other side.  He dismounts about 600 meters from the dangerous armed intruders, on the other side of a hill, and leads his horse. Pop! Pop! More gunfire as another kangaroo dies.   He feeds and waters his horse then leaves it hobbled and tied. Horse and boy are ready for a quick departure along a covered escape route.

David cautiously approaches the ridgeline.   The youth suppresses his fear and controls his breathing. He readies his binoculars and checks his rifle.  He crawls the last few meters to the ridgeline and peeks between big rocks towards the intruders below.

The dry valley far below is a scene of violence.  The ute is a big four seater, what most non- Australians would call a large four-by-four pickup truck.   Four men equipped with rifles and beer are shooting kangaroos.  About 20 'roos lie dead or dying.  Red blood still flows.  David watches through binoculars.  One man raises his rifle, shoots another 'roo, then swigs more beer.  David feels his blood rise.  He lets loose a quiet Kuringai curse at these awful men. He will not kill them but he is determined to delay them until the police arrive.

The boy has a clear line of fire down on the intruders and their vehicle.  They are about 400 meters away and still unaware of his presence.  David again checks his rifle and spare ammo clips.   He carries a powerful bolt action .303 Lee Enfield rifle. The Lee Enfield was the primary battle rifle of the British Army through two world wars. His uncle, a special forces SAS combat veteran who fought in several wars, taught him to shoot this rifle. A powerful .303 bullet, besides being lethal to most living things, is an effective anti-materiel round.  A .303 bullet will destroy an engine block, flatten tires, and can otherwise disable a vehicle. David takes careful aim at the big ute far below.  400 meters is a long range shot through iron sights but the engine is a big target.  He will have to shoot past a rifleman but judges he can safely do so without hitting the man.  BANG! goes his first round.  The boy palms the bolt, reloads, and keeps firing. It takes the intruders a little while to realize someone is shooting at them and more time to recover from their initial shock and panic.  David's mad minute of shooting puts 5 bullets through the ute's engine block and flattens multiple tires.  That ute isn't driving away anytime soon.

The intruders now have a disabled vehicle, are in unfriendly territory, don't know the terrain, and lack adequate supplies to hike out.  They still manage to locate the shooter and return fire.  David feels bullets crack the air above his head and hears more bullets strike the rocks nearby. A small chip of rock lands harmlessly on his leg.  David's blood is up and he can clearly see one of the men shooting at him. David is tempted to kill the man but decides against such murderous foolishness. Instead he stops shooting. Still prone, he crawls back out of sight.  They continue to fire sporadically at his previous location.  David withdraws, recovers his horse, then rides away and circles around the intruders.

Twenty minutes later, from a distant ridgeline, the boy cautiously and stealthily observes the angry armed intruders. They do not spot him. The intruders are not prepared to escape on foot and their car's engine is scrap metal. David does not reveal his location and fires no more bullets that day.  He patiently observes.

By the time police arrive the intruders are glad to be rescued, even by police.   The police check everyone for safety, provide the intruders with water, then arrest them.  David keeps to his hidden vantage point and watches.   The intruders are handcuffed and locked in police vehicles.  Just before they drive away one police officer gets out of his car, looks towards the ridgeline where David watches from concealment, gives a gesture of approval, then gets back in his car and drives away with the crims.  David rides away, reports the action to Koori, and continues the serious job of riding the fence line of his parents' ranch.

A few days later the police visit David's parents.  The no-longer-a-boy accurately described the entire incident to his mum and dad.  The version told by police matches his story.  The police laud the lad for helping them catch the criminals, who are still in jail. They'll be released soon but are unlikely to return. The wrecked ute is towed away for scrap.

The same police officer who gestured his approval, we'll call him Officer Kevin (or Ken --- the name is uncertain after decades), wants to meet the lad.  They meet.  Officer Kevin, who is a military veteran, begins to keep tabs on David.  Kevin became David's friend and mentor.  Years later, when teenage David gets in trouble with the law, Officer Kevin is there to help.

I never verified any of David's stories about his childhood. I had no way to. They arrived fully formed, vivid, and internally consistent. Which is exactly what you'd expect from a good storyteller. Regardless of whether the stories were true.

\section*{The Method}
\label{pos05:the-method}

What I could verify was the teaching. David never lectured. He had a method, though I didn't recognize it as a pattern until years after he was gone.

He would ask me a question --- something I thought I already understood. I'd give him the standard answer, the one I'd learned from textbooks or from my own professional experience. David would look at me and say, ``Are you sure about that, mate?''

That was all. No correction. No alternative theory. Just five words and a look that said he already knew something I didn't.

I'd go home and research. Sometimes for hours, sometimes for days. And I would discover that the standard answer was incomplete, or wrong, or that it depended on assumptions nobody examined. The real answer was always stranger and more interesting than the textbook version. David had known it before he asked. He was never testing my knowledge --- he was testing whether I could find the edge of my own ignorance and push past it.

This is the Dunning-Kruger curve weaponized as pedagogy. Most people who think they understand a subject are sitting on the peak of Mount Stupid --- confident in an answer that happens to be incomplete. David's method was to tap that confidence with a single question and let gravity do the rest. The student falls off the peak, discovers how much they don't know, and has to climb back up on the other side with real understanding. He never pushed. He just asked the question and waited.

Every technical chapter in this book exists because David asked me a question I thought I could answer and couldn't. The autocatalytic sets, the edge of chaos, the topological protection, the thermal ladder --- I found all of them because David said ``Are you sure about that?'' and I went looking.

The method only works on someone who will actually do the research. David chose his students carefully.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

\section*{Srebrenica, July 1995}
\label{pos05:srebrenica-july-1995}

Healer told this story only once. What follows is his account, as close to his words as memory allows.

\vspace{0.5cm}

July 1995, near the city of Srebrenica in the war-torn country of Bosnia-Herzegovina, in the heart of central Europe. Srebrenica has been under siege for years. NATO officials suspect there will soon be a massacre of Bosniak Muslim civilians. Serbian regular army and paramilitary forces, equipped with tanks and artillery, have broken the Bosniak Muslim defenders in Srebrenica. Only 400 lightly armed Dutch peacekeepers stand between the Serbs and genocidal ethnic cleansing against tens of thousands of Bosniak civilians. UN nations watch in horror but do not intervene militarily.

I fall out of the moonlit night sky at terminal velocity from 11,000 meters. Far below lies the beleaguered city of Srebrenica. There has been no electricity down there for months, so all is dark. This will probably be my last military HALO jump. Compared to many other combat missions I've survived this one should be relatively low risk. My recent successful work as a DARPA scientist for GCHQ will soon make me unavailable for high-risk SAS military operations. My membership in a notorious elite hacker collective is still not known to my military superiors.

No Serbian anti-aircraft battery sees me fall out of the overcast sky. My chute opens a few hundred meters above the ground. The parachute slows my descent just before I touch the ground. My feet touch down a few seconds later\ldots{} ooof\ldots{} for a safe landing. The landing is clean and very hard to detect, even for a modern military. I land within a few meters of my objective on a high ridge overlooking the towns of Srebrenica and Potocari.

Training takes over while I hide my chute, stow my gear, and relocate to a suitable hiding place. I've done it literally dozens of times. That's why they sent me. Nearby hostile Serbian Army and paramilitary soldiers outnumber us thousands to one. On a different ridge is an ally whom I'll call Bourne. They must not detect our presence.

I find my concealed nest without seeing a hostile patrol. It is a large, well-shaped bush, located in an ideal position, perfect for hiding a sniper. I carefully hollow out a space in the center and conceal myself and my gear within. The hillside provides an excellent view of the city of Srebrenica. I observe strict scent-discipline in the hope of avoiding detection by hostile K9 patrols.

This operation is strictly `observe and record'. I observe carefully as Serbian Regular Army units and allied paramilitary forces prepare to storm the besieged city. Tens of thousands of terrified civilians shelter behind a few hundred lightly armed UN Peacekeepers. For both logistical and political reasons the UN was unable to send reinforcements to the UN Peacekeepers. The only nation both willing and able to send sufficient military force is nearby recently-reunited Germany. Living memory still recalls with horror the last time German soldiers occupied Bosnia-Herzegovina, during World War II\@. German soldiers in the Balkans are politically too contentious. Instead the UN sends me to witness and document a horror it can not stop.

The United Nations could not send reinforcements despite what looked like an impending genocidal massacre. It could send a few senior special forces officers to observe and document. This is my mission: observe, record, and give witness. I was explicitly ordered not to reveal my presence under any circumstance.

The Serbian Army is closing in. About 30,000 Bosniak Muslim refugees, in justifiable fear for their lives, huddle behind a few hundred lightly-armed blue-helmeted Dutch UN Peacekeepers.

The city is about to fall after three years of siege. The well-equipped besieging army consists of many thousands of both Serbian regular soldiers and paramilitary troops. The Serbs have tanks and artillery. The remainder of the defending Bosniak Muslim militia is lightly armed and has poor unit cohesion. The Bosnian regular army already retreated from Srebrenica to avoid total destruction. The small local UN Peacekeeping force had previously confiscated weapons from local Bosniak militia. The defenders are now so militarily weak that the attacking army is directly assaulting the city with infantry and getting little resistance. The sun rises day by day for five days, providing a good view of the beleaguered city. I observe the fighting with naked eye, telescopic scope, camera, and occasionally with top-of-the-line military night vision gear. The Serbian army and paramilitary forces sack the city. Their usual practice, at this stage of the bitter war, is to kill all the men then rape and forcibly relocate the women and children. This is the army I'm hiding from. My mission is to observe and report what happens here. Thousands of heavily armed Serbian soldiers assault the city's weakened defenders. The Serbs overrun several UN observation posts. Some UN peacekeepers fight on the Bosniak side before surrendering.

The Serbs soon offer the UN Protection Force (UNPROFOR) commander an awful choice: either fight and die against hopeless odds, or surrender and abandon the civilians. I watch as the 400 lightly armed UN Peacekeepers surrender to the numerous and heavily armed Serbian Army with their heavy tanks, artillery, and swarming infantry.

The massacre of civilians begins shortly thereafter. All I can do is watch in mounting horror. I witness and take careful notes. I see the very worst of what humans do to other humans. I document murder, rape, and torture on a massive scale. The victims are men, women, and children. The horror I witness in Srebrenica will haunt me to my dying day. This is the closest I ever come, in my long and distinguished military career, to disobeying direct orders. My orders are to remain concealed and report back. In my rifle sights I have the Serbian officers who give orders. They give orders to kill, torture, and rape. My blood boils in helpless anger. I could probably kill at least five Serbian officers before they kill me in turn, thus ending my mission with failure. With regrets for those I watch suffer and die I chose to follow orders. My primary recording tool is pen on waterproof paper. I also have a camera with good lenses.

A Serbian Paramilitary K9 patrol approaches my position! I freeze. The dog handlers stop 10 meters away from the bush where I hide. I know my scent discipline has been good enough when a dog urinates on me. They eventually leave. I continue my mission of observation and recording.

Days of horror follow. Finally, when the coast is clear and I can bear it no more, shaken as never before, I radio for extraction. The actual escape and extraction is easy. By then I know the movement patterns of the patrols. I escape overland on foot with light gear. I am eager to leave this place. The extraction goes as planned. I am thoroughly debriefed. My notes are copied. For the rest of my life I have fevered dreams where I relive the terrible river of death and horror I witnessed in Srebrenica.

My handwritten notes form the basis of my testimony at the UN War Crimes Tribunals for Slobodan Milosevic, the Serbian President who ordered the genocide, and also for Ratko Mladic and other Serbian leaders. There are hundreds of other witnesses but few with my field of view. Video recordings of the 1800 hours of trial are still available on the archive.org website.\footcite{milosevicarchive}

% SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line
During his trial Slobodan Milosevic acted as his own lawyer and personally cross-examined witnesses. He said some very strange things. I nearly choked when I heard that Milosevic had asked the distinguished human rights statistician, Dr.\ Patrick Ball, about some pro bono work he had done recently for Hacktivismo. ``So, Dr.\ Ball. Vaht can you tell me about zees Dead Cow Cult?'' How did Milosevic know about that? Hacktivismo is an organization dedicated to applying the 1948 Universal Declaration of Human Rights and the International Covenant on Civil and Political Rights to the Internet. That organization, an offshoot of the Cult of the Dead Cow hacker group, helped popularize the concept of hacktivism.

Several Serbian leaders were sentenced to lengthy prison terms for committing War Crimes. Some escape or are set free. Some stay on the run for years. One famously kills himself in court. Their punishment can not undo the evil they perpetrated.

Some months later I heard a strange thing while some SAS mates \& I drank a few pints at The Barrels pub. After a few pints a man will say things he'd usually be wise enough to keep to himself. Sergeant Doe usually kept quiet about what he saw. We all knew Doe did a lot of liaison work with the US' Agency (aka the CIA). Most of us had. Sometimes they wanted bodyguards, sometimes they needed someone in an official uniform to clear them past checkpoints. Always they wanted someone who could keep his mouth shut. Anyway, this is what Doe said on his fourth pint.

Doe looked around the table and said, ``What really burns me is the pedos. If you ask me, scum that prey on children should be shot.'' We all got more serious at that and listened closely to what Doe had to say. Doe went on.

``In the aftermath of the Bosnian war I was assigned to protect some Agency spooks and their guests from Dyncorp. They flew in from somewhere on an Agency plane. There was a translator, the Agency spook, the Dyncorps rep, some rich guy named Epstein, and a black haired woman who was obviously Epstein's handler. Their cover was `Humanitarian fact finding mission'. I was supposed to guard them and fill a uniform while they visited refugee camps.''

``At first it was on the up-and-up. They spoke to refugees. They asked questions. They even wrote down some answers. Strange thing was, it was obvious they didn't care about the answers. Like they were just going through the motions. What they really did was pick out the prettiest young refugee girls and fly them off to this Epstein character's private island. I watched them do it and heard them discuss it. They obviously had bad intentions towards those young girls they took. Normally I keep my mouth shut, you guys know that, but I really don't like pedos, nor those who assist them. Yet they have clearance from the top. What can we do?''

None of us could fault Doe for violating direct order and talking about a secret mission. We all hated pedos. At the same time, none of us could think of anything we could do against such corruption at the top. None of us were ready to turn whistleblower, knowing we would certainly fail and then certainly be destroyed. If only we could blow the whistle yet somehow remain anonymous then it might be an option. Hmm. Now there's an idea.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

\section*{The Patrick Ball Nexus}
\label{pos05:patrick-ball-nexus}

That exchange between Milosevic and Ball gnawed at me for years. I later learned why.

Dr.\ Patrick Ball --- Director of Research at the Human Rights Data Analysis Group --- is the publicly documented bridge between the ICTY and the Cult of the Dead Cow. Ball was the first expert witness called by the prosecution in the trial of Slobodan Milosevic, testifying on March 13--14, 2002, presenting a sixty-seven-page statistical analysis demonstrating a systematic campaign against Kosovar Albanians resulting in more than ten thousand deaths.\footcites{ballicty2002}{ictydecision} He later co-authored ``The Bosnian Book of the Dead'' (2007), the most complete database of Bosnian War casualties --- 97,207 names, including approximately 6,886 at Srebrenica.\footcite{foreignpolicy2012}

Ball's connection to the cDc is also public record. He spoke on a cDc-sponsored hacktivism panel at DEF CON 9 in July 2001 and served as an advisor to Hacktivismo. On March 14, 2002, Milosevic --- acting as his own defense attorney --- had obtained Ball's DEF CON talk transcript. He asked: ``Who is this Dead Cow Cult?'' and ``Are you in the advisory board of the Hacktivism group of international computer hackers?''

A human rights statistician who testifies about genocide and advises a hacker collective dedicated to the Universal Declaration of Human Rights. That is not a contradiction. That is a through-line.

The presence of British special forces at Srebrenica is independently documented. The Netherlands Institute for War Documentation confirmed that SAS and SBS personnel were embedded as ``Joint Commission Observers'' conducting ``reconnaissance missions.''\footcite{niod2002} A two-man SAS reconnaissance team was covertly inserted into the safe area. Three JCOs received British military honours with classified details. The Ministry of Defence sued the SAS patrol commander in 2002 to prevent publication. London forbade NIOD from interviewing the SAS operatives.

My proposition is that Stephen Wolfram and Stuart Kauffman were founding members of the cDc in 1984, and that Healer joined later. The cDc was a meritocracy based on hacking skill. The ``3 COWS'' --- Wolfram, Kauffman, and Healer --- are the three scientists who walked the room-temperature QNN out of the DARPA lab. ``BULLS are senior COWS'' --- a wordplay on BULLRUN, the NSA's known cryptanalysis program. The cDc is well documented for its identity practices: false identities, shared handles, and tradecraft misdirection. Oxblood Ruffin is publicly admitted as a shared handle.\footnote{One anomalous file in the cDc archive --- text \#283, ``Possibilities'' (1994) --- contains quantum mechanics and network topology vocabulary entirely outside the art-student corpus. Determined researchers should read it.}

I have seen the very worst of what humans do to each other. My dreams ever since have been filled with the horror of it, especially around the anniversary of the massacre. On the tenth anniversary of the massacre, the United Nations formally commemorated what Secretary General Kofi Annan had called, in his 1999 report, the worst mass murder perpetrated on European soil since World War II.\footcites{sudetic2010}{annan1999}

\chapterreturn
