% Chapter: The Alpha Farm (2003)
% Track: 2
% Plan: 0002

\settrack{tracktwo}
% VOICE: 1st-person personal (Bruce) — "I" not "Bruce". Personal/investigative voice.

\chapter{Alpha Farm (2003)}
\label{t2:ch01:alpha-farm}
\label{pos02:alpha-farm}

% Epigraph — Healer's HALO jump, Srebrenica 1995 (flash-forward hook)
\begin{quote}
\textit{I fall out of the moonlit night sky at terminal velocity from 11,000 meters. Far below lies the beleaguered city of Srebrenica. There has been no electricity down there for months, so all is dark. This will probably be my last military HALO jump.}

\hfill --- Healer,\footnote{For biographical details on Healer and the author, see Appendix: The Recruitment Dossiers.} July 1995
\end{quote}

\vspace{0.5cm}

It's snowing, getting cold, and visibility is near zero. Dusk settled an hour earlier. My tiny bicycle light shines the way toward Alpha Farm --- warmth, food, and shelter. Had I not scouted out this road already I'd fear death by exposure. My limbs are leaden.

I'm nearly there after a fifty-one-mile bike ride through wilderness. It's been mostly logging roads, in the snow, over the mountains of the Oregon Coast Range. Thirty-four hundred feet vertical climb. I began at dawn in a snowstorm. Movement is the only thing keeping me warm. Ahead I see the lights of Alpha Farm. I knew I could make the journey. My bags are strapped to my tiny aluminum Dahon folding bike. I started walking thirty minutes back --- it's too dark to ride.

A few weeks earlier I visited Alpha Farm, so I know my way around.

They know I am coming. Someone is watching. They see my tiny light out a window. Light floods from a door and someone with a flashlight comes out. ``Bruce? Is that you?'' A female voice. I walk the last twenty meters and park my bike under cover. Several people crowd out. ``He actually made it!''

Someone takes my bags to safety. They hustle me to the warm common room where a fire blazes on the hearth. They strip off my sodden clothing, wrap me in a warm blanket, and hand me a warm drink. The fire begins to warm my bones. I start the process of rehydrating and warming up.

I realize I'm wearing a pink muumuu.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

I should explain what kind of person arrived at Alpha Farm that night.

For the previous five years I've been a self-employed technical trainer, flying around the United States teaching three-to-five-day seminars for Fortune 500 companies. Basic and Advanced Java Programming, Secure Software Engineering, that sort of thing. Classes of eight to fifteen engineers, roughly twenty seminars a year. Training agencies sought me out, never the reverse. I was good at it. Occasionally I even got standing ovations.

The official curriculum was only half the job. I was a cypherpunk.  I used every class as an opportunity to seed the full cypherpunk message: your own government surveils you, encryption is a civil right,  build strong crypto anyway. In class I'd hint. At lunch I'd deliver --- half the class would join me for lunch, eager to pump me for everything I'd only hinted at on the clock. I taught them about public key cryptography, about government attempts to subvert it, about the brave Australians who leaked that Five Eyes nations didn't surveil their own citizens but got around it by surveilling each other's and exchanging data. I was ahead of the curve on Echelon, heard about the first releases. I was running a two-layer pedagogy before I knew there was a name for it.

One incident captures the dynamic. At IBM's Chicago office --- where I'd once arrived as an emergency substitute teacher, called up on a Saturday to fill in by Monday, knocked it out of the park, became a regular --- I was teaching about vulnerability scanning. I wrote a single word on the whiteboard: \textit{SATAN}. The Security Administrator Tool for Analysis of Networks, Dan Farmer and Wietse Venema's 1995 vulnerability scanner. The students were mesmerized. They laughed, they leaned in, they wanted more. At lunch, half the class treated me and pumped me for cypherpunk information. When we got back, a program director and two security guards were waiting to escort me out. The janitor had reported that the whiteboard was blank except for the word SATAN writ large. I understood instantly what had happened and was horrified. My students understood too. They told the program director about the vulnerability scanner and vouched for my class. The program director stood down security. The misunderstanding cleared in minutes. But the structure of that incident --- ignorance creating panic, pedagogy as defense, truth defusing the crisis --- would turn out to be the structure of my life. What I didn't see then is that a structure can be real even when the story it organizes is not.

IBM Learning Services offered me a full-time position at a hundred and forty-three thousand dollars a year. I turned it down. The job had me on the road forty-five weeks a year. My daughters Gillian and Fiona were six and three. My marriage was failing. If I took that job I'd be outsourcing all parenting, and I wanted to parent. Instead I took a job at FiServ in Corvallis for eighty-five thousand, writing enterprise Java security code. Less glamorous, less money, but I could be home for dinner.

That choice put me in Corvallis, Oregon, thirty miles as the crow flies from Alpha Farm. At FiServ I built a homebanking system middleware that became a primary product for several years. I had also co-founded several dotcoms. By 2003, at thirty-five, I'd collected my first real payout and retired. FiServ invited me back to train my replacements, then leave for a few weeks, then return and provide additional pedagogy. Friends gave me a small retirement party. I was thirty-five, financially independent, and for the first time in my adult life, I had time.

What I haven't said yet is that I was also broken. My marriage to a severely bipolar woman had ended after years of extreme caregiver burnout --- the kind where you stop sleeping, stop eating properly, and one day realize you can't remember what your own preferences are because you've spent so long managing someone else's crises. I'd had what I can only call a nervous breakdown. My psyche and spirit was broken. I came to Alpha Farm seeking something monastic --- silence, manual labor, a place where no one needed me to hold their reality together. I'd visited once alone that summer, then again with Gillian and Fiona in late summer. The Thanksgiving arrival was my third visit. Anyone reading my email traffic with Alpha Farm residents would have known exactly where I was headed and why.

\aidraft{Bletchley narrative pending Gen issue 4}

That choice also kept me in the intellectual orbit I'd been building for fifteen years. Michael Ruppert's newsletter \textit{From The Wilderness} --- CopVsCIA --- which I'd first encountered at Reed College. Jay Hanson and the dieoff.org community, where I'd co-administered discussion lists with Nate Hagens while physicists, petroleum geologists, intelligence analysts, and policy researchers argued about the end of industrial civilization. My Gmail signature for a decade was a quote from Dmitri Orlov "further economic growth is neither possible nor desirable". I was not a subtle person.

All of this was in my file. I know that now. Five Eyes had my browsing history, my email, my list correspondence, my training evaluations, my cypherpunk lunch lectures, my Gmail signature. When the odd Australian came downstairs at Alpha Farm and introduced himself as David, he already knew exactly who I was. The selection was surgical. He didn't find a translator by accident. He found one who'd spent five years teaching cryptographic literacy to Fortune 500 engineers, who understood the parable of the tribes, who had a father recruited by the CIA and smart enough to refuse, and who could follow guided deduction because he'd been trained to read the signatures of what isn't there.

I didn't know any of this yet. I was just a tired cyclist in a pink muumuu.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

A couple of hours later, yet still early in the evening, I'm warm and full and rapidly recovering. I'm thirty-five and an endurance athlete. I arrived after Thanksgiving feast but they've fed me aplenty.

The people at Alpha mention that there's a new visitor, a strange Australian man. They tell me he arrived ten days ago, got very sick, and has been resting upstairs ever since. I think nothing of it.

I hear footsteps coming down stairs and a man enters the common room. He's about six feet tall --- showing I still don't truly think in metric, growing up American --- and thin, with heavy shoulders and a medium frame, short curly hair, and a distinctive hair split. He greets me. ``Hello, mate. Name's David.'' He sounds like the Crocodile Hunter, strong Australian accent.

He sits down on a chair about two meters away. I'm on the couch, still wrapped in the pink muumuu but now warm and comfortable with a full belly. The fire crackles. Other people come and go. Someone tends the hearth. I've already chatted with the other eight or so people here --- three men and five women.

We do the basic introductions that people do, with the usual social back and forth. He mentions that he arrived in the USA ten days ago from Europe, entering through a port in Louisiana. Customs officials were a bit taken aback by him --- asked some questions, thought his presence odd --- but had to let him in. He's usually very healthy but got sick and fell over immediately upon arrival. He says he'd just recovered the day before. We talk about ordinary stuff. I tell him about my ride and I notice he raises an eyebrow. He comments that he used to do extreme sports but is taking it easy due to ongoing health issues. He says he's a military veteran now engaged in humanitarian work. He says he's on ``walkabout'' --- seeing the world after mustering out of government work.

After the basic introductions we keep talking. We're alone in the room. Others are about nearby and sometimes pass through. He mentions that he's a mathematician and scientist and I say, ``So am I.'' He starts mentioning more and more abstruse topics and I realize I know them. We talk about cryptography, physics, history of math and science, global politics and history. It quickly becomes clear that he's highly intelligent and has an excellent grasp of history, science, math, and computer engineering. In retrospect I realize he was holding back and giving me opportunity to talk. He seemed more interested in listening than speaking. I tell him about my background and activities. Despite his obvious erudition --- we'd long ago moved on to discussing topics that most people know nothing about --- he has an obvious blue-collar salt-of-the-earth quality associated with hard physical work. He is no ivory tower intellectual, even though his erudition could put to shame most of the professors I know.

We discuss athleticism and the connection between body and mind. We both opine that endurance activity boosts applied intelligence. He mentions that his sharpest intellectual accomplishments were concurrent with endurance athletic activity. I ask about these and he defers his answer for another day. The conversation turns to martial arts. I learn that he's an advanced black belt with a background in Chinese Shaolin kung fu and also in taijutsu --- the body combat aspect of ninjutsu. I mention my background in pencak silat --- six years of formal training plus another decade of informal sparring with brown and black belt fighters. I realize he's a much more skilled and experienced martial artist. He mentions that he's also a combat veteran and tells a single short combat story. I forget which one.

Other people join us for a while. Someone produces a cannabis joint or three --- quasi-legal in Oregon, and this is a remote hippie farm commune. A lady asks him how he's recovering from illness. She has apparently been taking care of him for his ten days of illness. She is a nurse midwife. He expresses gratitude for her care.

The evening begins to grow fuzzy. I started travelling at dawn. Exhaustion overcomes me. Around nine o'clock someone helps me find my bed.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

I wake up well after dawn. I'm in a second floor bedroom in an old farmhouse with a big central wood stove hearth. Light streams in through the window. Outside is white with snow. My room is cold but I'm warm under the pleated blankets. My bag is nearby, as is my water bottle. I'm still recovering water from yesterday's ride but wake up bouncy and ready to explore. I put on extra layers and a wool hat and get up. Delicious smells come from the kitchen: feast leftovers and coffee. The fire is roaring. Snow fell overnight --- not deep but just enough to ski. Out the windows is a winter wonderland.

Before breakfast I notice David standing near the little-used dining room table. Alpha Farm has the usual hippie commune library --- permaculture, alternative medicine, spiritual texts. I'd left something there on my previous visit: a printed paper copy of Jay Hanson's ``A Means of Control,'' the last scheduled Brain Food essay from dieoff.org. I'd written my name on it in pen, as well as a challenge to engage for anyone willing to do so. It was a litmus test. Most people would see the title and move on. Anyone who read it would understand that I thought civilization was running on borrowed time.

David is holding it. He looks up at me. ``Interesting,'' he says.

I soon realize that he is already quite familiar with the content. Not just the essay --- the entire intellectual framework. Energy depletion, the parable of the tribes, the impossibility of voluntary restraint, the thermodynamic argument against infinite growth. He knows all of it. He's not learning from my pamphlet. He's assessing whether I know it.

Years later, I understood what that moment meant. Intelligence agencies were watching dieoff.org. Of course they were. It was a private email list where physicists, petroleum geologists, rice farmers, retired intelligence analysts, and policy researchers discussed the end of industrial civilization. That list had Dick Cheney's people on it. It had Matthew Simmons, the biggest energy investment banker in the world and Dick Cheney's golf buddy. It had intelligence operatives crowdsourcing on geophysics. And it had me, later co-administering with Nate Hagens, broadcasting the parable of the tribes to anyone who would listen. David already knew all of this. He'd read my file.  At the time I did not understand and thought little of it.

Over coffee and breakfast, I resume conversation with David. Other people about so we're chatting lightly over the breakfast table. He's good at keeping his discussions grounded for the folk about us. I notice he gets into a conversation with the nurse about medical stuff. David is a trained combat medic. They speak in medic jargon. They have both delivered babies and they discuss the process. David mentions that he liked bringing babies into the world and that he preferred it to treating combat injuries.

This is when I give him the mental appellation ``Healer.''

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

% Merged from pos03 (The Mentor), Session 29.
% Original pos03 preserved but commented out of main.tex.
% 3rd-person biography reframed to 1st-person.
% Kangaroo story is in pos05. Two conflicting versions published as evidence for A.

Over the days and weeks that followed, David told me about his life. He'd grown up on a horse ranch in New South Wales, riding fence and shooting a .303 Lee Enfield rifle. He'd stolen dozens of cars by the age of thirteen --- a thrill, until Juvenile Detention. A police officer named Ken, who'd seen him handle a crisis on the ranch, testified for him in court and arranged a job as a mechanic instead of juvi. David liked work. He liked what money could do. He taught himself computers, learned how the telephone system worked, and enlisted in the Australian military at seventeen.

Testing showed his aptitudes were off the charts. He trained as a combat medic, a communications specialist, and a soldier. In 1984, at nineteen, he was assigned to patrol duty in Northern Ireland during the Troubles. He killed his first enemy combatant --- an IRA sniper --- at extreme range, while under fire and with his wounded mate beside him. That drew the attention of the Special Air Service. He made it through selection. He was field promoted to officer --- a rare distinction in Commonwealth special forces, where the path from enlisted to commissioned usually requires formal coursework and a selection board. Bruce recalls that he claimed he held the rank of Captain in the SAS.

His unit was called Bravo Two Zero. Books and movies tell that story. In those accounts he is known as Steven ``Legs'' Lane, and supposedly dies of hypothermia after swimming across the Euphrates. His ``death'' freed him for classified work.

\section*{An Intellectual Equal}
\label{pos02:an-intellectual-equal}

David fascinated me. I was a child prodigy, unusually talented at both language and mathematics, yet it quickly became clear to me that he was more than my intellectual equal. We were able to communicate privately in a strange way: we could hold perfectly lucid conversations in front of other people, yet the others would have little idea what topic we had really discussed. It helped that we both had a solid understanding of mathematics, physics, computer programming and networks, history, politics, current events, military strategy and tactics, psychology, anthropology, literature, history of science, philosophy, and especially computer security systems. We both liked to speak using double and triple layered meanings, and were perfectly able to understand each other this way.

\section*{The Kitchen}
\label{pos02:the-kitchen}

David never lectured. He asked questions. We discussed fourier transforms. He probed what I knew about steganography. I was reasonably familiar, although not as much as I thought. We talked about the TCP/IP stack, common messaging protocols, and practical techniques for hiding data in network traffic. He challenged me to describe likely candidates for steganographic messaging over existing public networks. I suggested embedding jittered time-phase shifts in a ubiquitous protocol like NTP.  That's the same principle GPS uses to help receivers acquire a weak signal. He acknowledged that timing jitter in TCP/IP packets could embed multiple layers of data with substantial bandwidth --- again, just like GPS. Checksums would be required. David was very into checksums and their various approaches. Checksums help with error correction. He claimed to have played some part in developing the SHA-2 protocol. This caught my attention, because that is top-cryptographer work. We reviewed most of the Physics Nobel Prize work between 1980 and 2002.

Some of his questions led to quantum teleportation, which was newly confirmed in laboratories. State of the art in 2003 was measured in centimeters or meters. From the Nobel review I knew about the fractional quantum Hall effect, Laughlin's quantum fluids, anyons with their exotic statistics. I knew that decoherence was the killer problem. Quantum states are fragile, thermal noise destroys them, and no one had a clean solution for maintaining coherence over useful distances. I also knew that FQHE anyons behaved strangely. The mechanism wasn't clear to me yet. What \textit{was} clear was that quantum teleportation could theoretically send information if you had a classical backchannel --- and jittered NTP could serve as that backchannel. I realized it was theoretically possible, if the decoherence problem could be solved, to cause a message to tunnel across a network and emerge, as if out of an EPR wormhole, at any node. My brain had been operating at maximum efficiency the entire time. The intellectual challenge was comparable to my Advanced QM class from Crandall, only I had my own private professor asking questions to motivate my education. I wasn't thinking about \textit{why} he was doing it. I was just accepting the event. From the start it felt like a historic event occurring in real time.

So there we were in the kitchen doing KP duty, three or four weeks into this process. Several other people were around. David and I had learned to communicate in layers --- we could discuss seemingly mundane topics while the real conversation ran underneath, carried by references, quotes, and resonances that only we would catch. That private channel let him \textit{say} things without actually saying them.

While we worked he told what sounded like an amusing story about some silly thing his friends had done involving minor misuse of computers. Not illegal but definitely naughty. But the story, heard through our layered channel, implied they had added an additional layer to the TCP/IP stack between the physical layer and the logical layers. This implied actually having found a way to send a message through an EPR-style wormhole. This implied the ability to cut through any security perimeter by already being inside it. It implied that all computer security measures were fruitless. An attacker which could simply extrude a tendril of itself from any node anywhere in the system. This was crazy-advanced science fiction stuff already on the cusp of reality.

Everything came together in my brain at once. Steganography. Quantum teleportation. EPR correlations. Anyons that resisted decoherence for reasons I didn't yet fully understand. Truly random noise indistinguishable from no signal at all. A classical backchannel hidden in the timing jitter of infrastructure protocols that automated scanners treated as background noise. The decoherence problem wasn't solved in my head --- but I could see that possible solutions might exist. All of it synthesized into a single coherent picture. I realized that his absurd story could be true. That it was physically plausible. That the implications were staggering.

I started laughing uncontrollably. I fell down on the kitchen floor, howling with hilarity. It caused quite a disruption. People surrounded my prone body, still wracked with paroxysms of laughter. ``What's wrong? Is something funny?'' I could not answer except to gasp ``I'm alright!'' and keep laughing.

David waited. Then he made eye contact. A few raised eyebrows said everything. He could tell that the pieces had all come together and I was laughing at the implications of what he had just said. Without actually \textit{saying} it. His expression asked: \textit{are you laughing about what I think you are?} I expressioned back: \textit{yes.}

Perhaps I passed another test that day, because our lessons grew more intense. I started consuming science papers as fast as I could --- many each night, then meditation and time for synthesis. He knew not to overload me and quickly learned to match my rate of absorption. I didn't understand what strange game this was but it was exhilarating.

\section*{A Soldier's Contradictions}
\label{pos02:a-soldiers-contradictions}

David also told me stories of death and violence that chilled my blood. He described witnessing the very worst of human behavior. In the worst of these, even worse than his counter-insurgency actions in Central America, he had been a NATO sniper hiding in a bush during the Balkans War after the surrender of the Dutch peacekeeping force outside Srebrenica. He witnessed the Srebrenica Massacre, the largest mass murder in Europe since WWII, through a telescopic rifle scope, such that it seemed very close.

In the course of learning all this military stuff it became apparent that my friend had killed people many times over. I initially found this a bit disturbing, but I got over it. He must have been at least a fourth degree black belt, probably higher. I believe that the video game character of Timothy Hanley, in the computer game Tom Clancy's Rainbow Six, was loosely based on David. He vividly described the interview with Tom Clancy.

\section*{The Ninja Aspect}
\label{pos02:the-ninja-aspect}

David claimed the third floor of the main house at Alpha Farm. I managed to secure a second story room with a telephone line, which I could use for Internet access via my laptop's 56K modem. David had arrived with hardly any possessions, and sometimes expressed fury at the *\&\%*\$! who stole his kit bag, laptop computer, and all his gear, during airport baggage handling just before he came to Alpha Farm. My room looked out on the only stair. I began to realize that David occasionally came and went from his third floor nest without using the only stairway. He must have (de)fenestrated and climbed. For someone who climbed K2 this would be trivial. This was another hint that he had many of the skills one would expect of a master ninja.

\section*{The Hacker}
\label{pos02:the-hacker}

One time I saw David leave the Alpha Farm office. The single public computer on the premises was in there. He slipped a CD into his pocket. He had already mentioned that he was an active software developer and Unix user who contributed to Open Source software projects. I pointed to the CD in his pocket and asked which flavour of Live Unix OS lived on that boot-able CD. He realized I had just ``made'' him. He asked how I had known. He admitted that it was a ``very custom'' Linux OS live-CD. From then on he began to open up to me about computer-related topics.

\section*{K2}
\label{pos02:k2}

David told me his personal story of redemption. He had already told me some stories about the things he had to do when he served as a ``Special Operative,'' with no name or identity. He was deeply ashamed of certain violent acts he had committed, always under orders, and of certain ways in which he had killed. He told how, in 1996, he was mountain climbing with other off-duty SAS soldiers on leave. He stood on the summit of K2, a thousand meter drop in all directions, and had a moment of introspection. He thought, ``What the hell am I doing with my life?'' He resolved, right then, to try to Give Something Back to Humanity, in order to partially atone for the evil and violent acts he had committed during some of his nastiest missions. That was the moment when he resolved to use his formidable abilities to accomplish something more than ``killing bad guys.''

He would later anchor that vow to a specific document: the Universal Declaration of Human Rights, drafted in 1948 in the aftermath of atrocities that had overwhelmed every existing ethical framework.

\section*{Someone Who Fits a Particular Profile}
\label{pos02:someone-who-fits-a-particular-profile}

It was early January 2004, about six weeks into this process. Each resident at Alpha Farm was expected to work a forty-hour week doing farm work --- splitting firewood, repairing roofs and cabinets, cleaning, large animal care, motor pool maintenance. All the unglamorous labour that keeps a working farm functional. We were out splitting firewood together.

By then I understood what an exceptional individual he was. I understood that he was a man in black, a high-ranking SAS officer, a genius-level hacker and scientist. The question had been building in me for weeks. I asked him: ``Mate, what are you doing here? What's someone with your exceptional skills and aptitudes doing at a place like this?''

His answer went right over my head. I only understood it years later, after he departed.

``I'm looking for someone, mate. Someone who fits a particular profile.''

This completely baffled me and I dropped the topic.

\section*{The Water System}
\label{pos02:the-water-system}

On my earlier visit to Alpha Farm --- a week or two, a few weeks before the bike ride --- I had learned that the farm's water system was inadequate. It pulled from a surface stream that went dry in August some years. The terrain was right for a gravity-fed system. I had already said in my emails that I hoped to gift Alpha Farm with a proper water supply.

In retrospect, someone reading my email would have seen that. David had arrived at Alpha Farm claiming to be a water systems expert. The perfect coincidence. He actually was one --- a legacy of growing up on a horse station in outback New South Wales, where you either understood water infrastructure or you watched your livestock die. He could survey a watershed, calculate head pressure, and design a gravity-fed system from memory.

We proposed exactly that: a gravity-fed water system for Alpha Farm. I would fund it --- roughly seventy thousand dollars' worth of materials. David, myself, and two younger residents named Gabe and Celina would provide the labor. David would design and oversee the engineering. The system would give the farm reliable, sustainable water independent of electric pumps. I wanted to embed a bronze plaque in the pump house with my daughters' fingerprints cast into it --- insurance, so that if Gillian or Fiona ever showed up at this remote farm in need of help, they could prove their father had built the place's water supply. I was already thinking about contingencies I couldn't articulate.

We held meetings. Every resident unanimously wanted the water system. Caroline, half of the elderly couple who controlled the farm's governance after thirty years, blocked it anyway --- not by voting against it openly, but by scheduling our work assignments such that we never had time to do the project. She never admitted that's what she was doing. It was dishonest obstruction dressed up as farm management.

We had been warned. The locals we saw regularly while working at the Alpha Bit restaurant had told us exactly what would happen. Alpha Farm would take in ragged people who needed help, but would block anyone who tried to give anything to the community. It was how the people on top maintained their supremacy. The locals shook their heads and said, ``We told you so.''

David told me he had hoped to conduct the entire process at Alpha Farm. The remote location, the self-sufficiency, the cover of communal life --- it was ideal for what he needed. But the place was corrupt, not honest, and he would not work in a dishonest environment.

All four of us left together, frustrated and embittered. Gabe and Celina made two trips in their VW bus --- one to stage our belongings, then a second where we all loaded in and drove away. I offered temporary space at the Veggie House, a communal house in Corvallis near the OSU campus where I had a studio room. But a studio was too small for me, David, and a young couple --- Gabe and Celina lived in their van and used my place as a home base, but it was too crowded. I rented a three-bedroom house in Adair Village, a rural community north of Corvallis, and all four of us relocated there.

\section*{The Ground Rules}
\label{pos02:the-ground-rules}

Before we settled in at Adair Village, David set out the rules for our collaboration. He hoped I would volunteer for a particular project that needed someone with my particular profile. He had conditions.

First: the military method. If he insisted on something, I agreed to follow his instructions. This was not negotiable. He used the phrase ``military method'' often --- it was a refrain throughout our time together, a shorthand for disciplined execution. One maxim: once a plan is made, stick to the plan unless there's a reason to change it.

Second: cleanliness. He could not tolerate mess or sloppiness. His living and working spaces would be orderly. So would mine.

Third: don't play music by the Grateful Dead out loud. Headphones were fine. It wasn't my kind of music anyway, but he made his preferences --- and what he could tolerate --- very clear.

Fourth: don't peek at his computer screen when he wasn't around. He wanted to be able to trust me with his privacy so he didn't need to secure his terminal every time he walked away. This was a given for me. In my mind, violating someone's privacy is taboo.

I agreed to all four. The rules told me something about the man: disciplined, orderly, specific about boundaries, and willing to state them clearly rather than impose them by force. The Grateful Dead rule told me he was human.

\section*{The Mentorship}
\label{pos02:the-mentorship}

I spent the next thirty months doing everything I could to assist David with his various projects. Honestly, that was not very much. I provided logistical support, brought him food and medicine, cleaned house, and provided the best company I could, all while conducting my own research into the tidbits of information he tossed my way. I learned all I could about tradecraft. I also applied myself to abstruse but relevant topics in mathematics, theoretical physics, computer science, and, oddly enough, theoretical biology.

David worked in long cycles --- twelve hours on, twelve hours off. He needed reliable Internet, and when the DSL connection went down he was furious. ``Did you remember to pay the bill, mate?'' I had. It was a provider outage. But the lost connectivity cost him days of work, and I learned how seriously he took his operational windows. Whatever he was doing on that custom Linux CD required sustained, uninterrupted network access.

Near the end of our time in Adair Village, David announced that ``man does not live on bread alone'' and that he was interested in meeting women. I had been separated from my wife for about nine months. I still hoped to repair my marriage, so I'd been celibate --- faithful to my soon-to-be ex-wife. I'd also never really dated. Not as an adult, not properly. Years of a failing marriage had left me with a low self-image, and the whole ritual of courtship was foreign territory. But I immediately took his request as a mission --- one I was starting to think about for myself too.

I reached out to some of the women I knew from the Veggie House and arranged social visits. Some turned out well for everyone involved. The same sort of woman who might be drawn to a fictional James Bond or a real Oliver North was drawn to David. He had no problem getting dates. His greatest struggle was being normal --- not bringing up esoteric mathematics or gruesome combat stories over dinner. Pretty soon we had a lady or two visiting our place regularly. He treated them very well, as a gentleman should.

Eventually I acknowledged that my marriage was likely over and created a profile on an online dating site. My first date was with a sweet woman I'll call VR --- ZB's mother. After the second or third date she invited me to move in with her, and said I could bring my odd friend, who could live in the garage. She understood what she was signing up for and knew there was risk involved, especially from government agencies. She knew David and I were up to something naughty.

That was how we came to live at 15th and Grant in Corvallis. By spring of 2005 or 2006, ZB, a teenager who posted daily math homework assignments on her high school's class website with her teacher's authorization, tried to update the page and found the school's Windows NT server down. She brought it up in conversation while David was in the adjacent room, apparently absorbed in his own work. David called out to her from the next room: ``Try the server again now, ZB.'' She tried. The server came back up. ZB and I looked at each other. Neither of us had told David the school's name, let alone its network address. He had simply overheard the problem and fixed it --- remotely, from the next room, on a machine he had no business being able to reach.

\section*{Assessment}
\label{pos02:assessment}

Everyone around me told me they thought David was a dangerous madman up to some mysterious mischief. I was mostly not able to share what I learned with others, because the information was too strange, complicated, and outside their frame of reference. In truth, he was something of a madman. I can't imagine anyone surviving what he went through without going partially mad. Also, of course, geniuses often seem mad to those who don't understand them. I never lost faith in David's essential good intentions. Although my trust was shaken several times, it was never broken. Several times I suspected he was lying to me. Years later I learned that he had actually been speaking plain truth to me in every case I was able to verify.

\chapterreturn
