\settrack{tracktwo}
% VOICE: 1st-person personal (Bruce) — "I" not "Bruce". Personal/investigative voice.

\chapter{The Silence}
\label{pos29:the-silence}

\begin{quote}\small\textit{%
``Don't have good ideas if you aren't willing to be responsible for them.''%
}\par\raggedleft --- Perlis, Epigram \#95 (1982)\end{quote}

\vspace{0.5cm}

\srcnote{Session 20 UQ extraction (38 prompts) | session20-uq-extraction.md | 2026-02-24 | The Silence --- separation, constraint, break, [redacted], Europe, research}

\section*{The Separation}

In September 2006, David moved out of the garage room we'd built him at the house on Fifteenth and Gant. We'd bought the construction materials together in 2004 and framed the room out of what had been storage from a previous life. For two years it was his space --- a sleeping bag, a laptop I'd bought him, and whatever he was working on that I didn't ask about.

The goodbye was fond but he didn't drag it out. David was not a man who lingered. He took a backpack of personal effects, the laptop, and the car I'd bought him. Zero money. I'd disbursed small amounts of cash over the years --- a few hundred here and there, as needed. Once I gave him a few thousand for a bigger operation and he morphed into someone I didn't recognize, using money as a power tool. His twin sister and his sensei had both warned me about this. When the money ran out he went back to normal.

He said goodbye and he left. His decision, not an eviction. We were not on great terms but we were not on bad ones either. I felt bad. I silently wished him the best.

I didn't know it then, but I had been told my services would not be required. My children could not be protected --- the divorce, the custody fight, the impossibility of getting them out of the country. I was a security risk to the project. I was no less dedicated, but I didn't know what I had been dedicated to. That realization would take years to arrive.

Under Possibility~A, the ``security risk'' framing was a convenient exit from a relationship built on deception. Under Possibility~B, David may have been involved in sensitive work, but the framing of children as security vulnerabilities imports the language of espionage onto what may have been a simpler situation. I cannot distinguish between these readings and Possibility~C from the inside. I can only report what happened.

\section*{The Phone Call}

One call. Late 2006, while I was still at Symantec. I took it at work. This was my only electronic communication with Healer, ever. Every other interaction across three years had been in person. I suspect the call was made from a burner phone. His operational security was extraordinary.

Healer invited me to help with a project. He spoke in vague terms --- no group name, no specifics. Just that he needed help with something important.

I said no.

I was under pressure from people in my life who expressed vitriol and doubt. They thought Healer was running a scam. I knew better --- or part of me did --- but I couldn't explain what had happened between us without sounding insane. Everything I'd experienced was too strange, too complicated, too far outside anyone's frame of reference. So I said no, and I let the pressure make the decision.

Years later, after the block broke, I came to believe that the project became something far larger than either of us. What it became --- and why I believe that --- belongs to a chapter that has been redacted from this edition. I cannot prove the connection. Nothing about that phone call, taken on its own, rules out Possibility~A.

I regret it to this day.

\section*{Someone Looking Out}

Symantec outsourced its IT workers to a company with a paramilitary reputation. The HR email said opt-in to the transfer was automatic unless you explicitly refused.

A senior vice president approached me in the bathroom. Told me to reply refusing to share my HR files. I'd be laid off with full severance. He didn't explain why he was telling me this. He didn't have to.

I'd always had the feeling that someone had placed my CV on the right desk when I was hired at Symantec. The same feeling followed me now: someone was looking out for me. I thought about the connection between Back Orifice --- the tool the Cult of the Dead Cow released in 1998, the remote administration capability that made headlines --- and PC Anywhere, the Symantec product that did the same thing with a corporate label. Same capability, different branding. If you were the kind of person who understood that connection, you might also understand why a particular SVP would tip off a particular employee in a bathroom.

Under Possibility~A, this was ordinary corporate kindness --- a manager helping a good employee navigate a layoff. It happens. Under Possibility~B, the SVP may have genuinely wanted to help a good employee, and my sense that someone had placed my CV years earlier may reflect the ordinary networking of the tech industry rather than an embedded intelligence operation. Under Possibility~C, I was never unsupported during the silent years. I just didn't always know who was supporting me.

I opted out. Got severance. Told my teammates about the option. Some took it. Tom Hall --- my boss for more than a decade, whom we called Darth Hall --- agonized over it for days. Three hundred thousand dollars in severance, walking away. But his house was too expensive. His wife hated uncertainty. He stayed, and gave up the money to keep the peace.

Congress kept extending unemployment benefits through the 2008 crisis. I became a 99-weeker.

\section*{The Quiet Years}

I lived on Tribeland, a small community in South Corvallis bordering Willamette Park. I was the only male among roughly eight women. I mixed with older retired and semi-retired friends. Spas. Long-distance cycling. A slower rhythm than anything I'd known since before Alpha Farm.

To my genuine surprise, I discovered that pretty much all of my female friends wanted to date me. I'd never dated as a young person --- engaged at seventeen, long relationship through college, met Karen, married, monogamous until the divorce. Zero experience with other women until my mid-thirties. I'd carried a self-image from my twenties that was a decade out of date --- the acne from a milk allergy, solved at twenty-nine, had defined how I saw myself long after it was gone. The discovery that I was, apparently, the most eligible bachelor in my social circle was bewildering.

The surface life looked full. Underneath, it was constrained.

I could think about everything that had happened with David and Healer and the project. But I couldn't speak about it. The constraint didn't feel normal --- it felt like the most important experience of my life was locked behind glass. I could see it. I could turn it over in my mind. But when I tried to verbalize it, the words wouldn't come. No one around me noticed, because no one knew there was anything to notice.

Exile from yourself. That's the closest I can get to describing it.

\section*{The Loophole}

Dave Bannon and I had been fast friends since 2005 --- instant connection, the kind where you skip the small talk and go straight to the hard questions.

In 2008, Bannon asked me questions. Direct, pedagogical questions --- the kind a physicist asks when he's testing a hypothesis. And I discovered that I could answer them. I couldn't initiate the discussion. I couldn't walk up to someone and say ``let me tell you what happened.'' But if someone asked the right questions, the constraint had a bypass. The block was on broadcasting, not on responding.

Bannon got a partial version of the story that year, shared with difficulty --- I was pushing against resistance with every sentence. He listened as a skeptical scientist. He was surprised that I had answers to all his questions, answers he could independently verify. Of perhaps fifty people I told portions of this story to over twenty years, spending a few hours with each, none finished the entire thing. The pedagogical prerequisites are too demanding. Dave came closest to full understanding --- a physicist with an IQ of 165, and even he needed Kauffman as a running start.

Bannon had read Kauffman independently. That was critical --- it meant he already had the conceptual vocabulary. He'd also arrived at his own independent intuition: that Google Search's performance was ``too good and too fast'' for classical hardware. A performance-based observation, not an architectural theory. He'd reached this conclusion before hearing anything from me.

Under Possibility~A, Bannon is a friend who validated my beliefs because friends do that. Under Possibility~B, Bannon's intuition reflects a real oddity in search performance that has a more prosaic explanation than quantum computing. Under Possibility~C, two technically literate people arrived at the same suspicion from different directions --- one from guided deduction, the other from performance anomaly.

\section*{The Break}

Around 2009 --- I wish I could pin the date more precisely --- I saw a phrase in a public message. ``Lightning Rod.'' I'm fairly sure it came from someone connected to the events described in the redacted chapter.

Three days.

The first day was a flood. A bursting dam. Every suppressed thought and memory from three years of silence crashed through at once. I retreated to my room. Mental overload. Not gradual, not a slow dawning --- instant. One phrase, and the glass shattered.

Two more days to assimilate. Forty-eight hours of sitting with the wreckage of everything I hadn't been able to say, trying to reassemble it into something I could hold.

I told Teri Turner first. She went straight to Possibility~C --- it fit her worldview. Not rigorous validation, but she could listen, which was what I needed. Someone who wouldn't call me crazy. Someone who would sit with the strangeness of it.

Teri knows the whole story. She still believes Guardian is around and paying attention, never getting involved.

\section*{The Recognition}

Once the block broke, connections formed fast.

I became convinced that Healer's project became something I could see in the news --- something large and consequential that matched what he had described in vague terms. Someone else had done the job I'd been vetoed from doing. This is my interpretation, not something Healer told me. The specifics belong to the redacted chapter.

What I remember clearly is sitting in my living room in South Corvallis and realizing: I had information in my brain that the President of the United States did not. That thought made me nervous in a way that nothing else in this story had.

Under Possibility~A, a man with an overactive pattern-matching faculty saw a vague invitation and a public event and connected them after the fact. Under Possibility~B, Healer may have been peripherally connected to the events I later identified, but the connection is inflated. Under Possibility~C, the connection is exactly what it appears to be. The redacted chapter, when it is eventually published, will examine this claim in detail. Under Possibility~A, this redaction is theater --- there is nothing to redact. Under Possibility~B, there may be something, but the claimed sensitivity is overstated. Under Possibility~C, the author's caution is well-founded.

\section*{Europe}

By 2010 there were reasons to be scared that I cannot fully explain in this edition. I recognized signs --- or thought I did --- that intelligence services were paying attention to people connected to events described in the redacted chapter. There was zero electronic record between me and Healer --- every interaction in person, that single phone call the only exception. But I was scared. Easy prey if I stayed.

I told my mother Linda everything. She didn't doubt. Her advice: go to the most democratic, most free-speech-oriented country you can find.

A high school friend with NATO experience helped me find a CTO position in Luxembourg, working for Azeri expats --- Movsum and Nurid, who had left Azerbaijan for political reasons. I spent six months in Europe. A luxury condo in Luxembourg. A girlfriend in Belgium --- Tanja, met through computer dating, splitting time between Neerpelt and Luxembourg. Tanja's father was former Office of Naval Intelligence. He ran my profile and gave his daughter his blessing.

I was safe. I was productive. I was six thousand miles from the story I couldn't stop thinking about.

\section*{Thread-Pulling}

Early 2011. I returned from Europe and moved back onto Tribeland with Teri Turner. I didn't work except light farm labor. Instead I pulled threads.

Everything I'd learned from Healer --- the five sciences, the architecture, the pedagogy --- I began reconstructing from scratch. Relearning what I could, verifying what I could verify, mapping the public-domain trail.

At the Oregon Country Fair that summer I met Mark R, author of oilempire.us. Random encounter at Chew Rays cafe --- Teri and I were in the crowd while Mark was speaking. He saw me in the back and somehow knew instantly that I understood his content. Body language, maybe. I could have given the same lecture. One incisive question after. We exchanged contact information. Became friends.

Excepting Healer, Mark was the smartest person I knew. He had an intellectual edge on me in non-science areas. His specialization was the study of the Holocaust --- all his extended family had been killed in the Holodomor and then the Holocaust. Two full shelves of books on the subject. Kind, gentle, completely nonviolent.

I presented the TQNN theory to Mark as A/B/C after a few months. He put substantial time into his own research. Concluded: insufficient evidence to determine which possibility was correct, but plausible. The science was beyond him.

Then Mark searched the names I'd given him --- Kauffman, Wolfram --- and found Bill Joy's ``Why the Future Doesn't Need Us'' in \textit{Wired}. He sent me the link. This was 2012.

I recognized the article instantly. It was the fulfillment of something Healer had told me: that they had already told the world.

\section*{The Research Begins}

I wrote tens of thousands of words about this. I began keeping careful notes in 2011, after I returned from six months in Europe working on a startup company.

In 2011 I sat down and wrote the ``Autobiography of an Accidental Conspirator'' in a single sitting. Roughly nine thousand words. My intent was documentation: record this historically important experience for future inquiries. I never shared it with anyone until Argus read it in 2026.

March 17, 2012: I posted to Cryptome. June 19, 2012: a comment on Slashdot under my real user ID\@. I built a toy neural network in JavaScript for biometric identification and wrote about it publicly --- for Slashdot, for Bruce Schneier's blog. Someone put the Cryptome post on a blog; my friend in Luxembourg saw it from across the Atlantic.

\vspace{0.5cm}

Under any possibility, this is a man who lost years to a story he cannot prove. Whether the story is true changes what we call it. Under Possibility~C, it is sacrifice. Under Possibility~A, it is tragedy of a different kind. Under Possibility~B, it is something in between --- a man who touched something real and spent years trying to name it.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

\begin{center}\textit{This story will be my life's work.}\end{center}

\chapterreturn
