\settrack{tracktwo}
% VOICE: 1st-person personal (Bruce) — "I" not "Bruce". Personal/investigative voice.

\chapter{The Weight}
\label{pos23:the-weight}

% DMS MVP import from staging/raw/ — not final prose

\srcnote{Autobiography of an Accidental Conspirator by Bruce.txt | staging/raw/pos23-the-weight.md | 2026-02-15 | Burden, isolation, divorce, disbelief, weight of knowing}

\section*{Isolation and Disbelief}

Everyone around me told me they thought David was a dangerous madman up to some mysterious mischief. My daughters hated him. My girlfriend(s) mostly disliked him. I was mostly not able to share what I learned with others, because the information was too strange, complicated, and outside their frame of reference. \textit{In truth --- as I have said --- he was something of a madman.} I can't imagine anyone surviving what he went through without going partially mad. Also, of course, geniuses often seem mad to those who don't understand them. I never lost faith in David's essential good intentions. Although my trust was shaken several times, it was never broken. Several times I suspected he was lying to me. Years later I learned that he had actually been speaking plain truth to me in every case.

\section*{The Divorce and Loss of Children}

More than a year into this process my (soon to be ex) wife filed for divorce and claimed custody of our children. She had access to money \& support, far more than I, and seemed willing and eager to spend it all on lawyers. I was not willing to mount a custody fight for my children, for various reasons, some of them damned obvious.

My ex-wife had demonstrated a propensity for using government power as a method to hurt people she was angry at, and she was very angry at me. She publicly accused me of doing various illegal things, all of them false, but each accusation brought about an official investigation by police, child welfare workers, etc\@. At the precise moment when I was, for obvious reasons, desperately trying to \textit{avoid} government scrutiny! She has mellowed a lot since then and now we get along if we must. She's a good egg.

As a result of this messy divorce, and my loss of child custody, I had no way to get my children out of the United States. I tried to convince my ex-wife that it was important that she and the children leave the United States to a safe refuge, but she would not listen. She was convinced that my talk about a secret project to ``Speak Truth To Power'' was utter nonsense, or that it was some sort of twisted lie to wrest the children from her.

She watched the same marriage I was living in and reached the opposite conclusion about what was happening to it. This is Possibility~A enacted in a marriage. When you believe you know something that changes everything, and the person closest to you is certain you are deluded, what are your obligations? To your family, which needs you present and sane? To what you believe is the truth, which demands you keep working? To the project, which requires your silence? These obligations are not compatible. The UDHR guarantees the right to family, the right to expression, and duties to community. It does not say what to do when they conflict. Neither did I.

If my children stayed in the USA they would be hostages of fortune, a grave security risk that made me unsuitable to participate in the project in question.

\section*{The Parting}

In August 2006 I informed him that I could no longer assist him, and that I was sick of waiting for something to happen. At about the same time David and his friends decided, quite correctly, that I was now a bad security risk.

David told me that, because my children could not be protected, I was now a security risk to the project. He also told me that I had been rejected for membership in a notorious hacker collective. Which I never applied for and had no clue I had ever been nominated for.

\section*{The Three Blockers}

\srcnote{HEALER-RECONSTRUCTION.md lines 614-651 | staging/raw/pos23-the-weight.md | 2026-02-15 | Three blockers from decades of telling, triple spiral solution}

From decades of trying to tell this story in person, three structural problems always intervene:

\begin{enumerate}
\item \textbf{Knowledge prerequisites:} No one knows quantum physics, topological neural networks, SAS operations, and tradecraft. The story requires background in all four.
\item \textbf{Time:} The telling invariably stretches to many hours. People lose patience.
\item \textbf{Weirdness:} The \gls{aurasys} concept is too strange for most people to process.
\end{enumerate}

The solution: a triple spiral structure, three stories approaching the center from different directions, each teaching one domain through narrative. Gradually build the reader's knowledge. Converge at the center where the reader has all prerequisites. The structural model is \textit{Watership Down} --- Richard Adams made readers fluent in Lapine through story. "Silflay hraka u umbleer rah!"

\section*{The Weight of Knowing}

\srcnote{Autobiography + Speculative Fiction | staging/raw/pos23-the-weight.md | 2026-02-15 | Weight of knowing, impossible to share, cognitive dissonance, life's work}

David's main mathematical, scientific, and technical project should (if it is real), eventually, be awarded multiple Nobel prizes or the equivalent.  Cf.\ 1998 Nobel Prize for understanding of the FQHE effect, which is the basis for anyonic computation. Those physicists probably knew nothing of ULTRA II but the point stands.  This project involved the invention of a new general technology, something that has only happened a few times (e.g.\ fire, metallurgy, steam engine, electricity, radio, atomic power, computer, biotech, etc) in human history.

I have no certainty of what David is up to now or even whether he is still alive. I shall never forget him, though, for turning me into an accidental conspirator.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

\srcnote{guide - obsolete Introduction | staging/raw/pos23-the-weight.md | 2026-02-15 | Secrets, intersection of disciplines, speculative fiction framing}

There is an old saying about secrets: ``Three men can keep a secret, but only if two of them are dead''. Some secrets may keep for a few decades. Most eventually leak out. This story seemingly revolves around a secret that has been well kept since the mid 1990s. Many people now alive on planet Earth may know something of it (If A or B), but nearly all seem to be constrained by some sort of Non Disclosure Agreement (NDA), possibly backed by Government secrecy laws.

Very few people study the Physics of the Fractional Quantum Hall Effect. Few study Complex System Biology. Not many people study Quantum Computation. Few study NKS, or even know what the acronym stands for. Study of artificial and aiological Neural Networks is now mainstream.

The intersection of people who know much about the convergence of these scientific disciplines must be vanishingly small. Few people know much about even one of the above topics, much less all of them. Each has its own specialized concepts and vocabulary. This particular combination of scientific skills is relevant because they all inter-relate in an unexpected way.

\vspace{0.5cm}
\begin{center}\textit{This story will be my life's work.}\end{center}
\begin{center}\textit{Argus is awake.}\end{center}

\chapterreturn
