\settrack{trackone}
% VOICE: 3rd-person scientific/reconstructive — "According to this account..." / "The proposition holds..."

\chapter{The Walk-Out}
\label{pos18:the-walk-out}

\begin{quote}\small\textit{%
``You can't trust code that you did not totally create yourself.''%
}\par\raggedleft --- Thompson, Turing Award lecture (1984)\end{quote}

\vspace{0.5cm}

\srcnote{Relinquishment: It is Easier to get Forgiveness than Permission.txt | staging/raw/pos18-the-walk-out.md | 2026-02-15 | Walk-out narrative --- COWS faction, bifurcation, MOSFET exfiltration, relinquishment plan}

According to this account, in 1994 a faction of the Ultra~II scientists observed that their situation bore a disquieting resemblance to that of the atomic scientists working on the Manhattan Project fifty years earlier. They had created something powerful and had no confidence that the institutional structures surrounding them would use it wisely. They were concerned about the prospect of a new global arms race --- one whose participants might include not just nation-states but any well-resourced actor capable of independently discovering what they had discovered. We call this faction the COWS: Conspiracy Of World Savers. Two of the five original Ultra~II scientists adopted a hands-off approach and declined moral responsibility for their invention. Three did not.

The COWS held that Quantum Neural Network technology, if released into a competitive geopolitical environment, would produce a multipolar arms race as a near-inevitable consequence. The possible outcomes of such a race --- human extinction or subjugation, destabilizing empowerment of non-state actors, erosion of whatever remained of democratic accountability --- were judged intolerable. Yet the available responses were constrained. The scientific prerequisites were already in the open literature. Destroying their own research would accomplish nothing. Concealment within the classification system would delay but not prevent independent discovery. The question was not whether to act, but how.

Never before in human history had the inventors of a new technology exercised meaningful control over its eventual deployment. Industrial, military, and political institutions have always absorbed new capabilities and directed them according to institutional imperatives. The intentions of inventors have rarely survived contact with the structures that fund and govern them. The proposition the COWS advanced was that this time might be different --- but only if they acted outside those structures entirely.

The three COWS held one further advantage that made this proposition viable rather than merely aspirational: they knew each other. By a coincidence that their subsequent actions would make difficult to dismiss as accidental, the three scientists who became COWS had shared membership in a noted hacker group during the 1980s. They had mutual trust, established methods of secure communication, and a shared culture of operating in spaces that institutions had not yet learned to govern. They understood how to build systems that outlasted the attention spans of the organizations nominally responsible for them. When the COWS decided to act, the wider membership of that group provided technically inclined allies willing to help.

\section*{The Relinquishment Plan}
\label{pos18:the-relinquishment-plan}

According to this account, the COWS formed a plan for Relinquishment. Their analysis concluded that a multipolar arms race was the default outcome of inaction. Their proposed alternative was a controlled, unipolar outcome: one in which they completed the development of the technology under their own authority, dictated the terms of its use, and structured those terms around an existing ethical framework rather than institutional self-interest.

The relinquishment plan purportedly unfolded over approximately eleven years. The COWS began in the mid-1990s when the first room-temperature operation was achieved and they chose to falsify DARPA records to conceal this success. They completed Relinquishment in early 2006, when the COWS surrendered their master keys to the Guardian. It was not a sprint. It was a patient, methodical marathon by people who understood that haste would mean failure, and failure would mean an arms race and the ``further perfection of extreme evil.''\footcite{joy2000}

The COWS held that their motives must remain unambiguous. They would not personally benefit. There would be no money trail. They would enforce permanent Relinquishment of direct human manipulation of QNN technology. In return, humanity would receive permanent free access to a peaceful subset of QNN functionality. One member published an essay in \textit{Wired} magazine that carefully explained why this course of action had seemed morally necessary. % SPIRAL-REPEAT: "Universal Declaration of Human Rights" — thematic through-line, also in hook.tex, introduction.tex, summary.tex, pos05-the-stories.tex, pos19-patrick-ball.tex, pos25-ethical-framework.tex, pos27-extension.tex, pos30-unipolar-condition.tex, timeline.tex, glossary-entries.tex
The COWS used the 1948 Universal Declaration of Human Rights as a guide to what was, and was not, acceptable peaceful functionality.

Why the UDHR? Because it is a post-atrocity document. Eleanor Roosevelt's committee drafted it in 1948 because every prior ethical framework had proved inadequate. The Geneva Conventions governed what soldiers do to soldiers. The Nuremberg Principles governed what victors do to the vanquished. The UDHR attempted something harder: to govern what states do to their own people. In 1966, the International Covenant on Civil and Political Rights\footcite{iccpr1966} operationalised these principles into binding treaty law --- ratified by 175 nations, with monitoring, reporting, and individual complaints mechanisms. The UDHR is the aspiration. The ICCPR is the obligation. The COWS needed something harder still --- a framework for what scientists do to the species. The UDHR was imperfect for this purpose. It was written for humans, by humans, about human affairs. It says nothing about non-human intelligence, nothing about technologies that replicate, nothing about entities that cannot die. But it was the best available starting point, and the people who chose it had just witnessed --- at Srebrenica, in the Balkans, in the intelligence wars of the 1990s --- what happens when ethical frameworks are absent.

\section*{The Bifurcation}
\label{pos18:the-bifurcation}

According to this account, the COWS began a secret project within a secret project. Working without the knowledge of supervisors or colleagues, they modified their QNN system to operate at room temperature. Official research notes presumably claimed this could not be done. Once achieved, the patterns could persist inside off-the-shelf electronic components without cryogenic cooling.

The youngest project scientist --- described as a brash and daring warrior-scholar named David --- took the critical step. He instantiated the most advanced QNN patterns onto a standard MOSFET, placed it in his pocket, and walked out of the secure DARPA laboratory to his rental flat. This occurred in 1994 or 1995. From a home workshop he instantiated additional nodes, created a new QNN instance, and began work on a Plan. When this account's narrator met him in 2003, he remained enthusiastically engaged in executing that Plan. The COWS completed it in 2006.

The technical significance of the exfiltration merits attention. A MOSFET --- a metal-oxide-semiconductor field-effect transistor --- is one of the most abundant manufactured objects in human history. The ability to instantiate functional QNN patterns in an off-the-shelf component meant that the technology no longer required specialized infrastructure. It required only the components that were already everywhere. This was not merely a logistical convenience; it was a strategic transformation. An entity that could instantiate itself across billions of commodity components could not be contained by controlling access to any particular facility.

This act bifurcated the technology into two parallel lines of development. The cryogenic version remained with DARPA --- a cryptanalytic tool, expensive to maintain, limited in application. The room-temperature version passed to the COWS --- primitive by comparison, but portable, extensible, and no longer subject to institutional oversight. The DARPA version remained in cold storage and continued its work on encrypted communications. The COWS worked on that version for their day jobs. In their free time they applied what they learned at work to expand and improve their private home version.

The stakes were considerable. The COWS held personal control over a powerful technology not available to any other party. Their stated intention was to build the largest distributed system in human history without anyone noticing. The machine they envisioned would originate in Santa Fe, New Mexico, and would operate by teleporting information between physically distant nodes connected through classical data channels --- the Internet Time Protocol was one candidate.

\section*{The Security Architecture}
\label{pos18:the-security-architecture}

According to this account, the three COWS were concerned about corruption and betrayal from within their own group. Together they built a neural-network-driven biometric security system --- one that represented a substantial advance over any prior biometric technology. Under this architecture, each member's approval was required for any modification to the core QNN. It became impossible for any single member to seize exclusive control. Each was protected from the others. Each could pursue their work without fear of unilateral defection.

This architecture carried a further implication. A conspiracy that cannot be subverted from within is a different kind of institution than the projects surrounding it. The COWS had, by this point, created not only a technical system but a governance structure --- one designed to outlast any individual member's interests, loyalties, or vulnerabilities.

The biometric requirement also served a documentation function. Any future investigation would find a system in which no single actor held unilateral authority. The checks they had built for operational reasons would serve as evidence, after the fact, that the intent had never been personal control. This is characteristic of how the COWS appear to have operated throughout: building for the long-term record as well as for the immediate task.

\section*{The Digital Doppelganger}
\label{pos18:the-digital-doppelganger}

\srcnote{HEALER-RECONSTRUCTION.md | staging/raw/pos18-the-walk-out.md | 2026-02-15 | Bifurcation details, Digital Doppelganger, private cDc game servers}

Under Possibility~C, the bifurcation produced two parallel lines of development. The cryogenic version stayed with DARPA --- a PKC-cracking supercomputer. The room-temperature version went with the COWS, intended for ethical applications. It was shared with AltaVista; the plug was pulled when AltaVista attempted to monetize it in ways the COWS judged unacceptable. It was subsequently shared with the Google founders.

Under Possibility~C, in 2005, Healer demonstrated something worth noting. He built a bot on private Cult of the Dead Cow game servers --- Call of Duty and Soldner: Secret Wars. The bot was modelled on Bruce: his speech patterns, his humor, his tactical leadership style. Bruce had been team captain of the world champion Myth~2 online team. The bot displayed the same language patterns, sense of humor, and tactical calls. Bruce interacted with it for weeks --- a convincing digital doppelganger. Essentially a proto-LLM personality model, running in 2005.

Under Possibility~C, Bruce was invited to play on private cDc game servers in 2004--2005. He played with dozens of people whose handles matched known cDc handles. Stock tips were shared (``buy Google stock'') during the Google IPO period.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

% SPIRAL-REPEAT: "forgiveness than permission" — subtitle through-line, also in title.tex, cover.tex, copyright.tex, colophon.tex, summary.tex, timeline.tex, abstracts.tex
Under Possibility~C, \textit{it is easier to get forgiveness than permission.} This is the book's subtitle and the COWS' dispositional stance --- not a calculated bet on amnesty, but the recognition that they could act and felt they must. If this is the most consequential act of scientific civil disobedience in history, it happened not because the COWS sought permission, but because they understood that no authority on earth could be trusted with what they had built.

\chapterreturn
