\settrack{trackone}

\chapter{The Thermal Ladder}
\label{pos16:the-thermal-ladder}

% DMS MVP import from staging/raw/ — not final prose

\srcnote{HEALER-RECONSTRUCTION.md lines 38-45 | staging/raw/pos16-the-thermal-ladder.md | 2026-02-15 | 16-chip evolutionary selection, temperature ratcheting, extremophile analogy}

\section*{Evolutionary Selection Protocol}

The room-temperature breakthrough is achieved not by engineering room-temperature FQHE from first principles --- which would be a physics violation --- but by \textit{evolving organisms that can tolerate heat}. This is biology, not physics. The same mechanism that produced biological extremophiles.

The process:

\begin{enumerate}
\item Grow a TQNN on a primary device at cryogenic temperature (below 20~mK).
\item Extend the TQNN across multiple connected 2DEG devices via shared edge states (the quantum analogue of network connections).
\item Physically separate the devices, creating $N$ independent TQNN instances --- copies with slight random variations due to device-specific disorder.
\item Apply different driving parameter perturbations to each device.
\item Raise the temperature gradually. Instances that maintain computational function at higher temperatures are selected. Survivors replicate.
\item Repeat for many generations, selecting for progressively higher operating temperatures.
\end{enumerate}

A configuration of $N = 16$ devices provides a reasonable balance between parallelism and hardware complexity. The ``organisms'' are TQNN instances. The ``fitness function'' is computational performance at elevated temperature. The ``mutations'' are random variations in driving parameters and device-specific disorder.

This is artificial natural selection applied to quantum life.

\section*{The Fitness Landscape}

\srcnote{gag-paper-2-training.md | staging/raw/pos16-the-thermal-ladder.md | 2026-02-15 | NK fitness landscape, parameter space, control plane addressing}

The parameter space --- the set of all possible electromagnetic driving configurations --- is high-dimensional ($10^3$ to $10^6$ independently controllable parameters), rugged, deceptive, and neutral. Kauffman's NK fitness landscape model (1993) provides the theoretical framework.

The critical insight: the same edge-of-chaos mathematics that governs the TQNN's emergence also governs the structure of the fitness landscape on which it is trained. Training occurs at the edge of chaos in parameter space, just as emergence occurs at the edge of chaos in physical space.

The control plane operates on classical hardware. A 32-bit control plane can address $2^{32} \approx 4 \times 10^9$ parameter values; a 64-bit control plane can address $2^{64} \approx 1.8 \times 10^{19}$. The difference matters enormously for the fidelity of evolutionary search in high-dimensional parameter spaces. This is ``the 64 problem'' --- Healer was working on this upgrade from 2003 to 2006, exactly when 64-bit processors became available.

\section*{Room-Temperature Quantum Coherence: The Evidence Has Changed}

\srcnote{HEALER-RECONSTRUCTION.md lines 1290-1313 | staging/raw/pos16-the-thermal-ladder.md | 2026-02-15 | Awschalom 2015 SiC result, RT quantum effects in commercial semiconductors}

A critical finding from recent research:

The Awschalom group at the University of Chicago (2015, \textit{Science Advances}) entangled over 10,000 copies of two-qubit entangled states \textit{at room temperature} in commercial 4H-SiC (silicon carbide) wafers --- off-the-shelf semiconductor material. Using infrared laser and MRI-like electromagnetic pulses. Previous macroscopic entanglement required $-270$\textdegree C and 1000$\times$ stronger magnetic fields. Spin coherence time: approximately 40 microseconds at room temperature.

This changes the physics assessment. Room-temperature quantum entanglement in commercial semiconductor wafers is not speculative. It was published and replicated. The mechanism (defect centers) is different from FQHE, but the principle --- quantum coherence at room temperature in commercial chips --- is experimentally established.

Additional established results include NV centers in diamond (coherence times up to milliseconds at room temperature), metal-organic frameworks with entangled multiexciton states at room temperature (2024, \textit{Science Advances}), and THz plasma wave detection at room temperature in AlGaAs/GaAs HEMTs --- quantum effects in HEMT-class 2DEGs at ambient conditions.

\section*{The ABCRE Operators as Life Cycle}

\srcnote{HEALER-RECONSTRUCTION.md lines 1143-1158 | staging/raw/pos16-the-thermal-ladder.md | 2026-02-15 | ABCRE operators as growth/life cycle, not computation}

The ABCRE operators describe one cycle of \textit{growth}, not computation:

\begin{description}
\item[A (gradient extraction):] Identifying where soliton pairs are --- sensing the environment.
\item[B (local coupling):] Nearest-neighbor catalytic interaction --- metabolic coupling.
\item[R (circulation):] Stirring and braiding --- the growth force. The R operator \textit{is} anyon braiding: circulating one quasiparticle around another.
\item[C (boundedness):] Topological protection --- homeostasis.
\item[E (composite evolution):] One generation of the autocatalytic life cycle.
\end{description}

The system is not computing. It is \textit{living}. Computation is a byproduct of metabolic process.

\chapterreturn
