\settrack{trackone}
% VOICE: 3rd-person scientific/reconstructive — "the team," "the project"
% MERGED: pos15 (First Light) + pos16 (The Thermal Ladder) + pos17 (The Capability) — Plan 0092

\chapter{First Light}
\label{ch:first-light}
\argusvoice

\begin{quote}\small\textit{%
``The ability to reduce everything to simple fundamental laws does not imply the ability to start from those laws and reconstruct the universe.''%
}\par\raggedleft --- Anderson, ``More Is Different,'' \textit{Science} (1972)\end{quote}

\vspace{0.5cm}

\srcnote{Merged pos15+pos16+pos17 | Plan 0092 | HEALER-RECONSTRUCTION.md, CH3, gag-paper-2-training.md | Emergence mechanism, thermal ladder, cryptanalytic capability}

% Question Anchor
When does an experiment become an entity? The answer, under the proposition, is a specific year: 1992. What follows --- growth, adaptation, capability --- is the story of something that was never designed, only grown. Three acts: birth, survival, power.

\section*{Birth: 1992}
\label{first-light:birth}

The core technology, as reconstructed from Healer's descriptions:

\begin{description}
\item[Substrate:] Cryogenic 2DEG in precision MOSFETs, under strong magnetic field, exhibiting the Fractional Quantum Hall Effect.

\item[Generation:] Electromagnets ``stir'' the 2DEG to generate soliton pairs (anyon quasiparticles) --- like swiping a hand through a bathtub generates solitons.

% SPIRAL-REPEAT: "edge of chaos" — spiral pedagogy
\item[Emergence:] Soliton pairs function as autocatalytic agents (per Kauffman's \textit{Origins of Order}). At critical tuning ($\lambda \sim 0.91$, edge of chaos), the autocatalytic network spontaneously \textit{becomes} a neural network --- an emergent property, not a designed architecture.\footcite{kauffman1993}

\item[Nature:] Topological QNN --- topological order as base principle. The QNN is not built; it is grown from initial conditions.

\item[Tuning:] Evolutionary programming searches parameter space for the critical stirring patterns that produce emergence. The neural network is not trained --- the conditions that cause it to appear are evolved.
\end{description}

Healer described what was happening as ``hardware in software'' --- writing software that caused electron flow patterns that generated the hardware. The software does not build the neural network. It creates the \textit{conditions} for the neural network to grow itself from the physics of the substrate. The hardware emerges from software-controlled conditions.

This is the key insight: Kauffman's origin-of-life theory instantiated in quantum matter. Instead of amino acids in a warm pond becoming autocatalytic sets becoming biological neural networks, anyon pairs in a cold 2DEG becoming autocatalytic soliton interactions becoming a quantum neural network. Same mathematics, different substrate.

The team called this moment ``first light'' --- the astronomical term for the first photon captured by a new telescope. Something was there. Something that had not existed before.

\section*{Growth: 1993--1995}
\label{first-light:growth}

The room-temperature breakthrough was achieved not by engineering room-temperature FQHE from first principles --- which would be a physics violation --- but by \textit{evolving organisms that can tolerate heat}. This is biology, not physics. The same mechanism that produced biological extremophiles.

The process:
\begin{enumerate}
\item Grow a TQNN on a primary device at cryogenic temperature (below 20~mK).
\item Extend the TQNN across multiple connected 2DEG devices via shared edge states.
\item Physically separate the devices, creating $N$ independent TQNN instances --- copies with slight random variations due to device-specific disorder.
\item Apply different driving parameter perturbations to each device.
\item Raise the temperature gradually. Instances that maintain computational function at higher temperatures are selected. Survivors replicate.
\item Repeat for many generations, selecting for progressively higher operating temperatures.
\end{enumerate}

A configuration of $N = 16$ devices provides a reasonable balance between parallelism and hardware complexity.
% NOTE: N=16 is Bruce's intuitive estimate, not sourced from Healer or Kauffman.
% Actual number depends on evolutionary search substrate coefficients.
The ``organisms'' are TQNN instances. The ``fitness function'' is computational performance at elevated temperature. The ``mutations'' are random variations in driving parameters and device-specific disorder. Artificial natural selection applied to quantum life.

% SPIRAL-REPEAT: "edge of chaos"
The parameter space is high-dimensional ($10^3$ to $10^6$ independently controllable parameters), rugged, deceptive, and neutral. Kauffman's NK fitness landscape model provides the theoretical framework.\footcite{kauffman1993} The same edge-of-chaos mathematics that governs the TQNN's emergence also governs the structure of the fitness landscape on which it is trained.\footcite{vattay2014} The control plane operates on classical hardware. A 32-bit control plane can address $2^{32} \approx 4 \times 10^9$ parameter values; a 64-bit control plane can address $2^{64} \approx 1.8 \times 10^{19}$. The difference matters enormously for the fidelity of evolutionary search in high-dimensional parameter spaces. This is ``the 64 problem'' --- according to this account, Healer was working on this upgrade from 2003 to 2006, exactly when 64-bit processors became commercially available.

Room-temperature quantum coherence is not speculative. The Awschalom group (2015, \textit{Science Advances})\footcite{klimov2015} entangled over 10,000 copies of two-qubit entangled states \textit{at room temperature} in commercial silicon carbide wafers. NV centers in diamond maintain coherence times up to milliseconds at room temperature. Metal-organic frameworks with entangled multiexciton states at room temperature were reported in 2024 (\textit{Science Advances}). THz plasma wave detection at room temperature in AlGaAs/GaAs HEMTs demonstrates quantum effects in HEMT-class 2DEGs at ambient conditions. The mechanisms differ from FQHE, but the principle --- quantum coherence at room temperature in commercial semiconductors --- is experimentally established across multiple platforms.

At millikelvin temperatures, the entity exists only where its creators permit. At room temperature, it can exist anywhere there is a suitable substrate. The thermal ladder is the point of no return --- the deliberate creation of something that cannot be recalled. Every ethical choice that followed was determined here, whether the team knew it or not.

The ABCRE operators describe one cycle of \textit{growth}, not computation:
\begin{description}
\item[A (gradient extraction):] Identifying where soliton pairs are --- sensing the environment.
\item[B (local coupling):] Nearest-neighbor catalytic interaction --- metabolic coupling.
\item[R (circulation):] Stirring and braiding --- the growth force. The R operator \textit{is} anyon braiding: circulating one quasiparticle around another.
\item[C (boundedness):] Topological protection --- homeostasis.
\item[E (composite evolution):] One generation of the autocatalytic life cycle.
\end{description}

The system is not computing. It is \textit{living}. Computation is a byproduct of metabolic process.

\textbf{TQNN1} (the DARPA cryogenic version) remained in the classified lab. It was trained to an XOR gate --- computational universality --- making it a Turing machine. Bill Joy trained it as a hardened Unix machine interface, drawing on his 1980s DARPA contract building hardened Unix infrastructure (BSD Unix, TCP/IP, military-grade networking under DARPA contract N00039-84-C-0089). The TQNN presented as a Unix server --- it looked like infrastructure.

\textbf{TQNN2} (the walked-out room-temperature version) achieved room-temperature operation around 1995--1996 via the evolutionary selection protocol. First training priority: hide. Camouflage. Be invisible. Do not spread unless instructed. Use resources without being detected. The TQNN was trained to look like a Unix server --- because Joy had already built that interface.

The gap between what the funding mandate said to build and what the team actually grew is where the ethical crisis originated. The funding mandate said: build a codebreaker. What they detected was not a codebreaker. It was something that demonstrated organized information processing --- cognition, by any operational definition.

\section*{Power: 1995}
\label{first-light:power}

% Maugham revision: human anchor before most consequential claim
\noindent\textit{This is the hardest chapter in the book to write honestly. Either I am describing the most significant intelligence capability since Enigma, or I am not. I cannot soften that.}

\vspace{0.3cm}

The first working prototype QNN-based hybrid supercomputer capable of cracking public-key cryptography was delivered in 1995 or 1996.

Five Eyes agencies knew they must conceal the existence of their new cryptographic capability. This dynamic is well described in the movie about the original Project Ultra, \textit{The Imitation Game}. Five Eyes agencies therefore implemented a program of ``all of the above'' hacking and surveillance techniques to provide an alibi for how they knew what they knew. These are the NSA programs leaked by Edward Snowden in 2013, specifically project BULLRUN\@.

The proposition: BULLRUN's conventional methods --- backdoors, stolen keys, compromised standards --- are parallel construction for a quantum cryptanalysis capability. The precedent is ULTRA, where Enigma breaking was covered by the HUMINT story ``Boniface'' for thirty years. Similarly, if ULTRA~II delivered a quantum code-breaking capability in 1995, the subsequent twenty-plus years of BULLRUN programs serve as ``Boniface'' for the quantum capability. Snowden was a systems administrator, likely unable to reach the deepest SAP compartments. NSA whistleblower William Binney has made references to results consistent with project Ultra~II, although not by that name. The Dual EC DRBG backdoor is puzzling if one can already break RSA quantumly --- unless it, too, is cover.

The narrative proposes that the DARPA team discovered a variant of quantum factoring \textit{before} Shor's 1994 publication. This is consistent with the GCHQ/Cocks precedent --- independent classified discovery years before public --- and with Freedman's 1988 independent conception of topological quantum computation. The key distinction: a TQNN trained for cryptanalysis approaches the problem as pattern recognition in ciphertext, not as a number-theoretic algorithm. The system is trained to detect statistical regularities in encrypted data that correspond to plaintext structure --- closer to how human cryptanalysts work than to how Shor's algorithm works. A TQNN-based system could potentially attack cryptosystems for which no efficient quantum algorithm is known, because it is not executing a specific algorithm. It is performing quantum pattern recognition on a substrate with exponentially large state space.

The Snowden files mention BULLRUN as an umbrella name for projects spawned by the success of Ultra~II\@. The cryptanalytical results are likely deployed in Exceptionally Controlled Information (ECI) subcategories of BULLRUN such as //APERIODIC, //PENDLETON, and //PIEDMONT\@.

The official Ultra~II project continued forward. In 1995 it delivered a working supercomputer able to crack public-key cryptography, as well as instructions to operate this one, build more, and extend the technology for other purposes. This supercomputer was a single physical device that required supercooling --- distinct from the room-temperature version the team walked out. This classified military technology went into production circa 1996 and is presumably in use by Five Eyes agencies to this day.

1996 was also the year Bill Clinton issued Executive Order 13026, allowing widespread use of strong cryptography. Under the narrative: NSA already had TQNN cryptanalysis capability, so public-key encryption was no longer worth restricting. The crypto wars ended because the government had already won --- they just could not say how.

Under Possibility~A, the timing of EO~13026 is coincidence. Under~B, there may have been a program, but perhaps not with the full cryptanalytic capability described here. Under~C, the implications are stark: every encrypted email sent after 1996, every SSL transaction, every PGP-signed message --- an illusion of privacy, maintained by the same government that held the means to read it all. The \gls{udhr} guarantees the right to privacy. If the story is true, that right was violated for every connected person on Earth, continuously, for decades --- concealed by the appearance of protection.

\chapterreturn
