\settrack{trackone}
% VOICE: 3rd-person scientific/reconstructive — "According to this account..." / "The proposition holds..."

\chapter{The Capability}
\label{pos17:the-capability}

\begin{quote}\small\textit{%
``The enemy knows the system.''%
}\par\raggedleft --- Shannon, ``Communication Theory of Secrecy Systems'' (1949)\end{quote}

\vspace{0.5cm}

% DMS MVP import from staging/raw/ — not final prose

\srcnote{CH3 - Practical Cryptography \& project ULTRA II.txt | staging/raw/pos17-the-capability.md | 2026-02-15 | Code-breaking machine, BULLRUN, ULTRA precedent, EO 13026}

% Maugham revision: human anchor before most consequential claim
\noindent\textit{This is the hardest chapter in the book to write honestly. Either I am describing the most significant intelligence capability since Enigma, or I am not. I cannot soften that.}

\section*{The Code-Breaking Machine}
\label{pos17:the-code-breaking-machine}

The first working prototype QNN-based hybrid supercomputer capable of cracking public key cryptography was delivered in 1995 or 1996.

Five Eyes agencies knew they must conceal the existence of their new cryptographic technology. This dynamic is very well described in the movie about the original Project Ultra, ``The Imitation Game''. Five Eyes agencies therefore implemented a program of `all of the above' hacking and surveillance techniques to provide an alibi for how they know what they know. These are the NSA programs leaked by Edward Snowden in 2013, specifically project BULLRUN\@. NSA whistleblower William Binney has made references to the results of project Ultra~II, although not by that name.

\section*{BULLRUN as Cover Story}
\label{pos17:bullrun-as-cover-story}

\srcnote{HEALER-RECONSTRUCTION.md lines 77-83 | staging/raw/pos17-the-capability.md | 2026-02-15 | BULLRUN parallel construction, ULTRA precedent, Snowden compartmentation}

The proposition: BULLRUN's conventional methods (backdoors, stolen keys, compromised standards) are parallel construction for a quantum cryptanalysis capability. The precedent is ULTRA --- Enigma breaking was covered by the HUMINT story ``Boniface'' for thirty years.

The original Project ULTRA was kept secret from 1943 to 1974. During that time, the Allies created elaborate cover stories to explain intelligence derived from decrypted German communications. Similarly, if ULTRA~II delivered a quantum code-breaking capability in 1995, the subsequent twenty-plus years of BULLRUN programs serve as ``Boniface'' for the quantum capability --- parallel construction that explains what they know without revealing how they know it.

Supporting evidence: Snowden was a systems administrator, likely unable to reach the deepest SAP compartments. The Dual EC DRBG backdoor is puzzling if you can already break RSA quantumly --- unless it is cover.

\section*{Pattern Recognition, Not Algorithm}
\label{pos17:pattern-recognition-not-algorithm}

\srcnote{HEALER-RECONSTRUCTION.md lines 1788-1815 | staging/raw/pos17-the-capability.md | 2026-02-15 | Pre-Shor variant, GCHQ/Cocks precedent, pattern recognition vs algorithmic}

The narrative proposes that the DARPA team discovered a variant of quantum factoring \textit{before} Shor's 1994 publication. This is consistent with the GCHQ/Cocks precedent (independent classified discovery years before public) and Freedman's 1988 independent conception.

The key distinction: a TQNN trained for cryptanalysis approaches the problem as pattern recognition in ciphertext, not as a number-theoretic algorithm. The system is trained to detect statistical regularities in encrypted data that correspond to plaintext structure. This is closer to how human cryptanalysts work --- recognizing patterns --- than to how Shor's algorithm works.

A TQNN-based cryptanalytic system could potentially attack cryptosystems for which no efficient quantum algorithm is known, because it is not executing a specific algorithm. It is performing quantum pattern recognition on a substrate with exponentially large state space. The specific capability depends on training --- each cipher type is a feature set.

\section*{BULLRUN and ECI Subcategories}
\label{pos17:bullrun-and-eci-subcategories}

The Snowden files mention a project BULLRUN\@. BULLRUN seems to be an umbrella name for projects spawned by the success of Ultra~II\@. This author suspects the cryptanalytical results of Ultra~II are deployed in Exceptionally Controlled Information (ECI) subcategories of BULLRUN such as //APERIODIC, //PENDLETON, and //PIEDMONT\@.

\section*{The Official Delivery}
\label{pos17:the-official-delivery}

The official Ultra~II project continued forward. In 1995 it delivered a working supercomputer able to crack PKC, as well as instructions to operate this one, build more, and extend the technology for other purposes. This supercomputer was a single physical device that required supercooling. This classified military technology went into production circa 1996 and is presumably in use by NSA and other Five Eyes agencies to this day.

\section*{Executive Order 13026}
\label{pos17:executive-order-13026}

1996 was also the year when Bill Clinton issued Executive Order 13026, which allowed the widespread use of strong cryptography. Under the narrative: NSA already had TQNN cryptanalysis capability (1995), so public-key encryption was no longer worth restricting. The crypto wars ended because the government had already won --- they just could not say how.

Under Possibility~A, the timing of EO~13026 is coincidence --- governments adjust export policy for many reasons, and deregulation was already underway. Under~B, there may have been a program, but perhaps not with the full cryptanalytic capability described here. Under~C, the implications are stark: every encrypted email sent after 1996, every SSL transaction, every PGP-signed message --- an illusion of privacy, maintained by the same government that held the means to read it all. The UDHR guarantees the right to privacy. If the story is true, that right was violated for every connected person on Earth, continuously, for decades --- concealed by the appearance of protection.

\chapterreturn
