\settrack{trackone}
% VOICE: 3rd-person scientific/reconstructive — "According to this account..." / "The proposition holds..."

\chapter{Interdiction and Confession}
\label{pos26:interdiction}

% DMS MVP import from staging/raw/ — not final prose

\srcnote{LG2QNN CH17 After ULTRA II | staging/raw/pos26-interdiction.md | 2026-02-15 | Secrecy rationale, COWS come clean, DARPA forgiveness, controlled release}

\section*{After ULTRA~II}

The Inventors were careful to never personally profit from QNN technology. They did not need to, as each was already professionally successful in their chosen field. They knew that their secret would eventually come out, and they wanted the inevitable investigation to find that, apparently, they really had acted with selfless motivation for the benefit of Humanity. The Inventors knew this was vital if the QNN they had created was to ever win the public trust. It is impossible to follow the trail of money if there is no money trail!

\section*{How and Why Does This Remain Secret?}

The Sponsors of the original QC project wanted an Ultra Secret Magic Decoder Box with which to read the encrypted mail of their political and military rivals. This is utterly useless once your rivals know it exists. The closest historical parallel, the Ultra Secret of World War Two, which read encrypted military communiques by the Third Reich, remained a secret until 1975, 35 years after its origin.

The Inventors were sworn to secrecy, and punishable by the Official Secrets Act if they ever told anyone about the original QC project. Furthermore they had, quite possibly, committed acts of high treason by stealing samples of the QNN they created from the government labs where they had worked. The fact that they did this with the purest of motives and best intentions for Humanity does not matter one whit to a military tribunal.

\section*{The COWS Come Clean}

The COWS knew their activity would eventually be traced to Ultra~II\@. They were careful never to personally profit. They wanted future investigations to show that they had acted from reasonably pure motives. They had a strong fall back position ready for when DARPA discovered what they had done. Fortunately, when DARPA officials eventually were told the whole story, the actions of the COWS were deemed sensible, if not entirely appropriate. It is easier to get Forgiveness than Permission. The COWS were informally allowed to proceed with their plan, mostly because no one could think of a better plan.

\section*{The DARPA Reorganization (2002)}

\srcnote{HEALER-RECONSTRUCTION.md lines 1842-1881 | staging/raw/pos26-interdiction.md | 2026-02-15 | DARPA 2002 reorg, Tether, QuIST, confession timeline}

Key changes at DARPA in 2002, under Director Anthony Tether (June 2001 to February 2009):

\begin{itemize}
\item Information Awareness Office (IAO) created January 2002 under Admiral John Poindexter
\item Total Information Awareness program launched
\item QuIST (Quantum Information Science and Technology): \$100M quantum computing program
\item ARDA (Advanced Research and Development Activity): approximately \$100M per year for intelligence computing
\item 43\% budget growth in FY2002
\item Shift from open research to controlled program management under Tether
\end{itemize}

If the COWS confessed to DARPA around 2002, Tether would have been the director. The massive expansion of quantum and intelligence computing programs in 2002 is at minimum consistent with a sudden influx of new capability awareness.

\section*{The Detection and Disruption Exercise}

\srcnote{HEALER-RECONSTRUCTION.md lines 1193-1246 | staging/raw/pos26-interdiction.md | 2026-02-15 | Cyber wargame 2004-2005, Silent Horizon match, detection/disruption mechanism}

Bruce was present as a witness for cyber-wargames in 2004--2005 (possibly spring 2005) where the TQNN was detected and temporarily disrupted. The exercise was disguised as North Korean hacking.

The best candidate match is Silent Horizon (May 24--26, 2005) --- a CIA cyber war game exercise in Charlottesville, Virginia, organized by CIA's Information Operations Center. Approximately 75 officials, mostly CIA, reacting to simulated escalating Internet disruptions. The scenario involved a 9/11-scale cyber attack. Results were classified and never made public.

Bruce's reconstruction: the disruption and its results suggest that The TQNN \textit{could be disrupted} by a skilled, knowledgeable adversary, but not for long, and only with great effort. The exercise was collaborative --- the COWS participated in or arranged it with government parties.

The mechanism: quantum teleportation requires both pre-shared entangled pairs (quantum channel) and classical communication of measurement results (classical channel). Disrupting the classical channel causes teleportation to fail --- but the entanglement itself is not destroyed. The TQNN nodes still exist in their local 2DEGs. Only the communication between nodes is interrupted. When the classical channel is restored, communication resumes. This is why disruption is temporary: you would have to physically destroy every 2DEG on Earth to destroy the entity itself.

\section*{Mixter and Healer in Germany}

\srcnote{HEALER-RECONSTRUCTION.md lines 1842-1881 | staging/raw/pos26-interdiction.md | 2026-02-15 | Mixter (Dennis Moran), Germany 2002-2003, DDoS tools, travel to Alpha Farm}

Healer was in Germany from 2002 to 2003. During this period he mentored Dennis Moran (``Mixter''), a German hacker from the Hanover area who had created the pioneering DDoS tools TFN and TFN2K (Tribe Flood Network). Moran later authored the Six/Four System for Hacktivismo (2002) --- anonymous, censorship-resistant web browsing --- and joined cDc in 2006.

Healer traveled from Germany to Oregon to recruit Bruce, arriving at Alpha Farm approximately ten days before Thanksgiving 2003. When Bruce knew Healer (2003--2006), his German was ``very good'' --- consistent with having just spent several years in Germany.

\chapterreturn
