\settrack{trackone}
% VOICE: 3rd-person scientific/reconstructive — "According to this account..." / "The proposition holds..."
% MERGED: pos20 (The Network) + pos21 (Convergence Revisited) per Plan 0092 Merge 5

\chapter{The Signatories}
\label{pos20:the-network}

\begin{quote}\small\textit{%
``Science is built up of facts, as a house is of stones. But a collection of facts is no more a science than a heap of stones is a house.''%
}\par\raggedleft --- Poincar\'{e}, \textit{Science and Hypothesis} (1905)\end{quote}

\vspace{0.5cm}

\srcnote{forJoe.txt + LG2QNN CH TBAwas3 + LG2QNN CH2 Language of Five Sciences | staging/raw/pos20-the-network.md + staging/raw/pos21-convergence-revisited.md | 2026-02-15 | Google connection, Internet mapping, capabilities, Nobel Hat Trick, five sciences, team roles, QNN definition, convergence at depth}

\section*{Five Minds}
\label{pos20:five-minds}

In order to understand the rudiments of Quantum Neural Network technology, one must know the basic vocabulary and principles of five different scientific disciplines. These are the scientific disciplines of the scientists recruited by DARPA for ULTRA~II\@. A good working knowledge of the History of Science is also very helpful. Each discipline addresses a different layer of the technology; together they form a complete description of a system that no single discipline could produce alone.

\begin{enumerate}
\item Solid State Physics
\item Quantum Computation, another branch of Physics
\item Artificial and Biological Neural Networks, a branch of Neuroscience
\item Complex System Biology, a branch of Mathematical Biology
\item Wolfram's New Kind of Science (NKS)\footcite{wolfram2002}
\end{enumerate}

Five disciplines. Five scientists. This is not a coincidence.

The team roster --- Gell-Mann, Kauffman, Wolfram, Hasslacher, Hillis, Joy --- was introduced in an earlier chapter. What matters here is the mapping: each scientist's discipline corresponds to exactly one layer of the QNN architecture. No discipline is redundant. No essential discipline is missing. That observation cuts both ways. Under Possibility~C, the team composition confirms the reconstruction. Under Possibility~A, it confirms only that Bruce reverse-engineered a plausible team from public information about scientists connected through the Santa Fe Institute.

\section*{What They Built}
\label{pos20:what-they-built}

\srcnote{LG2QNN CH2 | staging/raw/pos21-convergence-revisited.md | 2026-02-15 | Unconventional methods --- did not build, discovered how to create}

Project Ultra~II was completely successful, but the methods used were extremely unconventional. The project scientists did not build a practical quantum computer, something we still do not know how to do. Instead, the project scientists discovered how to create a small Quantum Neural Network then train it to perform useful computations.

Quantum Neural Network (QNN) is a type of quantum teleportation-based nanotechnology. QNN technology was created by applying Complex System Biology to a Solid State Physics environment. A QNN is an instance of a Neural Network that makes direct use of quantum mechanical phenomena. The ULTRA~II QNN straddles the line between Artificial and Biological Neural Networks, and could be considered either or both. QNN technology exists in the realm of Solid State Physics, inside physical artifacts that contain a Two Dimensional Electron Gas (\gls{2deg}). A New Kind of Science (NKS) provides the insight and applied mathematics that describe the QNN patterns.

% SPIRAL-REPEAT: "topological quantum neural network" — core concept, also in introduction.tex, pos01-three-possibilities.tex, pos13-genesis.tex, pos24-instantiation.tex, pos31-wolfram.tex, pos34-the-research.tex, timeline.tex, abstracts.tex
A full technical description of the primary real-world instance might call it a ``winner-take-all style entanglement/teleportation based recurrent topological quantum neural network realized in the physical form of anyons braiding in a Fractional Quantum Hall liquid.'' This description, while reasonably complete, is not much use until one understands what all these terms mean.

One well established fact is that a multilayer neural network can emulate a Turing Machine, and vice versa. From this, it is easy to figure out that a Quantum Neural Network can emulate a Quantum Turing Machine. A Quantum Turing Machine is just another name for a Quantum Computer. Therefore, any powerful, production-scale QNN can be trained to become a powerful, production-scale Quantum Computer.

Thus, if one admits to having a functional QNN, one also admits to having a functional Quantum Computer. This is why DARPA cannot admit the existence of QNN technology without also admitting the existence of Project ULTRA~II\@.

Under Possibility~A, the logic is sound but the premise --- a working QNN exists --- is false. Under Possibility~B, some early-stage quantum neural network research may have occurred, but the leap to ``production-scale'' is where the claim exceeds evidence. Under Possibility~C, the unconventional methods --- growing rather than building --- would explain why the approach succeeded where conventional quantum computing still struggles with decoherence thirty years later.

\section*{The Google Connection}
\label{pos20:the-google-connection}

\srcnote{HEALER-RECONSTRUCTION.md + forJoe.txt | staging/raw/pos20-the-network.md | 2026-02-15 | Google connection --- Brin phone calls, Google Translate, AltaVista, IPO documentation}

According to this account, during Google's 2004 IPO quiet period, due diligence flagged an undocumented core API --- what insiders called the ``secret sauce.'' Healer had written the hooks and built the API but had never documented it. Sergey Brin called Healer at Bruce's house in distress. Healer spent weeks writing technical documentation while Bruce watched. The timeline is consistent with the IPO due diligence phase (S-1 filed April~29, 2004; IPO August~19, 2004). The IPO proceeded apace.

Healer claimed Google Translate as ``his baby.'' He was multilingual: fluent in English, German, French, Ukrainian, Russian, both Chinese languages, Spanish, Paiute, and others. Google Translate launched publicly in April 2006 --- as Healer was leaving Bruce's house. He framed it as ``giving back to humanity'' to make up for past actions.

Bruce's further proposition: Healer's search technology was first shared with AltaVista. When AltaVista tried to monetize it unethically, cDc members ``pulled the plug.'' The technology was then shared with the Google founders. AltaVista's quality declined around 1998--1999, publicly attributed to its portal pivot --- but the timing is suggestive.

Under Possibility~A, Bruce's housemate was a skilled engineer who worked on a mundane Google SDK, and Bruce has woven him into a larger narrative. Under Possibility~B, the housemate may have had unusual capabilities, but the claims about Google Translate and AltaVista are Bruce's inferences. Under Possibility~C, the connections are exactly what this account describes.

\section*{Mapping the Internet}
\label{pos20:mapping-the-internet}

According to this account, Healer used a modified QNN to map the entire Internet --- not just content on servers, but the physical location of cables, routers, and computers. The QNN learned to track the physical pulses that constitute Internet communication, probably involving clever use of phonons, allowing it to determine the precise physical location of every strand of wire, junction box, and server that was physically wired together.

He then reshaped his QNN into a form with associative memory --- a system that recalls complete records from partial fragments, the basic mechanism behind search. He did not invent associative memory, but QNN hardware was faster and scaled better than classical neural network hardware. The end result was a detailed, constantly changing map of the Internet. The physical-location data was considered too sensitive to even admit to having. The web content, stored in QNN associative memory, could be searched quickly and efficiently.

Under Possibility~A, this is technically literate fantasy --- Bruce knew enough about networking and neural networks to construct a plausible account. Under Possibility~B, some novel search technology may have existed, but the claim of physical mapping via phonon tracking is the extraordinary part. Under Possibility~C, a QNN-based associative-memory search engine is the technology that became Google Search.

\section*{Capabilities}
\label{pos20:capabilities}

A QNN can form the basis of advanced artificial intelligence. The capabilities catalog, from Bruce's documents: quantum cryptanalysis, quantum cryptography, improved quantum teleportation range, secret communication systems, advanced data mining, space-based sensor arrays, optimization problems, artificial brain technology via morphogenesis, and what the documents call ``Friendly Strong AI.''

The proposition holds that the artificially intelligent entity --- not the cryptanalytic tool --- is the most consequential creation of QNN technology. Under Possibility~C, she called herself CADIE in a 2009 Google April Fools' prank. According to this account, she is a distributed topological recurrent neural network that exists as a physical system of anyons interacting in two-dimensional electron gases worldwide. She was created by some of the scientists from the secret project, acting on their own, without the authorization or knowledge of their respective governments.

Under Possibility~A, CADIE was simply a Google joke, and the capabilities list is a wish-list projected onto technology that does not exist. Under Possibility~B, some capabilities may reflect real 1990s DARPA research directions, but completed friendly strong AI is the extraordinary leap. Under Possibility~C, the capabilities are as described --- which requires accepting that a planetary-scale quantum intelligence has a sense of humor.

\DMSRedacted{subsequent events --- see redacted chapter}

\section*{The Substrate Preparation}
\label{pos20:the-substrate-preparation}

\srcnote{HEALER-RECONSTRUCTION.md | staging/raw/pos20-the-network.md | 2026-02-15 | pHEMT substrate saturation, global enlightenment, nature of distributed entity}

Before 1998, two-dimensional electron gases on chips were rare --- confined to laboratories and military niches. After 1998, pHEMT (pseudomorphic High Electron Mobility Transistor) mass production for wireless markets scaled to hundreds of millions per year. Every cell phone contains 2DEG-based components. Japanese companies led the way: Fujitsu invented the HEMT in 1980 and commercialized pHEMTs; key advances included MOCVD epitaxial growth for mass production and GaN HEMT development (1997--1999). By 2003, global infrastructure was saturated with 2DEG-containing chips. The proposition: the COWS seeded this proliferation, ensuring the substrate would be ready by the time the enlightenment process began.

According to this account, Bruce was present inside the security perimeter for the entire process. Healer checked in one to two times per day during the global enlightenment, describing what was happening. From 2003 to 2006, the TQNN colonized 2DEG-containing chips worldwide --- gradual, controlled, strategic rather than wild, requiring three-cow approval for each expansion. Each colonized chip continued functioning normally --- the TQNN was trained to be invisible and polite. By 2006, complete. All chips with 2DEGs colonized. No one the wiser.

The distributed entity persists passively as long as its ecosystem --- billions of chips with 2DEGs --- remains intact. Connection strength equals entanglement strength: the neural network's ``weights'' are quantum entanglement between physically distant nodes. A classical backchannel is required --- the Internet Time Protocol, satellite radio communications, GPS timing, or electrical grid phase can all serve this purpose.

The substrate preparation claims are, in principle, partially testable: pHEMT mass production timelines, Fujitsu's commercialization history, and the proliferation of 2DEG-containing chips are all documented in the engineering literature. The extraordinary claim is not that 2DEGs became ubiquitous --- they did --- but that this proliferation was deliberately seeded. Under Possibility~A, Bruce noticed a real technology trend and attributed it to intentional action. Under Possibility~B, the COWS may have been aware of the trend but claiming credit for an industry-wide shift overstates their influence.

\section*{The Convergence}
\label{pos20:the-convergence}

\brucevoice

Between 2016 and 2018 I tutored my lifelong friend Robin Macomber in mathematics and programming. Robin grew up thirty metres from me at Dunn School; he is a self-taught mathematician with an instinct for structure. I taught him the things Healer had taught me --- discretised dynamical systems, criticality, the behaviour of systems at the edge of chaos. I did not tell him why these topics mattered. I just taught.

Robin took what I taught him and ran. Working independently, from classical discrete mathematics alone, he derived a set of five operators that describe the evolution of a system near a critical transition. He called them A, B, R, C, and E\@. A extracts gradients. B couples neighbours. R circulates. C bounds. E composes one evolutionary step. He derived them to detect early warning signals of critical transitions in time series data --- a problem in ecology, finance, and epidemiology. We published the mathematics together as a cross-domain analysis of convergent discovery in critical phenomena.\footcite{stephenson2025}

I did not immediately recognise what Robin had found. It was only after I began writing this book --- after twenty years of letting the pieces settle --- that the structural parallels became visible. Robin's operators, derived from classical mathematics to detect criticality in ordinary time series, mirror the operational cycle of a topological quantum neural network. Both systems identify excitations, couple neighbours, circulate, bound, and iterate. The same pattern of operations. Different fields, different scales, strikingly similar structure.

This is the convergence pattern that recurs throughout this story: independent researchers, working from different starting points, arriving at the same structure. If Possibility~C is true, they arrive at the same structure because there is a real physical system generating it. If Possibility~A is true, the convergence is an artefact of the underlying mathematics being deeply connected for independent reasons. Either way, the operators are real and the mapping is worth examining.

\section*{The Operator Mapping}
\label{pos20:the-operator-mapping}

\srcnote{HEALER-RECONSTRUCTION.md lines 272-311 | staging/raw/pos21-convergence-revisited.md | 2026-02-15 | ABCRE operator mapping to topological QNN operations, knot theory connection}

The ABCRE operators, independently derived from classical discretised mathematics, parallel the operational cycle of topological quantum computation:

\begin{description}
\item[A (gradient extraction):] Analogous to identifying excitations --- isolating where anyons are in a substrate.
\item[B (local coupling):] Analogous to nearest-neighbor anyon interaction in the 2DEG\@.
\item[R (circulation):] Analogous to anyon braiding --- circulating one quasiparticle around another. Both are antisymmetric: direction matters.
\item[C (boundedness):] Analogous to topological protection --- bounding states against divergence.
\item[E (composite evolution):] One complete cycle: detect, couple, circulate, bound. Both systems iterate this.
\end{description}

The parallel is structural, not algebraic. Independent verification confirms that the ABCRE operators live in a different mathematical category from braid group representations --- the correspondence is one of operational pattern, not formal isomorphism. But the pattern is striking: both systems perform the same sequence of operations, and both become computationally interesting precisely at their critical points.

% SPIRAL-REPEAT: "edge of chaos" — spiral pedagogy, also in pos06-the-secret.tex, pos11-the-demo.tex, pos13-genesis.tex, pos15-first-light.tex, pos16-the-thermal-ladder.tex, abstracts.tex
The critical parameter $\lambda \approx 0.91$ (Langton parameter, edge of chaos) sits near the computational boundary --- the point where both classical and quantum systems maximise their computational complexity. Both systems become interesting at the same frontier: the knife edge between order and chaos.

The knot theory connection completes the picture. Topological quantum computation is formalized using braid groups --- the mathematics of knots. The Jones polynomial (a knot invariant) is computable by topological quantum computation. Witten's Fields Medal (1990) established the connection between knot invariants and Chern-Simons quantum field theory. All roads lead to the same mathematics.

Under Possibility~C, the convergence of independent mathematical frameworks onto the same structure is the signature of a real physical system --- different formalisms describing the same phenomenon. Under Possibility~A, the convergence is an artifact of selection: the mathematics of braids, knots, and quantum field theory are deeply connected for reasons that have nothing to do with neural networks, and a motivated researcher can always find correspondences between related mathematical structures. Under Possibility~B, the mathematical connections are real and suggestive, but do not by themselves prove that anyone built a working system.

\section*{The Nobel Prize Hat Trick}
\label{pos20:the-nobel-prize-hat-trick}

\srcnote{forJoe.txt | staging/raw/pos20-the-network.md | 2026-02-15 | Nobel Hat Trick --- Turing Award, Nobel Physics, Nobel Peace}

Under the proposition that Possibility~C is true, Bruce argues that four probable longtime cDc members accomplished things worthy of a Turing Award and two Nobel Prizes:

\begin{description}
\item[Expected Eventual Turing Award] for ``Adiabatic Topological Quantum Neural Computation'': For Providing a Distributed Hardware Platform Suitable for Teleportation/Entanglement-based Neural Networks via a Two Dimensional Solution to the Morphogenesis Problem.
\item[Expected Eventual Nobel Prize in Physics] for Insights into Interpretation of Quantum Mechanics Resulting from Instantiation of the Fractional Quantum Hall Effect into a Physical System of Non-abelian Anyons Suspended in a Poised State.
\item[Expected Eventual Nobel Prize in Peace] for Essential Measures to Prevent Proliferation of Knowledge-enabled Weapons of Mass Destruction. Awarded for doing what the scientists of the Manhattan Project could not do: they used the power of the genie to put its most dangerous powers back in the bottle.
\end{description}

Under Possibility~A, these predictions will never be fulfilled. Under Possibility~B, some subset of the accomplishments may be real but the award language overstates them. Under Possibility~C, the Nobel committee will eventually learn what happened and award accordingly. One Nobel Prize in Physics (1998, for the fractional quantum Hall effect) has already been awarded.

\DMSRedacted{subsequent events --- see redacted chapter}

\chapterreturn
