% Evidence Interlude
% Maugham revision: evidence assessment after narrative, not before
% Content from: summary.tex "The Evidence" section + introduction.tex "What the book does"
% The reader has now encountered the evidence in narrative form.
% This section becomes a structured summary of what they already know.

\settrack{trackbridge}

\chapter*{The Evidence}
\addcontentsline{toc}{chapter}{The Evidence}

You have now read the story and the reckoning. Before the book turns to implications, a pause for assessment.

This story makes extraordinary claims. Extraordinary claims require extraordinary evidence. Here is what exists --- both for and against.

\vspace{0.5cm}

\textbf{Evidence that supports the story:}

\begin{itemize}
\item The Dutch government's official investigation (known as NIOD) independently confirmed that SAS soldiers were present at Srebrenica in July 1995. This is consistent with Healer's claimed HALO jump into the massacre zone. It does not prove he was there, but the type of operation he described did happen.
\item Patrick Ball, a statistician, is confirmed as both an advisor to the hacker group Cult of the Dead Cow AND the first expert witness at the International Criminal Tribunal for the former Yugoslavia. During the trial of Slobodan Milosevic in March 2002, Milosevic himself cross-examined Ball about the ``Dead Cow Cult.'' This unlikely connection between hackers and international war crimes prosecution --- which sounds absurd on its face --- is documented in official court transcripts.
\item Danny Hillis, another named scientist, co-founded a company called Applied Minds with Bran Ferren. Ferren has confirmed advisory roles with the CIA, NSA, the Director of National Intelligence, and the National Reconnaissance Office. He holds the Intelligence Community Seal Medallion. Applied Minds served every branch of the US military.
% SPIRAL-REPEAT: "Why the Future Doesn't Need Us" — key reference, also in pos07-the-departure.tex, pos06-the-secret.tex, pos27-extension.tex, timeline.tex, abstracts.tex
\item Bill Joy, co-founder of Sun Microsystems, published a famous and terrifying article in \textit{Wired} magazine in the year 2000 --- ``Why the Future Doesn't Need Us'' --- about the danger of technology threatening humanity. He described being told something by ``another person'' that deeply disturbed him. Its timing and emotional tone are consistent with the relinquishment narrative.
\item The GCHQ precedent --- keeping a world-changing cryptographic breakthrough secret for twenty-four years --- proves that this kind of classification is not just possible but has actually happened before.
\end{itemize}

\textbf{Evidence against the story:}

\begin{itemize}
\item There is only one source. Everything comes from one mentor. No other member of the team has stepped forward to confirm any part of the story.
\item There is no physical evidence. No one has produced a quantum computer, a piece of Custodian, or any artifact from the program.
\item The story has a troubling quality: it absorbs all contradictory evidence. Any lack of proof can be explained by ``it's classified.'' Any missing corroboration can be explained by ``people are afraid to talk.'' This makes parts of the story difficult or impossible to disprove --- which is a warning sign for any critical thinker.
\item Extraordinary claims require extraordinary evidence, and the evidence, while genuinely suggestive, is entirely circumstantial. None of it PROVES the story. All of it is consistent with other explanations.
\end{itemize}

There is no proof. There is a mountain of circumstantial evidence. The reader must weigh it for themselves.

\vspace{1cm}\noindent\rule{0.5\textwidth}{0.5pt}\vspace{0.5cm}

This book uses real names. Not for spectacle --- for falsifiability. Anonymized propositions cannot be checked. Fictional propositions carry no predictive weight. The names are the load-bearing structure that makes the claims testable.

The book makes specific, falsifiable predictions. If Possibility~C is correct, specific events should occur in specific timeframes. These are documented in Appendix~B with dates. They cannot be retro-fitted. Every confirmation strengthens the book's credibility. Every failure is absorbed by the three-possibilities framing. This asymmetry is deliberate.

If the predictions fail, you have your answer. If they don't, you have a different one.

\chapterreturn
